### 2019-01-30 manuscript/manuscript.tex
first commit

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\author{
FirstName LastName\\University/Company \And Second Author\\Affiliation
}
\title{A Capitalized Title: Something about a Package \pkg{foo}}

\Plainauthor{FirstName LastName, Second Author}
\Plaintitle{A Capitalized Title: Something about a Package foo}
\Shorttitle{\pkg{foo}: A Capitalized Title}

\Abstract{
The abstract of the article.
}

\Keywords{keywords, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, not capitalized, Java}

%% publication information
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{}
%% \Acceptdate{2012-06-04}

\Address{
    FirstName LastName\\
  University/Company\\
  First line Second line\\
  E-mail: \email{name@company.com}\\
  URL: \url{http://rstudio.com}\\~\\
    }

% Pandoc header

\usepackage{amsmath}

\begin{document}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

This template demonstrates some of the basic latex you'll need to know
to create a JSS article.

\hypertarget{code-formatting}{%
\subsection{Code formatting}\label{code-formatting}}

Don't use markdown, instead use the more precise latex commands:

\begin{itemize}
\item
  \proglang{Java}
\item
  \pkg{plyr}
\item
  \code{print("abc")}
\end{itemize}

\hypertarget{r-code}{%
\section{R code}\label{r-code}}

Can be inserted in regular R markdown blocks.

\begin{CodeChunk}

\begin{CodeInput}
R> x <- 1:10
R> x
\end{CodeInput}

\begin{CodeOutput}
 [1]  1  2  3  4  5  6  7  8  9 10
\end{CodeOutput}
\end{CodeChunk}

\end{document}

### 2019-01-31 manuscript/manuscript.Rmd
add content

---
author:
  - name: Nan-Hung Hsieh
    affiliation: Texas A&M University
  - name: Brad Reisfeld
    affiliation: Colorado State University 
  - name: Weihsueh A. Chiu
    affiliation: Texas A&M University
    address: >
      4458 TAMU, 
      College Station, TX 77843, USA
    email: \email{wchiu@cvm.tamu.edu}  
title:
  formatted: "\\pkg{pksensi}: An R package to apply sensitivity analysis in pharmacokinetic models"
  # If you use tex in the formatted title, also supply version without
  plain:     "pksensi: An R package an R package to apply sensitivity analysis in pharmacokinetic models"
  # For running headers, if needed
  short:     "**pksensi**: Sensitivity analysis for pharmacokinetic models in R"
abstract: >
  Sensitivity analysis is an essential tool for modelers to understand the influence of model parameters on model outputs. It is also increasingly used in developing and assessing pharmacokinetic models. In our previous work, we applied a global sensitivity analysis workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model. Although several sensitivity analysis algorithms are available, no comprehensive package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize sensitivity analysis results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration. Therefore, we developed an open-source R package, called **pksensi**, to fill this need and make sensitivity analysis more accessible in pharmacological and toxicological research. This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK model with multivariate model outputs (multiple time-points and tissue compartments). It also refined the extended Fourier Amplitude Sensitivity Test method for sensitivity analysis by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter. Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement. Utilizing pksensi, we successfully reproduced our previously published sensitivity analysis results of human PBPK modeling for acetaminophen and its two primary metabolites. Overall, pksensi improves the user experience of performing sensitivity analysis and can create robust and reproducible results for decision making in pharmacokinetic model calibration. 
keywords:
  # at least one keyword must be supplied
  formatted: [Pharmacokinetics, Physiologically based pharmacokinetics, Sensitivity analysis]
  plain:     [Pharmacokinetics, Physiologically based pharmacokinetics, Sensitivity analysis]
preamble: >
  \usepackage{amsmath}
output: rticles::jss_article 
#output: rmarkdown::html_vignette
---

# Introduction

Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs. An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions (Ferretti et al., 2016). Also, this technique has been applied in pharmacology and toxicology research (Loizou et al., 2015; McNally et al., 2012; Scherholz et al., 2018). Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time. The goal of sensitivity analysis in pharmacokinetic research is to examine the sensitivity of output variables (e.g. compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants (McNally et al., 2011). It can be further applied to parameter prioritization and parameter fixing before model calibration (Hsieh et al., 2018).

In our previous work (Hsieh et al., 2018), we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model. We used GNU MCSim (Bois, 2009), an effective simulation package for Bayesian population PBPK modeling, to calibrate the model. We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of global sensitivity analysis algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model. Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters. We also found a useful approach for communicating the parameter sensitivity in decision making.

Most of sensitivity analysis tools are constructed using scientific computing software, such as GNU Octave/Matlab (Pianosi et al., 2015), python (Herman and Usher, 2017), and R (Ghanem et al., 2017). Previously, we mainly used R with the sensitivity package (Pujol et al., 2017), which is a practical tool to conduct local and global sensitivity analysis. The sensitivity package includes some functions to generate the parameter sequences by using different computing algorithms. Also, it provides a feasible way to integrate external modeling results. This computational approach can effectively solve the heavy computing burden of numerical solutions within the pure R environment. Several R packages, such as FME, multisensi, and ODEsensitivity include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network. The FME package contains a basic method for performing local sensitivity analysis for dynamic models (Soetaert and Petzoldt, 2016). The multisensi package provides sensitivity analysis for models with multivariate output (Bidot et al., 2018). ODEsensitivity can perform sensitivity analysis in ordinary differential equation (ODE) models (Weber et al., 2018).  It uses the ODE interface from the deSolve package, connecting it with the sensitivity analysis from the sensitivity package, and can perform sensitivity analysis on models with multivariate output. Although these tools provide various approaches to perform sensitivity analysis, we have not find a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.

Here we develop an R package, called pksensi, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches. This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK, and advanced compartment absorption and transit models with multivariate model output. The design concepts of pksensi are:

*	Cross-platform: Models can run on Windows/MacOS/Linux
*	Freely available: All related packages are free and open source
*	Integrated application: Users can run pharmacokinetic models in R with script that were written in C or GNU MCSim 
*	Decision making: The output results and visualization tools can be used to easily determine which parameters have “non-influential” effects on the model output and can be fixed in model calibration.

# Workflow

## Installation

\pkg{pksensi} is available from the comprehensive R Archive Network at https://CRAN.R-project.org/package=pksensi. The latest up-to-date version is available through GitHub at https://github.com/nanhung/pksensi.

## Parameter matrix generation

We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package (Saltelli et al., 1999). Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the R \pkg{sensitivity} package (Pujol et al., 2017). To test the convergence and robustness of the sensitivity measurement, we included a random phase-shift approach to replicate sampling from random starting points across parameter space. This can be written as:

$$ x_i = \frac{1}{2} + \frac{1}{\pi}\arcsin(\sin(\omega_is + \varphi_i)) $$

where $x_i$ is the nominal value of the i-th parameter, and $\omega_i$  is a vector giving the set of frequencies, one frequency for each parameter. The default set of frequencies is based on the study from Saltelli et al. (1999). The first frequency of the vector $\omega$ is assigned to each factor $x_i$ in turn (corresponding to the estimation of Sobol indices of first and total order, respectively). Through the random phase-shift, the robust result of the sensitivity measurement should be similar across replications under the same sample size.

To investigate the convergence of sensitivity indices, we adopted the exam approach proposed by Sarrazin et al. (2016). This method quantitatively assesses convergence by computing the width of 95% confidence intervals for all parameters across all time-points and output variables. To integrate the random phase-shift with eFAST, we built the \proglang{rfast99} function, which can be used to perform the sensitivity test for the whole pharmacokinetic modeling process. The function \proglang{rfast99} in pksensi can sample and generate the testing parameter matrix based on a given argument of sample size n and probability distribution (e.g., mean/s.d. of normal distribution, meanlog/sdlog of log-normal distribution, and minimum/maximum of uniform distribution). The number of model evaluations is equal to the sample size times the number of model parameters.

## Pharmacokinetic modeling

The \pkg{pksensi} provides a useful function to solve ordinary differential equations in pharmacokinetic models using analytical and numerical approaches. The solution can be performed under the pure R coding environment by linking pksensi with the deSolve package (Soetaert et al., 2018). However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as FORTRAN, C, or C++. The \pkg{pksensi} includes a function to compile and create dynamic-link libraries (.dll) on Windows and shared objects (.so) on Unix-liked systems (e.g., Linux and MacOS). This compiled file can load and execute through the build-in function in R. However, Windows users need to install Rtools or MinGW to compile the source code by using the GNU GCC compiler. More requirements are detailed in the study from Soetaert et al. (2010).  The C implementation of the example model can be found as an example within MCSim v6.0.1 (Bois, 2009). The \pkg{pksensi} can also link with MCSim to compile the model code, used in solving each system of equations.

## Output visualization and decision making

The output of sensitivity analysis returns a list that contains the given time-points and values of all state variables (e.g., chemical concentrations in blood). This format is particularly suited for further graphical routines within pkseni. Also, a plot can can create time-dependent sensitivity measurements of first and total order effect for all parameters in one figure. Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement. We also developed a “cut-off”-based approach in this package to accurately distinguish between “influential” and “non-influential” parameters. Finally, pksensi includes numerous visualization tools for the effective investigation and communication of results in decision making.

# Examples

## One-compartment taxicokinetic model

```{r setup, include = FALSE}
library(pksensi)
#mcsim_install()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.height=4, 
  fig.width=6
)

fn = local({ # not used function for fig caption
  i = 0
  function(x) {
    i <<- i + 1
    paste('Figure ', i, ': ', x, sep = '')
  }
})

```

### Equations

In this example, We use a simple, one-compartment PK model from `httk` package [@JSSv079i04] to demonstrate how `pksensi` can be applied to pharmacokinetic studies. The differential equations for the one-compartment pharmacokinetic model can be written as:

$$\frac{dA_{gutlumen}}{dt} = -k_{gutabs} \cdot A_{gutlumen} + g(t)$$
$$\frac{dC_{rest}}{dt} = \frac{k_{gutabs}}{V_{dist}}-k_{elim} \cdot C_{rest}$$

where $A_{gutlumen}$ is the state variable that describes the quantity of compound in gut lumen (mol), $k_{gutabs}$ is the absorption rate constant that describes the chemical absorption from the gut lumen into gut tissue through first-order processes (/h), $V_{dist}$ is the volume of distribution (L), and $k_{elim}$ is the elimination rate constant (/h), which is equal to the total clearance divided by the volume of distribution. The time-dependent function $g(t)$ is used to describe the oral dosing schedule. $C_{rest}$ is the chemical concentration in plasma that can be used to compare with observed results in a pharmacokinetic experiment (mol/L).

### Model implementations with R deSolve package

In the beginning, we need to pre-install GNU MCSim [@JSSv002i09] and related compiler to let us generate .c file and executable file. The GNU MCSim can be installed by following the instruction in GNU MCSim’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function `mcsim_install()` in `pksensi`. The GNU compiler is necessary for users that use Linux or MacOS. For Windows users, you should install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and use `Sys.setenv()` to set the working path of compiler. The  `Sys.which("gcc")` and `system('g++ -v')` can check whether we can run compiler correctly.

We first implemented this model in R by compiling the file written in C. pksensi allows users to select the preferred method to solve the pharmacokinetic model, either with the `deSolve` package or with GNU MCSim through the compile function. However, running model under GNU MCSim native code can have faster speed to obtain the model outputs.

The following R script can download and compile the model description file (`pbtk1cpt.model`) and use `deSolve` package [@JSSv033i09] to solve ordinary differential equations in our model. The example model code of one-compartment PBTK model is available with pksensi package:

```{r, eval=T}
pbtk1cpt_model()
cat(readLines("pbtk1cpt.model"), sep = "\n")
```

Then, use `compile_model()` to generate the executable files (`pbtk1cpt.dll` on Windows or `pbtk1cpt.so` on other systems) and R file (`pbtk1cpt_inits.R`) with default input parameters and initial state settings with the definition of `application = "R"`.

```{r, eval=T}
mName <- "pbtk1cpt"
compile_model(mName, application = "R")
source(paste0(mName, "_inits.R"))
```

The parameter values and initial states can be customized to specify the properties and schedule for the given dosing scenario.

```{r, eval=T}
parms <- initParms()
parms["vdist"] <- 0.74
parms["ke"] <- 0.28
parms["kgutabs"] <- 2.18
initState <- initStates(parms=parms)
initState["Agutlument"] <- 10
```

```{r, eval=T}
parms
```

```{r, eval=T}
initState
```

```{r, eval=T}
Outputs
```

In the current setting, we assumed the initial condition of the intake chemical to be 10 mol. The `initParms` and `initStates` functions were used to customize the parameter values and the initial state that will be used in the `solve_fun` function. These parameter value can be adopted from the `httk` package, which includes physico-chemical and drug biological properties for 553 chemicals. In this case, we used the parameter value of theophylline in this example. The given `vdist`, `ke`, and `kgutabs` are 0.74, 0.28, and 2.18, respectively.

Through `ode` function in `deSolve` package, we can visualize the pharmacokinetic according to the given parameter conditions such as time points (`times`):

```{r, eval=T}
times <- seq(from = 0.01, to = 24.01, by = 1)
y <- deSolve::ode(initState, times, func = "derivs", parms = parms, 
                  dllname = mName, initfunc = "initmod", nout = 1, outnames = Outputs)
```

```{r, eval=T, fig.cap = 'Figure 1. Simulation results of one-compartment PBTK model.'}
plot(y)
```

To conduct sensitivity analysis for the parameters in one-compartment pharmacokinetic model in this case, we want to quantify the impact of these three parameters on the chemical concentration in plasma during 24-hour time period post intake. We assume a uniform distribution for the estimate for each parameter with the coefficient of uncertainty within 50%. The parameter ranges are assumed to be (0.37, 1.12) for `vdist`, (0.0058, 0.0174) for `ke`, and (0.045, 0.136) for `kgutabs`. The sample number determines the robustness of the result of sensitivity analysis. Higher sample numbers can generate narrow confidence intervals for sensitivity measurements across different replications. However, they might cause heavy computational burden for complex models. Here we use a sample number of 400 with 20 replications:

```{r, eval=T}
LL <- 0.5 
UL <- 1.5
q <- "qunif"
q.arg <- list(list(min = parms["vdist"] * LL, max = parms["vdist"] * UL),
             list(min = parms["ke"] * LL, max = parms["ke"] * UL), 
             list(min = parms["kgutabs"] * LL, max = parms["kgutabs"] * UL)) 
set.seed(1234)
params <- c("vdist", "ke", "kgutabs")
x <- rfast99(params, n = 200, q = q, q.arg = q.arg, replicate = 20)
```

Because the pharmacokinetic model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements can also show the time-dependent relationships for each model parameter. Here we define the output time points to examine the change of the parameter sensitivity over time. To solve the pharmacokinetic model through deSolve, we need to provide the details of the argument:

```{r, eval=T}
y <- solve_fun(x, times, initState = initState,
               outnames = Outputs, dllname = mName)
tell2(x,y)
```

To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours in the example. The `initParmsfun` is used to generate the sampling value for each parameter. The `outnames`, `dllname`, `func`, `initfunc` are based on the arguments from the ode function in `deSolve` package. The details of the model structure and these arguments are defined in `pbtk1comp.c`. and `pbtk1comp_inits.R`. Finally, the `tell2` function is used to integrate the parameter values and the output results of numerical analysis that were generated and stored in variables x and y. The result of object x is an object of rfast99, which has specific `print`, `plot`, and `check` method. The print function gives the sensitivity and convergence indices for main, interaction, and total order at each time point. In addition to print out the result of sensitivity analysis, the more efficient way to distinguish the influence of model parameter is to visualize them. The time-dependent sensitivity indices are shown in Figure 2.

```{r, eval=T, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24 hour time period intake.'}
plot(x)
```

Here, we can find that `vdist` and `ke` are dominating the plasma concentration in the before and after 5-hour post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration. Besides, the `kgutabs` only plays a crucial role to determine the plasma concentration in the first hour. The relationship between concentration and the parameters can be plotted as follow (Figure 3):

```{r, eval=T, fig.cap = 'Figure 3. The relationship between model parameter and estimated concentration under the time-point of 0.01, 2.01, and 24.01 hr'}
par(mfrow = c(3,3), mar = c(2,2,2,2), oma = c(2,2,1,1))
plot(x$a[,1,"vdist"], y[,1,"0.01",], main = "vdist")
text(1, .7, "t=0.01",cex = 1.2)
plot(x$a[,1,"ke"], y[,1,"0.01",], main = "ke")
plot(x$a[,1,"kgutabs"], y[,1,"0.01",], main = "kgutabs")
plot(x$a[,1,"vdist"], y[,1,"2.01",])
text(1, 18, "t=2.01",cex = 1.2)
plot(x$a[,1,"ke"], y[,1,"2.01",])
plot(x$a[,1,"kgutabs"], y[,1,"2.01",])
plot(x$a[,1,"vdist"], y[,1,"24.01",])
text(1, .7, "t=24.01",cex = 1.2)
plot(x$a[,1,"ke"], y[,1,"24.01",])
plot(x$a[,1,"kgutabs"], y[,1,"24.01",])
mtext("parameter", SOUTH<-1, line=0.4, outer=TRUE)
mtext("Ccompartment", WEST<-2, line=0.4, outer=TRUE)
```

The x is a list of class "rfast99", containing all the input arguments detailed before and the calculated sensitivity indices of first order (`mSI`), interaction (`iSI`), and total order (`tSI`). The convergence indices are also stored in the list named `mCI`, `iCI`, and `tCI`. The parameter values are stored in an array `x$a` with c(model evaluation, replication, parameters).

```{r, eval=T}
dim(x$a)
```

In addition, the output are also formated with c(model evaluation, replication, time, variable).

```{r, eval=T}
dim(y)
```

The `check()` is a useful function to determine which parameters have relative lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration. The argument of `SI.cutoff` is setting at 0.5 to detect the relative non-influential parameters in this case.

```{r, eval=T}
check(x, SI.cutoff = 0.5)
```

Based on the sensitivity measurement of the total order, the result shows that `kgutabs` has relative lower measurement of sensitivity index.

### Model implementations with GNU MCSim

Alternatively, to solve ODE by using GNU MCSim, we need to change the argument to `application = mcsim` in `compile_model()`. Rather than apply R `deSolve` to solve differential equations, the GNU MCSim can provide higher computational speed in global sensitivity analysis.

```{r, eval=T}
system.time(y<-solve_fun(x, times, initState = initState, 
                         outnames = Outputs, dllname = mName))
```

```{r, eval=T}
compile_model(mName, application = "mcsim")
```

Sililiar to `solve_fun` that can define the initial parameter and state values through input function, the `solve_mcsim` has a `condition` argument that is used to givien the specific input value such as oral dose or fixing parameter value or initial state variable.

```{r, eval=T}
conditions <- c("Agutlument = 10") # Set the initial state of Agutlument = 10 
system.time(y<-solve_mcsim(x, mName = mName, 
                           params = params,
                           vars = Outputs,
                           time = times,
                           condition = conditions))
```

Under the same given condition, it takes 5-6 (`deSolve`) and 1-2 (GNU MCSim) seconds to solve model. The `solve_mcsim()` shows the better computational performance than `solve_fun()` in `pksensi`.

## Acetaminophen-PBPK model

### Uncertainty and sensitivity analysis

The aim of this section is to reproduce our previous published [@fphar201800588] result of global sensitivity analysis for acetaminophen PBPK model through `pksensi`. The model codes are included in this package and can be generated through `pbpk_apap_model()`. We applied the global sensitivity analysis workflow to the original published model with 21 model parameters [@s13318-015-0253-x]. The descriptions of each parameter and the sampling ranges are list in Table 1.

<Table 1>

Same as the example of one-compartment PBTK model. The model parameter and the corresponding sampling range should be defined to create the parameter matrix. Previously, the probability distributions of model parameters were set to either truncated normal or uniform distribution when the parameters have informative prior information or not. To rapidly reach the acceptance convergence, we apply uniform distribution for all testing parameters. The ranges of informative parameters are set to 1.96-times difference for single side (approximate 54.6 times difference between minimum and maximum) under log-scaled. The nominal values of informative model parameters were defined as:

```{r, eval=T}
# Nominal value
Tg <- log(0.23)
Tp <- log(0.033)
CYP_Km <- log(130)
SULT_Km_apap <- log(300)
SULT_Ki <- log(526)
SULT_Km_paps <- log(0.5)
UGT_Km <- log(6.0e3)
UGT_Ki <- log(5.8e4)
UGT_Km_GA <-log(0.5)
Km_AG <- log(1.99e4)
Km_AS <- log(2.29e4)

rng <- 1.96 
```

Generally, The wide range of parameter value might cause the computational error in the solver. One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value. The `generate_infile()` provide the arguments of `rtol` and `atol` that can be adjusted to prevent the unwanted error. However, the modification will slow down the computational speed. Therefore, the alternative method to prevent the computational error is to detect the crucial parameter range that causes the problem. Also, setting the maximum number of (internally defined) steps to higher value instead of using the default value (500) can prevent this problem. The maximum number of step is set to 5000 in this case.

In this test case, we adjusted the range of `SULT_VmaxC` and `UGT_VmaxC` from U(0, 15) to U(0, 10). The relative and absolute error tolerance were set to 1e-7 and 1e-9, respectively, to prevent the computational error in MCSim,

```{r, eval=T}
params <- c("lnTg", "lnTp", "lnCYP_Km","lnCYP_VmaxC",
           "lnSULT_Km_apap","lnSULT_Ki","lnSULT_Km_paps","lnSULT_VmaxC",
           "lnUGT_Km","lnUGT_Ki","lnUGT_Km_GA","lnUGT_VmaxC",
           "lnKm_AG","lnVmax_AG","lnKm_AS","lnVmax_AS",
           "lnkGA_syn","lnkPAPS_syn", "lnCLC_APAP","lnCLC_AG","lnCLC_AS")
q <- "qunif"
q.arg <-list(list(Tg-rng, Tg+rng),
             list(Tp-rng, Tp+rng),
             list(CYP_Km-rng, CYP_Km+rng),
             list(-2., 5.),
             list(SULT_Km_apap-rng, SULT_Km_apap+rng),
             list(SULT_Ki-rng, SULT_Ki+rng),
             list(SULT_Km_paps-rng, SULT_Km_paps+rng),
             list(0, 10),
             list(UGT_Km-rng, UGT_Km+rng),
             list(UGT_Ki-rng, UGT_Ki+rng),
             list(UGT_Km_GA-rng, UGT_Km_GA+rng),
             list(0, 10),
             list(Km_AG-rng, Km_AG+rng),
             list(7., 15),
             list(Km_AS-rng, Km_AS+rng),
             list(7., 15),
             list(0., 13),
             list(0., 13),
             list(-6., 1),
             list(-6., 1),
             list(-6., 1))

times <- seq(from = 0.1, to = 12.1, by = 0.2)
set.seed(1234)
x <- rfast99(params = params, n = 512, q = q, q.arg = q.arg, replicate = 10) 
```

After creating the `pbpk_apap.model` in the working directory, the next step is to generate the executable files (`mcsim.pbpk_apap`) through `compile_model()`.

```{r, eval=T}
mName <- "pbpk_apap"
pbpk_apap_model()
compile_model(mName, application = "mcsim")
```

To improve the computational speed, this case only uses MCSim to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma. The setting oral dose of APAP is 20 mg/kg in this example. Generally, the input dosing method can be defined through the `condition` argument. Since the unit of the given dose is mg/kg, the `mgkg_flag` is set to 1 to declare the statement. More definition of input can be found in the section of input functions in GNU MCSim User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).

```{r, eval=T}
vars <- c("lnCPL_APAP_mcgL", "lnCPL_AG_mcgL", "lnCPL_AS_mcgL")
conditions <- c("mgkg_flag = 1",
                "OralExp_APAP = NDoses(2, 1, 0, 0, 0.001)",
                "OralDose_APAP_mgkg = 20.0")
generate_infile(params = params,
                vars = vars,
                time = times, 
                condition = conditions,
                rtol = 1e-7, atol = 1e-9)
system.time(y <- solve_mcsim(x, mName = mName, 
                             params = params,
                             vars = vars,
                             time = times,
                             condition = conditions,
                             generate.infile = F))
tell2(x,y)
```

The plotting function can output the result of time-dependent sensitivity measurement to determine the parameter impact on model output over time (Figure 1).

```{r, eval=T, fig.height=8, fig.width=8, fig.cap = 'Figure 4. '}
plot(x, vars = "lnCPL_AG_mcgL")
```

The uncertainty analysis is a crucial step before model calibration. We can apply uncertainty analysis through the `pksim()` by the given name of output variables (Figure 2). Through this visualization approach, we can recognize whether the simulated outputs can accurately simulate the same concentration profile as the in-vivo experiment under the setting of parameter ranges. In Figure 2, all experiment points are included in the intervals, representing the acceptable set of the parameter range.

```{r, eval=T, fig.cap = 'Figure 5. '}
par(mfrow = c(2,2), mar = c(2,2,1,1), oma = c(2,2,1,1))
pksim(y, vars = "lnCPL_APAP_mcgL")
text(1, 15, "APAP",cex = 1.2)
points(APAP$Time, log(APAP$APAP * 1000))
pksim(y, vars = "lnCPL_AG_mcgL", legend = F)
text(1, 15, "AG",cex = 1.2)
points(APAP$Time, log(APAP$AG * 1000))
pksim(y, vars = "lnCPL_AS_mcgL", legend = F)
text(1, 15, "AS",cex = 1.2)
points(APAP$Time, log(APAP$AS * 1000))
mtext("Time", SOUTH<-1, line=0.4, cex=1.2, outer=TRUE)
mtext("Conc.", WEST<-2, line=0.4, cex=1.2, outer=TRUE)
```

In addition, through using the `check()`, the parameter with sensitivity and convergence indices over the given condition can be easily detected for all output variables.

```{r, eval=T}
check(x)
```

The `check()` also provides some feasible argument to specify the target output or change the cut-off value.

```{r, eval=T}
check(x, vars = "lnCPL_APAP_mcgL", SI.cutoff = 0.1, CI.cutoff = 0.1)
```

### Heatmap visualization combined with an index “cut-off”

Based on our previous study, we proposed the heatmap visualization approach to distinguish "influential" and "non-influential" parameters with a cut-off. Through the given argument `order`, we can select the specific order of sensitivity measurement that we're interested in (Figure 3 & 4).

```{r, eval=T}
heat_check(x, order = "interaction")
```

```{r, eval=T}
heat_check(x, order = "total order")
```

Also, adding the `index = "CI"` in the function can further investigate the convergence of the sensitivity index. Based on the current setting of sampling number, most parameters cannot reach the acceptable criteria of convergence. Therefore, the higher number of sampling is necessary.

```{r, eval=T, fig.height=9}
heat_check(x, index = "CI", CI.cutoff = 0.05)
```

## Code formatting

Don't use markdown, instead use the more precise latex commands:

* \proglang{Java}
* \pkg{plyr}
* \code{print("abc")}

# Concluding remarks

# Acknowledgments

This work was funded, in part, by grant 1U01FD005838 from the U.S. FDA.  This abstract reflects the views of the author and should not be construed to represent FDA’s views or policies. We thank Dr. Yasha Hartberg and Dr. Barbara Gastel at the Texas A&M University for reviewing the manuscript and consultation.

# References

### 2019-01-31 manuscript/manuscript.tex
add content

@@
 \author{
-FirstName LastName\\University/Company \And Second Author\\Affiliation
+Nan-Hung Hsieh\\Texas A\&M University \And Brad Reisfeld\\Colorado State University \And Weihsueh A. Chiu\\Texas A\&M University
 }
-\title{A Capitalized Title: Something about a Package \pkg{foo}}
+\title{\pkg{pksensi}: An R package to apply sensitivity analysis in
+pharmacokinetic models}
 
-\Plainauthor{FirstName LastName, Second Author}
-\Plaintitle{A Capitalized Title: Something about a Package foo}
-\Shorttitle{\pkg{foo}: A Capitalized Title}
+\Plainauthor{Nan-Hung Hsieh, Brad Reisfeld, Weihsueh A. Chiu}
+\Plaintitle{pksensi: An R package an R package to apply sensitivity analysis in
+pharmacokinetic models}
+\Shorttitle{\textbf{pksensi}: Sensitivity analysis for pharmacokinetic models in R}
 
 \Abstract{
-The abstract of the article.
+Sensitivity analysis is an essential tool for modelers to understand the
+influence of model parameters on model outputs. It is also increasingly
+used in developing and assessing pharmacokinetic models. In our previous
+work, we applied a global sensitivity analysis workflow to reduce the
+computational burden in the Bayesian Markov Chain Monte Carlo-based
+calibration process of a physiologically based pharmacokinetic (PBPK)
+model. Although several sensitivity analysis algorithms are available,
+no comprehensive package exists that allows users to seamlessly solve
+the PBPK model differential equations, run sensitivity analyses,
+visualize sensitivity analysis results, and discriminate between the
+``non-influential'' model parameters that can be fixed and those that
+need calibration. Therefore, we developed an open-source R package,
+called \textbf{pksensi}, to fill this need and make sensitivity analysis
+more accessible in pharmacological and toxicological research. This
+package can investigate both parameter uncertainty and sensitivity in
+pharmacokinetic models, such as PBPK model with multivariate model
+outputs (multiple time-points and tissue compartments). It also refined
+the extended Fourier Amplitude Sensitivity Test method for sensitivity
+analysis by adding a random phase-shift to analyze the statistical
+variability of the sensitivity index for each model parameter.
+Furthermore, pksensi includes functions to check the convergence and
+sensitivity of model parameters, providing a means to assess the
+robustness of the sensitivity measurement. Utilizing pksensi, we
+successfully reproduced our previously published sensitivity analysis
+results of human PBPK modeling for acetaminophen and its two primary
+metabolites. Overall, pksensi improves the user experience of performing
+sensitivity analysis and can create robust and reproducible results for
+decision making in pharmacokinetic model calibration.
 }
 
-\Keywords{keywords, not capitalized, \proglang{Java}}
-\Plainkeywords{keywords, not capitalized, Java}
+\Keywords{Pharmacokinetics, Physiologically based pharmacokinetics, Sensitivity analysis}
+\Plainkeywords{Pharmacokinetics, Physiologically based pharmacokinetics, Sensitivity analysis}
 
@@
 \Address{
-    FirstName LastName\\
-  University/Company\\
-  First line Second line\\
-  E-mail: \email{name@company.com}\\
-  URL: \url{http://rstudio.com}\\~\\
+        Weihsueh A. Chiu\\
+  Texas A\&M University\\
+  4458 TAMU, College Station, TX 77843, USA\\
+  E-mail: \email{wchiu@cvm.tamu.edu}\\
+
   }
@@
 
-This template demonstrates some of the basic latex you'll need to know
-to create a JSS article.
+Sensitivity analysis is a mathematical technique to investigate how
+variations in model parameters affect model outputs. An increasing
+number of studies use sensitivity analysis to determine which model
+parameters contribute to high variation in model predictions (Ferretti
+et al., 2016). Also, this technique has been applied in pharmacology and
+toxicology research (Loizou et al., 2015; McNally et al., 2012;
+Scherholz et al., 2018). Pharmacokinetic modeling describes the changes
+in the concentrations or amounts of a substance within model
+compartments over time. The goal of sensitivity analysis in
+pharmacokinetic research is to examine the sensitivity of output
+variables (e.g.~compound concentration in blood or tissues) in
+pharmacokinetic models responds to input parameters, such as anatomical,
+physiological, and kinetic constants (McNally et al., 2011). It can be
+further applied to parameter prioritization and parameter fixing before
+model calibration (Hsieh et al., 2018).
+
+In our previous work (Hsieh et al., 2018), we developed an approach to
+apply global sensitivity analysis workflow to reduce the computational
+burden in the Bayesian, Markov Chain Monte Carlo (MCMC)-based
+calibration process of a physiologically based pharmacokinetic (PBPK)
+model. We used GNU MCSim (Bois, 2009), an effective simulation package
+for Bayesian population PBPK modeling, to calibrate the model. We found
+that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of
+global sensitivity analysis algorithm, had the best balance of
+efficiency and accuracy for a complex, multi-compartment, multi-dataset,
+and multi-metabolite PBPK model. Also, we found some efficient
+visualization approaches that can be used to distinguish between
+``influential'' and ``non-influential'' parameters. We also found a
+useful approach for communicating the parameter sensitivity in decision
+making.
+
+Most of sensitivity analysis tools are constructed using scientific
+computing software, such as GNU Octave/Matlab (Pianosi et al., 2015),
+python (Herman and Usher, 2017), and R (Ghanem et al., 2017).
+Previously, we mainly used R with the sensitivity package (Pujol et al.,
+2017), which is a practical tool to conduct local and global sensitivity
+analysis. The sensitivity package includes some functions to generate
+the parameter sequences by using different computing algorithms. Also,
+it provides a feasible way to integrate external modeling results. This
+computational approach can effectively solve the heavy computing burden
+of numerical solutions within the pure R environment. Several R
+packages, such as FME, multisensi, and ODEsensitivity include
+sensitivity analysis tools and are freely available in the Comprehensive
+R Archive Network. The FME package contains a basic method for
+performing local sensitivity analysis for dynamic models (Soetaert and
+Petzoldt, 2016). The multisensi package provides sensitivity analysis
+for models with multivariate output (Bidot et al., 2018). ODEsensitivity
+can perform sensitivity analysis in ordinary differential equation (ODE)
+models (Weber et al., 2018). It uses the ODE interface from the deSolve
+package, connecting it with the sensitivity analysis from the
+sensitivity package, and can perform sensitivity analysis on models with
+multivariate output. Although these tools provide various approaches to
+perform sensitivity analysis, we have not find a suitable package that
+provides functions with which users can visualize and distinguish the
+``non-influential'' model parameters and can further apply to parameter
+fixing in pharmacokinetic modeling.
+
+Here we develop an R package, called pksensi, which is designed to make
+sensitivity analysis more accessible and reproducible in pharmacological
+and toxicological researches. This package can investigate both
+parameter uncertainty and sensitivity in pharmacokinetic models, such as
+PBPK, and advanced compartment absorption and transit models with
+multivariate model output. The design concepts of pksensi are:
+
+\begin{itemize}
+\tightlist
+\item
+  Cross-platform: Models can run on Windows/MacOS/Linux
+\item
+  Freely available: All related packages are free and open source
+\item
+  Integrated application: Users can run pharmacokinetic models in R with
+  script that were written in C or GNU MCSim
+\item
+  Decision making: The output results and visualization tools can be
+  used to easily determine which parameters have ``non-influential''
+  effects on the model output and can be fixed in model calibration.
+\end{itemize}
+
+\hypertarget{workflow}{%
+\section{Workflow}\label{workflow}}
+
+\hypertarget{installation}{%
+\subsection{Installation}\label{installation}}
+
+\pkg{pksensi} is available from the comprehensive R Archive Network at
+\url{https://CRAN.R-project.org/package=pksensi}. The latest up-to-date
+version is available through GitHub at
+\url{https://github.com/nanhung/pksensi}.
+
+\hypertarget{parameter-matrix-generation}{%
+\subsection{Parameter matrix
+generation}\label{parameter-matrix-generation}}
+
+We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely
+used global sensitivity analysis approach in biomathematical modeling,
+especially for ordinary differential equation-based dynamical models, in
+this package (Saltelli et al., 1999). Most of the available eFAST
+functions, such as the algorithm to generate the parameter space and to
+set the sampling frequency, were sourced from the R \pkg{sensitivity}
+package (Pujol et al., 2017). To test the convergence and robustness of
+the sensitivity measurement, we included a random phase-shift approach
+to replicate sampling from random starting points across parameter
+space. This can be written as:
+
+\[ x_i = \frac{1}{2} + \frac{1}{\pi}\arcsin(\sin(\omega_is + \varphi_i)) \]
+
+where \(x_i\) is the nominal value of the i-th parameter, and
+\(\omega_i\) is a vector giving the set of frequencies, one frequency
+for each parameter. The default set of frequencies is based on the study
+from Saltelli et al.~(1999). The first frequency of the vector
+\(\omega\) is assigned to each factor \(x_i\) in turn (corresponding to
+the estimation of Sobol indices of first and total order, respectively).
+Through the random phase-shift, the robust result of the sensitivity
+measurement should be similar across replications under the same sample
+size.
+
+To investigate the convergence of sensitivity indices, we adopted the
+exam approach proposed by Sarrazin et al.~(2016). This method
+quantitatively assesses convergence by computing the width of 95\%
+confidence intervals for all parameters across all time-points and
+output variables. To integrate the random phase-shift with eFAST, we
+built the \proglang{rfast99} function, which can be used to perform the
+sensitivity test for the whole pharmacokinetic modeling process. The
+function \proglang{rfast99} in pksensi can sample and generate the
+testing parameter matrix based on a given argument of sample size n and
+probability distribution (e.g., mean/s.d. of normal distribution,
+meanlog/sdlog of log-normal distribution, and minimum/maximum of uniform
+distribution). The number of model evaluations is equal to the sample
+size times the number of model parameters.
+
+\hypertarget{pharmacokinetic-modeling}{%
+\subsection{Pharmacokinetic modeling}\label{pharmacokinetic-modeling}}
+
+The \pkg{pksensi} provides a useful function to solve ordinary
+differential equations in pharmacokinetic models using analytical and
+numerical approaches. The solution can be performed under the pure R
+coding environment by linking pksensi with the deSolve package (Soetaert
+et al., 2018). However, the most efficient way to achieve high
+computational speed is through compiled, lower-level languages, such as
+FORTRAN, C, or C++. The \pkg{pksensi} includes a function to compile and
+create dynamic-link libraries (.dll) on Windows and shared objects (.so)
+on Unix-liked systems (e.g., Linux and MacOS). This compiled file can
+load and execute through the build-in function in R. However, Windows
+users need to install Rtools or MinGW to compile the source code by
+using the GNU GCC compiler. More requirements are detailed in the study
+from Soetaert et al.~(2010). The C implementation of the example model
+can be found as an example within MCSim v6.0.1 (Bois, 2009). The
+\pkg{pksensi} can also link with MCSim to compile the model code, used
+in solving each system of equations.
+
+\hypertarget{output-visualization-and-decision-making}{%
+\subsection{Output visualization and decision
+making}\label{output-visualization-and-decision-making}}
+
+The output of sensitivity analysis returns a list that contains the
+given time-points and values of all state variables (e.g., chemical
+concentrations in blood). This format is particularly suited for further
+graphical routines within pkseni. Also, a plot can can create
+time-dependent sensitivity measurements of first and total order effect
+for all parameters in one figure. Furthermore, pksensi includes
+functions to check the convergence and sensitivity of model parameters,
+providing a means to assess the robustness of the sensitivity
+measurement. We also developed a ``cut-off''-based approach in this
+package to accurately distinguish between ``influential'' and
+``non-influential'' parameters. Finally, pksensi includes numerous
+visualization tools for the effective investigation and communication of
+results in decision making.
+
+\hypertarget{examples}{%
+\section{Examples}\label{examples}}
+
+\hypertarget{one-compartment-taxicokinetic-model}{%
+\subsection{One-compartment taxicokinetic
+model}\label{one-compartment-taxicokinetic-model}}
+
+\hypertarget{equations}{%
+\subsubsection{Equations}\label{equations}}
+
+In this example, We use a simple, one-compartment PK model from
+\texttt{httk} package \citep{JSSv079i04} to demonstrate how
+\texttt{pksensi} can be applied to pharmacokinetic studies. The
+differential equations for the one-compartment pharmacokinetic model can
+be written as:
+
+\[\frac{dA_{gutlumen}}{dt} = -k_{gutabs} \cdot A_{gutlumen} + g(t)\]
+\[\frac{dC_{rest}}{dt} = \frac{k_{gutabs}}{V_{dist}}-k_{elim} \cdot C_{rest}\]
+
+where \(A_{gutlumen}\) is the state variable that describes the quantity
+of compound in gut lumen (mol), \(k_{gutabs}\) is the absorption rate
+constant that describes the chemical absorption from the gut lumen into
+gut tissue through first-order processes (/h), \(V_{dist}\) is the
+volume of distribution (L), and \(k_{elim}\) is the elimination rate
+constant (/h), which is equal to the total clearance divided by the
+volume of distribution. The time-dependent function \(g(t)\) is used to
+describe the oral dosing schedule. \(C_{rest}\) is the chemical
+concentration in plasma that can be used to compare with observed
+results in a pharmacokinetic experiment (mol/L).
+
+\hypertarget{model-implementations-with-r-desolve-package}{%
+\subsubsection{Model implementations with R deSolve
+package}\label{model-implementations-with-r-desolve-package}}
+
+In the beginning, we need to pre-install GNU MCSim \citep{JSSv002i09}
+and related compiler to let us generate .c file and executable file. The
+GNU MCSim can be installed by following the instruction in GNU MCSim's
+manual on \url{https://www.gnu.org/software/mcsim/mcsim.html} or using
+the build-in function \texttt{mcsim\_install()} in \texttt{pksensi}. The
+GNU compiler is necessary for users that use Linux or MacOS. For Windows
+users, you should install Rtools on
+\url{https://cran.r-project.org/bin/windows/Rtools/} and use
+\texttt{Sys.setenv()} to set the working path of compiler. The
+\texttt{Sys.which("gcc")} and
+\texttt{system(\textquotesingle{}g++\ -v\textquotesingle{})} can check
+whether we can run compiler correctly.
+
+We first implemented this model in R by compiling the file written in C.
+pksensi allows users to select the preferred method to solve the
+pharmacokinetic model, either with the \texttt{deSolve} package or with
+GNU MCSim through the compile function. However, running model under GNU
+MCSim native code can have faster speed to obtain the model outputs.
+
+The following R script can download and compile the model description
+file (\texttt{pbtk1cpt.model}) and use \texttt{deSolve} package
+\citep{JSSv033i09} to solve ordinary differential equations in our
+model. The example model code of one-compartment PBTK model is available
+with pksensi package:
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> pbtk1cpt_model()
+R> cat(readLines("pbtk1cpt.model"), sep = "\n")
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> # ------------------------------------------
+#> # pbtk1cpt.model (Based on R httk package)
+#> # ------------------------------------------
+#> 
+#> #
+#> States  = { Agutlument, Acompartment, Ametabolized, AUC};
+#> 
+#> #
+#> Outputs = {Ccompartment};
+#> 
+#> # Parameters
+#> vdist = 0;
+#> ke = 0;
+#> kgutabs = 1;
+#> 
+#> #
+#> Dynamics {
+#>   Ccompartment = Acompartment / vdist;
+#>   dt (Agutlument)  = - kgutabs * Agutlument;
+#>   dt (Acompartment)  = kgutabs * Agutlument - ke * Acompartment;
+#>   dt (Ametabolized) = ke * Acompartment;
+#>   dt (AUC) = Ccompartment;
+#> }
+#> 
+#> End.
+\end{CodeOutput}
+\end{CodeChunk}
+
+Then, use \texttt{compile\_model()} to generate the executable files
+(\texttt{pbtk1cpt.dll} on Windows or \texttt{pbtk1cpt.so} on other
+systems) and R file (\texttt{pbtk1cpt\_inits.R}) with default input
+parameters and initial state settings with the definition of
+\texttt{application\ =\ "R"}.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> mName <- "pbtk1cpt"
+R> compile_model(mName, application = "R")
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> * Created file 'pbtk1cpt.so'.
+\end{CodeOutput}
+
+\begin{CodeInput}
+R> source(paste0(mName, "_inits.R"))
+\end{CodeInput}
+\end{CodeChunk}
+
+The parameter values and initial states can be customized to specify the
+properties and schedule for the given dosing scenario.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> parms <- initParms()
+R> parms["vdist"] <- 0.74
+R> parms["ke"] <- 0.28
+R> parms["kgutabs"] <- 2.18
+R> initState <- initStates(parms=parms)
+R> initState["Agutlument"] <- 10
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> parms
+\end{CodeInput}
+
+\begin{CodeOutput}
+#>   vdist      ke kgutabs 
+#>    0.74    0.28    2.18
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> initState
+\end{CodeInput}
+
+\begin{CodeOutput}
+#>   Agutlument Acompartment Ametabolized          AUC 
+#>           10            0            0            0
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> Outputs
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> [1] "Ccompartment"
+\end{CodeOutput}
+\end{CodeChunk}
+
+In the current setting, we assumed the initial condition of the intake
+chemical to be 10 mol. The \texttt{initParms} and \texttt{initStates}
+functions were used to customize the parameter values and the initial
+state that will be used in the \texttt{solve\_fun} function. These
+parameter value can be adopted from the \texttt{httk} package, which
+includes physico-chemical and drug biological properties for 553
+chemicals. In this case, we used the parameter value of theophylline in
+this example. The given \texttt{vdist}, \texttt{ke}, and
+\texttt{kgutabs} are 0.74, 0.28, and 2.18, respectively.
+
+Through \texttt{ode} function in \texttt{deSolve} package, we can
+visualize the pharmacokinetic according to the given parameter
+conditions such as time points (\texttt{times}):
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> times <- seq(from = 0.01, to = 24.01, by = 1)
+R> y <- deSolve::ode(initState, times, func = "derivs", parms = parms, 
+R+                   dllname = mName, initfunc = "initmod", nout = 1, outnames = Outputs)
+\end{CodeInput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> plot(y)
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-8-1}
+
+}
+
+\caption[Figure 1]{Figure 1. Simulation results of one-compartment PBTK model.}\label{fig:unnamed-chunk-8}
+\end{figure}
+\end{CodeChunk}
+
+To conduct sensitivity analysis for the parameters in one-compartment
+pharmacokinetic model in this case, we want to quantify the impact of
+these three parameters on the chemical concentration in plasma during
+24-hour time period post intake. We assume a uniform distribution for
+the estimate for each parameter with the coefficient of uncertainty
+within 50\%. The parameter ranges are assumed to be (0.37, 1.12) for
+\texttt{vdist}, (0.0058, 0.0174) for \texttt{ke}, and (0.045, 0.136) for
+\texttt{kgutabs}. The sample number determines the robustness of the
+result of sensitivity analysis. Higher sample numbers can generate
+narrow confidence intervals for sensitivity measurements across
+different replications. However, they might cause heavy computational
+burden for complex models. Here we use a sample number of 400 with 20
+replications:
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> LL <- 0.5 
+R> UL <- 1.5
+R> q <- "qunif"
+R> q.arg <- list(list(min = parms["vdist"] * LL, max = parms["vdist"] * UL),
+R+              list(min = parms["ke"] * LL, max = parms["ke"] * UL), 
+R+              list(min = parms["kgutabs"] * LL, max = parms["kgutabs"] * UL)) 
+R> set.seed(1234)
+R> params <- c("vdist", "ke", "kgutabs")
+R> x <- rfast99(params, n = 200, q = q, q.arg = q.arg, replicate = 20)
+\end{CodeInput}
+\end{CodeChunk}
+
+Because the pharmacokinetic model is being used to describe a continuous
+process for the chemical concentration over time, the sensitivity
+measurements can also show the time-dependent relationships for each
+model parameter. Here we define the output time points to examine the
+change of the parameter sensitivity over time. To solve the
+pharmacokinetic model through deSolve, we need to provide the details of
+the argument:
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> y <- solve_fun(x, times, initState = initState,
+R+                outnames = Outputs, dllname = mName)
+R> tell2(x,y)
+\end{CodeInput}
+\end{CodeChunk}
+
+To create the time-dependent sensitivity measurement, we set the time
+duration from 0.01 to 24.01 hours in the example. The
+\texttt{initParmsfun} is used to generate the sampling value for each
+parameter. The \texttt{outnames}, \texttt{dllname}, \texttt{func},
+\texttt{initfunc} are based on the arguments from the ode function in
+\texttt{deSolve} package. The details of the model structure and these
+arguments are defined in \texttt{pbtk1comp.c}. and
+\texttt{pbtk1comp\_inits.R}. Finally, the \texttt{tell2} function is
+used to integrate the parameter values and the output results of
+numerical analysis that were generated and stored in variables x and y.
+The result of object x is an object of rfast99, which has specific
+\texttt{print}, \texttt{plot}, and \texttt{check} method. The print
+function gives the sensitivity and convergence indices for main,
+interaction, and total order at each time point. In addition to print
+out the result of sensitivity analysis, the more efficient way to
+distinguish the influence of model parameter is to visualize them. The
+time-dependent sensitivity indices are shown in Figure 2.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> plot(x)
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-11-1}
+
+}
+
+\caption[Figure 2]{Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24 hour time period intake.}\label{fig:unnamed-chunk-11}
+\end{figure}
+\end{CodeChunk}
+
+Here, we can find that \texttt{vdist} and \texttt{ke} are dominating the
+plasma concentration in the before and after 5-hour post chemical
+intake, respectively, representing that the elimination is a key
+parameter to dominate the plasma concentration. Besides, the
+\texttt{kgutabs} only plays a crucial role to determine the plasma
+concentration in the first hour. The relationship between concentration
+and the parameters can be plotted as follow (Figure 3):
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> par(mfrow = c(3,3), mar = c(2,2,2,2), oma = c(2,2,1,1))
+R> plot(x$a[,1,"vdist"], y[,1,"0.01",], main = "vdist")
+R> text(1, .7, "t=0.01",cex = 1.2)
+R> plot(x$a[,1,"ke"], y[,1,"0.01",], main = "ke")
+R> plot(x$a[,1,"kgutabs"], y[,1,"0.01",], main = "kgutabs")
+R> plot(x$a[,1,"vdist"], y[,1,"2.01",])
+R> text(1, 18, "t=2.01",cex = 1.2)
+R> plot(x$a[,1,"ke"], y[,1,"2.01",])
+R> plot(x$a[,1,"kgutabs"], y[,1,"2.01",])
+R> plot(x$a[,1,"vdist"], y[,1,"24.01",])
+R> text(1, .7, "t=24.01",cex = 1.2)
+R> plot(x$a[,1,"ke"], y[,1,"24.01",])
+R> plot(x$a[,1,"kgutabs"], y[,1,"24.01",])
+R> mtext("parameter", SOUTH<-1, line=0.4, outer=TRUE)
+R> mtext("Ccompartment", WEST<-2, line=0.4, outer=TRUE)
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-12-1}
+
+}
+
+\caption[Figure 3]{Figure 3. The relationship between model parameter and estimated concentration under the time-point of 0.01, 2.01, and 24.01 hr}\label{fig:unnamed-chunk-12}
+\end{figure}
+\end{CodeChunk}
+
+The x is a list of class ``rfast99'', containing all the input arguments
+detailed before and the calculated sensitivity indices of first order
+(\texttt{mSI}), interaction (\texttt{iSI}), and total order
+(\texttt{tSI}). The convergence indices are also stored in the list
+named \texttt{mCI}, \texttt{iCI}, and \texttt{tCI}. The parameter values
+are stored in an array \texttt{x\$a} with c(model evaluation,
+replication, parameters).
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> dim(x$a)
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> [1] 600  20   3
+\end{CodeOutput}
+\end{CodeChunk}
+
+In addition, the output are also formated with c(model evaluation,
+replication, time, variable).
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> dim(y)
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> [1] 600  20  25   1
+\end{CodeOutput}
+\end{CodeChunk}
+
+The \texttt{check()} is a useful function to determine which parameters
+have relative lower sensitivity measurement across the given time
+interval, and therefore can be applied parameter fixing in model
+calibration. The argument of \texttt{SI.cutoff} is setting at 0.5 to
+detect the relative non-influential parameters in this case.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> check(x, SI.cutoff = 0.5)
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> 
+#> Sensitivity check ( Index > 0.5 )
+#> ----------------------------------
+#> First order:
+#>  vdist ke 
+#> 
+#> Interaction:
+#>   
+#> 
+#> Total order:
+#>  vdist ke 
+#> 
+#> Unselected factors in total order:
+#>  kgutabs 
+#> 
+#> 
+#> Convergence check ( Index > 0.05 )
+#> ----------------------------------
+#> First order:
+#>   
+#> 
+#> Interaction:
+#>   
+#> 
+#> Total order:
+#> 
+\end{CodeOutput}
+\end{CodeChunk}
+
+Based on the sensitivity measurement of the total order, the result
+shows that \texttt{kgutabs} has relative lower measurement of
+sensitivity index.
+
+\hypertarget{model-implementations-with-gnu-mcsim}{%
+\subsubsection{Model implementations with GNU
+MCSim}\label{model-implementations-with-gnu-mcsim}}
+
+Alternatively, to solve ODE by using GNU MCSim, we need to change the
+argument to \texttt{application\ =\ mcsim} in \texttt{compile\_model()}.
+Rather than apply R \texttt{deSolve} to solve differential equations,
+the GNU MCSim can provide higher computational speed in global
+sensitivity analysis.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> system.time(y<-solve_fun(x, times, initState = initState, 
+R+                          outnames = Outputs, dllname = mName))
+\end{CodeInput}
+
+\begin{CodeOutput}
+#>    user  system elapsed 
+#>   5.656   0.019   5.695
+\end{CodeOutput}
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> compile_model(mName, application = "mcsim")
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> * Created executable file 'mcsim.pbtk1cpt'.
+\end{CodeOutput}
+\end{CodeChunk}
+
+Sililiar to \texttt{solve\_fun} that can define the initial parameter
+and state values through input function, the \texttt{solve\_mcsim} has a
+\texttt{condition} argument that is used to givien the specific input
+value such as oral dose or fixing parameter value or initial state
+variable.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> conditions <- c("Agutlument = 10") # Set the initial state of Agutlument = 10 
+R> system.time(y<-solve_mcsim(x, mName = mName, 
+R+                            params = params,
+R+                            vars = Outputs,
+R+                            time = times,
+R+                            condition = conditions))
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> * Created input file "input.in".
+\end{CodeOutput}
+
+\begin{CodeOutput}
+#>    user  system elapsed 
+#>   1.604   0.084   1.523
+\end{CodeOutput}
+\end{CodeChunk}
+
+Under the same given condition, it takes 5-6 (\texttt{deSolve}) and 1-2
+(GNU MCSim) seconds to solve model. The \texttt{solve\_mcsim()} shows
+the better computational performance than \texttt{solve\_fun()} in
+\texttt{pksensi}.
+
+\hypertarget{acetaminophen-pbpk-model}{%
+\subsection{Acetaminophen-PBPK model}\label{acetaminophen-pbpk-model}}
+
+\hypertarget{uncertainty-and-sensitivity-analysis}{%
+\subsubsection{Uncertainty and sensitivity
+analysis}\label{uncertainty-and-sensitivity-analysis}}
+
+The aim of this section is to reproduce our previous published
+\citep{fphar201800588} result of global sensitivity analysis for
+acetaminophen PBPK model through \texttt{pksensi}. The model codes are
+included in this package and can be generated through
+\texttt{pbpk\_apap\_model()}. We applied the global sensitivity analysis
+workflow to the original published model with 21 model parameters
+\citep{s13318-015-0253-x}. The descriptions of each parameter and the
+sampling ranges are list in Table 1.
+
+\textless{}Table 1\textgreater{}
+
+Same as the example of one-compartment PBTK model. The model parameter
+and the corresponding sampling range should be defined to create the
+parameter matrix. Previously, the probability distributions of model
+parameters were set to either truncated normal or uniform distribution
+when the parameters have informative prior information or not. To
+rapidly reach the acceptance convergence, we apply uniform distribution
+for all testing parameters. The ranges of informative parameters are set
+to 1.96-times difference for single side (approximate 54.6 times
+difference between minimum and maximum) under log-scaled. The nominal
+values of informative model parameters were defined as:
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> # Nominal value
+R> Tg <- log(0.23)
+R> Tp <- log(0.033)
+R> CYP_Km <- log(130)
+R> SULT_Km_apap <- log(300)
+R> SULT_Ki <- log(526)
+R> SULT_Km_paps <- log(0.5)
+R> UGT_Km <- log(6.0e3)
+R> UGT_Ki <- log(5.8e4)
+R> UGT_Km_GA <-log(0.5)
+R> Km_AG <- log(1.99e4)
+R> Km_AS <- log(2.29e4)
+R> 
+R> rng <- 1.96 
+\end{CodeInput}
+\end{CodeChunk}
+
+Generally, The wide range of parameter value might cause the
+computational error in the solver. One of the effective ways to prevent
+this problem is to adjust the value of relative and absolute error
+tolerance to control the error appearance by resetting these parameters
+in a lower value. The \texttt{generate\_infile()} provide the arguments
+of \texttt{rtol} and \texttt{atol} that can be adjusted to prevent the
+unwanted error. However, the modification will slow down the
+computational speed. Therefore, the alternative method to prevent the
+computational error is to detect the crucial parameter range that causes
+the problem. Also, setting the maximum number of (internally defined)
+steps to higher value instead of using the default value (500) can
+prevent this problem. The maximum number of step is set to 5000 in this
+case.
+
+In this test case, we adjusted the range of \texttt{SULT\_VmaxC} and
+\texttt{UGT\_VmaxC} from U(0, 15) to U(0, 10). The relative and absolute
+error tolerance were set to 1e-7 and 1e-9, respectively, to prevent the
+computational error in MCSim,
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> params <- c("lnTg", "lnTp", "lnCYP_Km","lnCYP_VmaxC",
+R+            "lnSULT_Km_apap","lnSULT_Ki","lnSULT_Km_paps","lnSULT_VmaxC",
+R+            "lnUGT_Km","lnUGT_Ki","lnUGT_Km_GA","lnUGT_VmaxC",
+R+            "lnKm_AG","lnVmax_AG","lnKm_AS","lnVmax_AS",
+R+            "lnkGA_syn","lnkPAPS_syn", "lnCLC_APAP","lnCLC_AG","lnCLC_AS")
+R> q <- "qunif"
+R> q.arg <-list(list(Tg-rng, Tg+rng),
+R+              list(Tp-rng, Tp+rng),
+R+              list(CYP_Km-rng, CYP_Km+rng),
+R+              list(-2., 5.),
+R+              list(SULT_Km_apap-rng, SULT_Km_apap+rng),
+R+              list(SULT_Ki-rng, SULT_Ki+rng),
+R+              list(SULT_Km_paps-rng, SULT_Km_paps+rng),
+R+              list(0, 10),
+R+              list(UGT_Km-rng, UGT_Km+rng),
+R+              list(UGT_Ki-rng, UGT_Ki+rng),
+R+              list(UGT_Km_GA-rng, UGT_Km_GA+rng),
+R+              list(0, 10),
+R+              list(Km_AG-rng, Km_AG+rng),
+R+              list(7., 15),
+R+              list(Km_AS-rng, Km_AS+rng),
+R+              list(7., 15),
+R+              list(0., 13),
+R+              list(0., 13),
+R+              list(-6., 1),
+R+              list(-6., 1),
+R+              list(-6., 1))
+R> 
+R> times <- seq(from = 0.1, to = 12.1, by = 0.2)
+R> set.seed(1234)
+R> x <- rfast99(params = params, n = 512, q = q, q.arg = q.arg, replicate = 10) 
+\end{CodeInput}
+\end{CodeChunk}
+
+After creating the \texttt{pbpk\_apap.model} in the working directory,
+the next step is to generate the executable files
+(\texttt{mcsim.pbpk\_apap}) through \texttt{compile\_model()}.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> mName <- "pbpk_apap"
+R> pbpk_apap_model()
+R> compile_model(mName, application = "mcsim")
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> * Created executable file 'mcsim.pbpk_apap'.
+\end{CodeOutput}
+\end{CodeChunk}
+
+To improve the computational speed, this case only uses MCSim to
+estimate the concentration of acetaminophen (APAP) and its metabolites
+glucuronide (AG) and sulfate (AS) in plasma. The setting oral dose of
+APAP is 20 mg/kg in this example. Generally, the input dosing method can
+be defined through the \texttt{condition} argument. Since the unit of
+the given dose is mg/kg, the \texttt{mgkg\_flag} is set to 1 to declare
+the statement. More definition of input can be found in the section of
+input functions in GNU MCSim User's Manual
+(\url{https://www.gnu.org/software/mcsim/mcsim.html\#Input-functions}).
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> vars <- c("lnCPL_APAP_mcgL", "lnCPL_AG_mcgL", "lnCPL_AS_mcgL")
+R> conditions <- c("mgkg_flag = 1",
+R+                 "OralExp_APAP = NDoses(2, 1, 0, 0, 0.001)",
+R+                 "OralDose_APAP_mgkg = 20.0")
+R> generate_infile(params = params,
+R+                 vars = vars,
+R+                 time = times, 
+R+                 condition = conditions,
+R+                 rtol = 1e-7, atol = 1e-9)
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> * Created input file "input.in".
+\end{CodeOutput}
+
+\begin{CodeInput}
+R> system.time(y <- solve_mcsim(x, mName = mName, 
+R+                              params = params,
+R+                              vars = vars,
+R+                              time = times,
+R+                              condition = conditions,
+R+                              generate.infile = F))
+\end{CodeInput}
+
+\begin{CodeOutput}
+#>    user  system elapsed 
+#> 150.905   1.794 152.061
+\end{CodeOutput}
+
+\begin{CodeInput}
+R> tell2(x,y)
+\end{CodeInput}
+\end{CodeChunk}
+
+The plotting function can output the result of time-dependent
+sensitivity measurement to determine the parameter impact on model
+output over time (Figure 1).
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> plot(x, vars = "lnCPL_AG_mcgL")
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-23-1}
+
+}
+
+\caption[Figure 4]{Figure 4. }\label{fig:unnamed-chunk-23}
+\end{figure}
+\end{CodeChunk}
+
+The uncertainty analysis is a crucial step before model calibration. We
+can apply uncertainty analysis through the \texttt{pksim()} by the given
+name of output variables (Figure 2). Through this visualization
+approach, we can recognize whether the simulated outputs can accurately
+simulate the same concentration profile as the in-vivo experiment under
+the setting of parameter ranges. In Figure 2, all experiment points are
+included in the intervals, representing the acceptable set of the
+parameter range.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> par(mfrow = c(2,2), mar = c(2,2,1,1), oma = c(2,2,1,1))
+R> pksim(y, vars = "lnCPL_APAP_mcgL")
+R> text(1, 15, "APAP",cex = 1.2)
+R> points(APAP$Time, log(APAP$APAP * 1000))
+R> pksim(y, vars = "lnCPL_AG_mcgL", legend = F)
+R> text(1, 15, "AG",cex = 1.2)
+R> points(APAP$Time, log(APAP$AG * 1000))
+R> pksim(y, vars = "lnCPL_AS_mcgL", legend = F)
+R> text(1, 15, "AS",cex = 1.2)
+R> points(APAP$Time, log(APAP$AS * 1000))
+R> mtext("Time", SOUTH<-1, line=0.4, cex=1.2, outer=TRUE)
+R> mtext("Conc.", WEST<-2, line=0.4, cex=1.2, outer=TRUE)
+\end{CodeInput}
+\begin{figure}
+
+{\centering \includegraphics{manuscript_files/figure-latex/unnamed-chunk-24-1}
+
+}
+
+\caption[Figure 5]{Figure 5. }\label{fig:unnamed-chunk-24}
+\end{figure}
+\end{CodeChunk}
+
+In addition, through using the \texttt{check()}, the parameter with
+sensitivity and convergence indices over the given condition can be
+easily detected for all output variables.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> check(x)
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> 
+#> Sensitivity check ( Index > 0.05 )
+#> ----------------------------------
+#> First order:
+#>  lnTg lnTp lnSULT_VmaxC lnUGT_Km lnUGT_VmaxC lnKm_AG lnVmax_AG lnKm_AS lnVmax_AS lnkGA_syn lnkPAPS_syn lnCLC_APAP lnCLC_AG lnCLC_AS 
+#> 
+#> Interaction:
+#>  lnTg lnSULT_Km_apap lnSULT_VmaxC lnUGT_VmaxC lnVmax_AG lnVmax_AS lnkGA_syn lnkPAPS_syn lnCLC_APAP lnCLC_AG lnCLC_AS 
+#> 
+#> Total order:
+#>  lnTg lnTp lnSULT_Km_apap lnSULT_VmaxC lnUGT_Km lnUGT_VmaxC lnKm_AG lnVmax_AG lnKm_AS lnVmax_AS lnkGA_syn lnkPAPS_syn lnCLC_APAP lnCLC_AG lnCLC_AS 
+#> 
+#> Unselected factors in total order:
+#>  lnCYP_Km lnCYP_VmaxC lnSULT_Ki lnSULT_Km_paps lnUGT_Ki lnUGT_Km_GA 
+#> 
+#> 
+#> Convergence check ( Index > 0.05 )
+#> ----------------------------------
+#> First order:
+#>  lnTg lnTp lnSULT_VmaxC lnUGT_Km lnUGT_VmaxC lnVmax_AG lnKm_AS lnVmax_AS lnkGA_syn lnkPAPS_syn lnCLC_APAP lnCLC_AG lnCLC_AS 
+#> 
+#> Interaction:
+#>  lnSULT_Km_apap lnSULT_VmaxC lnUGT_VmaxC lnKm_AG lnVmax_AG lnKm_AS lnVmax_AS lnkGA_syn lnkPAPS_syn lnCLC_APAP lnCLC_AG lnCLC_AS 
+#> 
+#> Total order:
+#>  lnTg lnTp lnSULT_Km_apap lnSULT_VmaxC lnUGT_Km lnUGT_VmaxC lnKm_AG lnVmax_AG lnKm_AS lnVmax_AS lnkGA_syn lnkPAPS_syn lnCLC_APAP lnCLC_AG lnCLC_AS
+\end{CodeOutput}
+\end{CodeChunk}
+
+The \texttt{check()} also provides some feasible argument to specify the
+target output or change the cut-off value.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> check(x, vars = "lnCPL_APAP_mcgL", SI.cutoff = 0.1, CI.cutoff = 0.1)
+\end{CodeInput}
+
+\begin{CodeOutput}
+#> 
+#> Sensitivity check ( Index > 0.1 )
+#> ----------------------------------
+#> First order:
+#>  lnTg lnTp lnSULT_VmaxC lnCLC_APAP 
+#> 
+#> Interaction:
+#>  lnSULT_VmaxC lnkPAPS_syn 
+#> 
+#> Total order:
+#>  lnTg lnTp lnSULT_VmaxC lnkPAPS_syn lnCLC_APAP 
+#> 
+#> Unselected factors in total order:
+#>  lnCYP_Km lnCYP_VmaxC lnSULT_Km_apap lnSULT_Ki lnSULT_Km_paps lnUGT_Km lnUGT_Ki lnUGT_Km_GA lnUGT_VmaxC lnKm_AG lnVmax_AG lnKm_AS lnVmax_AS lnkGA_syn lnCLC_AG lnCLC_AS 
+#> 
+#> 
+#> Convergence check ( Index > 0.1 )
+#> ----------------------------------
+#> First order:
+#>  lnTg lnCLC_APAP 
+#> 
+#> Interaction:
+#>  lnkPAPS_syn 
+#> 
+#> Total order:
+#>  lnTg lnSULT_VmaxC lnkPAPS_syn lnCLC_APAP
+\end{CodeOutput}
+\end{CodeChunk}
+
+\hypertarget{heatmap-visualization-combined-with-an-index-cut-off}{%
+\subsubsection{Heatmap visualization combined with an index
+``cut-off''}\label{heatmap-visualization-combined-with-an-index-cut-off}}
+
+Based on our previous study, we proposed the heatmap visualization
+approach to distinguish ``influential'' and ``non-influential''
+parameters with a cut-off. Through the given argument \texttt{order}, we
+can select the specific order of sensitivity measurement that we're
+interested in (Figure 3 \& 4).
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> heat_check(x, order = "interaction")
+\end{CodeInput}
+
+\begin{center}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-27-1} \end{center}
+
+\end{CodeChunk}
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> heat_check(x, order = "total order")
+\end{CodeInput}
+
+\begin{center}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-28-1} \end{center}
+
+\end{CodeChunk}
+
+Also, adding the \texttt{index\ =\ "CI"} in the function can further
+investigate the convergence of the sensitivity index. Based on the
+current setting of sampling number, most parameters cannot reach the
+acceptable criteria of convergence. Therefore, the higher number of
+sampling is necessary.
+
+\begin{CodeChunk}
+
+\begin{CodeInput}
+R> heat_check(x, index = "CI", CI.cutoff = 0.05)
+\end{CodeInput}
+
+\begin{center}\includegraphics{manuscript_files/figure-latex/unnamed-chunk-29-1} \end{center}
+
+\end{CodeChunk}
 
@@
 
-\hypertarget{r-code}{%
-\section{R code}\label{r-code}}
+\hypertarget{concluding-remarks}{%
+\section{Concluding remarks}\label{concluding-remarks}}
 
-Can be inserted in regular R markdown blocks.
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
 
-\begin{CodeChunk}
-
-\begin{CodeInput}
-R> x <- 1:10
-R> x
-\end{CodeInput}
-
-\begin{CodeOutput}
- [1]  1  2  3  4  5  6  7  8  9 10
-\end{CodeOutput}
-\end{CodeChunk}
+This work was funded, in part, by grant 1U01FD005838 from the U.S. FDA.
+This abstract reflects the views of the author and should not be
+construed to represent FDA's views or policies. We thank Dr.~Yasha
+Hartberg and Dr.~Barbara Gastel at the Texas A\&M University for
+reviewing the manuscript and consultation.
 

### 2019-01-31 Example/Example.Rmd
Add example of problem

---
author:
  - name: FirstName LastName
    affiliation: University/Company
    address: >
      First line
      Second line
    email: \email{name@company.com}
    url: http://rstudio.com
  - name: Second Author
    affiliation: Affiliation
title:
  formatted: "A Capitalized Title: Something about a Package \\pkg{foo}"
  # If you use tex in the formatted title, also supply version without
  plain:     "A Capitalized Title: Something about a Package foo"
  # For running headers, if needed
  short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The abstract of the article.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: >
  \usepackage{amsmath}
bibliography: index.bib   
output: rticles::jss_article
---

# Example

* This is unwanted output

[@R-sensitivity]

* This is what I wany to show in the paper

[@R-sensitivity]

# Referenes

```{r, include=FALSE}
knitr::write_bib('sensitivity', 'index.bib')
```

### 2019-01-31 Example/Example.Rmd
typo
Cannot solve citation problem

@@
     affiliation: University/Company
-    address: >
-      First line
-      Second line
-    email: \email{name@company.com}
-    url: http://rstudio.com
-  - name: Second Author
-    affiliation: Affiliation
 title:
   formatted: "A Capitalized Title: Something about a Package \\pkg{foo}"
-  # If you use tex in the formatted title, also supply version without
   plain:     "A Capitalized Title: Something about a Package foo"
-  # For running headers, if needed
   short:     "\\pkg{foo}: A Capitalized Title"
@@
 keywords:
-  # at least one keyword must be supplied
   formatted: [keywords, not capitalized, "\\proglang{Java}"]
@@
 
-* This is what I wany to show in the paper
+* This is what I want to show in the paper
 

### 2019-01-31 html/manuscript.Rmd
Cannot solve citation problem

---
author:
  - name: Yihui Xie
    affiliation: RStudio
    address: |
      First line  
      Second line
    email: <name@company.com>
    url: <https://rstudio.com>
  - name: Romain Lesur
    affiliation: Affiliation
title: "A Capitalized Title: Something about a Package **foo**"
shorttitle: "**foo**:  A Capitalized Title"
abstract: |
  The abstract of the article.
# at least one keyword must be supplied
keywords: [keywords, not capitalized, "[Java]{.proglang}"]
links-to-footnotes: true
paged-footnotes: true
bibliography: index.bib
output: 
  pagedown::jss_paged:
    self_contained: false
---

# Introduction

This template demonstrates some of the basic [markdown]{.proglang} you'll need to know to create a JSS article with **pagedown** [@R-pagedown].

## Code formatting

You should use **Pandoc**'s [markdown]{.proglang} but you also can use $\LaTeX$ commands:

* [Java]{.proglang} or $\proglang{Java}$ 
* **plyr**, [plyr]{.pkg} or $\pkg{plyr}$
* `print("abc")` or $\code{print("abc")}$

# [R]{.proglang} code

Can be inserted in regular [R Markdown]{.proglang} blocks.

```{r}
x <- 1:10
x
```

# Render the document

If you use the option `self_contained: false` (see line \#22 of this `Rmd` file), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@R-xaringan] RStudio add-in _Infinite Moon Reader_.

[@R-httk]

# References {-}

```{r, include=FALSE}
knitr::write_bib(c(.packages(), 'pagedown', 'xaringan', 'httk'), 'index.bib')
```

### 2019-01-31 manuscript/manuscript.Rmd
Cannot solve citation problem

@@
   \usepackage{amsmath}
+bibliography: references.bib 
 output: rticles::jss_article
-#output: rmarkdown::html_vignette
+
 ---
@@
 # Introduction
+
+[@R-sensitivity]
+
+[@R-sensitivity]
 
@@
 Most of sensitivity analysis tools are constructed using scientific computing software, such as GNU Octave/Matlab (Pianosi et al., 2015), python (Herman and Usher, 2017), and R (Ghanem et al., 2017).
-Previously, we mainly used R with the sensitivity package (Pujol et al., 2017), which is a practical tool to conduct local and global sensitivity analysis.
+Previously, we mainly used R with the \pkg{sensitivity} package [@R-sensitivity], which is a practical tool to conduct local and global sensitivity analysis.
 The sensitivity package includes some functions to generate the parameter sequences by using different computing algorithms.
@@
 We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package (Saltelli et al., 1999).
-Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the R \pkg{sensitivity} package (Pujol et al., 2017).
+Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the R \pkg{sensitivity} package [-@R-sensitivity].
 To test the convergence and robustness of the sensitivity measurement, we included a random phase-shift approach to replicate sampling from random starting points across parameter space.
@@
 
-```{r, eval=T}
+```{r, eval=F}
 pbtk1cpt_model()
@@
 
-```{r, eval=T}
+```{r, eval=F}
 mName <- "pbtk1cpt"
@@
 
-```{r, eval=T}
+```{r, eval=F}
 parms <- initParms()
@@
 
-```{r, eval=T}
+```{r, eval=F}
 parms
@@
 
-```{r, eval=T}
+```{r, eval=F}
 initState
@@
 
-```{r, eval=T}
+```{r, eval=F}
 Outputs
@@
 
-```{r, eval=T}
+```{r, eval=F}
 times <- seq(from = 0.01, to = 24.01, by = 1)
@@
 
-```{r, eval=T, fig.cap = 'Figure 1. Simulation results of one-compartment PBTK model.'}
+```{r, eval=F, fig.cap = 'Figure 1. Simulation results of one-compartment PBTK model.'}
 plot(y)
@@
 
-```{r, eval=T}
+```{r, eval=F}
 LL <- 0.5 
@@
 
-```{r, eval=T}
+```{r, eval=F}
 y <- solve_fun(x, times, initState = initState,
@@
 
-```{r, eval=T, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24 hour time period intake.'}
+```{r, eval=F, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24 hour time period intake.'}
 plot(x)
@@
 
-```{r, eval=T, fig.cap = 'Figure 3. The relationship between model parameter and estimated concentration under the time-point of 0.01, 2.01, and 24.01 hr'}
+```{r, eval=F, fig.cap = 'Figure 3. The relationship between model parameter and estimated concentration under the time-point of 0.01, 2.01, and 24.01 hr'}
 par(mfrow = c(3,3), mar = c(2,2,2,2), oma = c(2,2,1,1))
@@
 
-```{r, eval=T}
+```{r, eval=F}
 dim(x$a)
@@
 
-```{r, eval=T}
+```{r, eval=F}
 dim(y)
@@
 
-```{r, eval=T}
+```{r, eval=F}
 check(x, SI.cutoff = 0.5)
@@
 
-```{r, eval=T}
+```{r, eval=F}
 system.time(y<-solve_fun(x, times, initState = initState, 
@@
 
-```{r, eval=T}
+```{r, eval=F}
 compile_model(mName, application = "mcsim")
@@
 
-```{r, eval=T}
+```{r, eval=F}
 conditions <- c("Agutlument = 10") # Set the initial state of Agutlument = 10 
@@
 
-```{r, eval=T}
+```{r, eval=F}
 # Nominal value
@@
 
-```{r, eval=T}
+```{r, eval=F}
 params <- c("lnTg", "lnTp", "lnCYP_Km","lnCYP_VmaxC",
@@
 
-```{r, eval=T}
+```{r, eval=F}
 mName <- "pbpk_apap"
@@
 
-```{r, eval=T}
+```{r, eval=F}
 vars <- c("lnCPL_APAP_mcgL", "lnCPL_AG_mcgL", "lnCPL_AS_mcgL")
@@
 
-```{r, eval=T, fig.height=8, fig.width=8, fig.cap = 'Figure 4. '}
+```{r, eval=F, fig.height=8, fig.width=8, fig.cap = 'Figure 4. '}
 plot(x, vars = "lnCPL_AG_mcgL")
@@
 
-```{r, eval=T, fig.cap = 'Figure 5. '}
+```{r, eval=F, fig.cap = 'Figure 5. '}
 par(mfrow = c(2,2), mar = c(2,2,1,1), oma = c(2,2,1,1))
@@
 
-```{r, eval=T}
+```{r, eval=F}
 check(x)
@@
 
-```{r, eval=T}
+```{r, eval=F}
 check(x, vars = "lnCPL_APAP_mcgL", SI.cutoff = 0.1, CI.cutoff = 0.1)
@@
 
-```{r, eval=T}
+```{r, eval=F}
 heat_check(x, order = "interaction")
@@
 
-```{r, eval=T}
+```{r, eval=F}
 heat_check(x, order = "total order")
@@
 
-```{r, eval=T, fig.height=9}
+```{r, eval=F, fig.height=9}
 heat_check(x, index = "CI", CI.cutoff = 0.05)
@@
 # References
+
+```{r, include=FALSE}
+knitr::write_bib(c('sensitivity'))
+```

### 2019-01-31 manuscript/manuscript.Rmd
update

@@
 
-[@R-sensitivity]
-
-[@R-sensitivity]
-
 Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs.
@@
 compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants (McNally et al., 2011).
-It can be further applied to parameter prioritization and parameter fixing before model calibration (Hsieh et al., 2018).
+It can be further applied to parameter prioritization and parameter fixing before model calibration [@fphar201800588].
 
-In our previous work (Hsieh et al., 2018), we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+In our previous work [@fphar201800588], we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
 We used GNU MCSim (Bois, 2009), an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
@@
 ```{r, include=FALSE}
-knitr::write_bib(c('sensitivity'))
+knitr::write_bib(c('sensitivity','FME', 'multisensi', 'ODEsensitivity'))
 ```

### 2019-02-01 manuscript/manuscript.Rmd
update manuscript

--- 2019-01-31 manuscript/manuscript.Rmd
@@
 Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs.
-An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions (Ferretti et al., 2016).
-Also, this technique has been applied in pharmacology and toxicology research (Loizou et al., 2015; McNally et al., 2012; Scherholz et al., 2018).
+An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions \citep{ferretti2016trends}.
+Also, this technique has been applied in pharmacology and toxicology research \citep{loizou2015application, mcnally2012reconstruction}.
 Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time.
 The goal of sensitivity analysis in pharmacokinetic research is to examine the sensitivity of output variables (e.g.
-compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants (McNally et al., 2011).
-It can be further applied to parameter prioritization and parameter fixing before model calibration (Hsieh et al., 2018).
+compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants \citep{mcnally2011workflow}.
+It can be further applied to parameter prioritization and parameter fixing before model calibration \citep{fphar201800588}.
 
-In our previous work (Hsieh et al., 2018), we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-We used GNU MCSim (Bois, 2009), an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
+In our previous work \citep{fphar201800588}, we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+We used GNU MCSim \citep{JSSv002i09}, an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
 We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of global sensitivity analysis algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
@@
 
-Most of sensitivity analysis tools are constructed using scientific computing software, such as GNU Octave/Matlab (Pianosi et al., 2015), python (Herman and Usher, 2017), and R (Ghanem et al., 2017).
-Previously, we mainly used R with the sensitivity package (Pujol et al., 2017), which is a practical tool to conduct local and global sensitivity analysis.
+Most of sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} \citep{pianosi2015matlab}, \proglang{python} \citep{herman2017salib}, and \proglang{R} \citep{iooss2017introduction}.
+Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package \citep{R-sensitivity}, which is a practical tool to conduct local and global sensitivity analysis.
 The sensitivity package includes some functions to generate the parameter sequences by using different computing algorithms.
 Also, it provides a feasible way to integrate external modeling results.
-This computational approach can effectively solve the heavy computing burden of numerical solutions within the pure R environment.
-Several R packages, such as FME, multisensi, and ODEsensitivity include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network.
-The FME package contains a basic method for performing local sensitivity analysis for dynamic models (Soetaert and Petzoldt, 2016).
-The multisensi package provides sensitivity analysis for models with multivariate output (Bidot et al., 2018).
-ODEsensitivity can perform sensitivity analysis in ordinary differential equation (ODE) models (Weber et al., 2018).
+This computational approach can effectively solve the heavy computing burden of numerical solutions within the pure \proglang{R} environment.
+Several \proglang{R} packages, such as \pkg{FME}, \pkg{multisensi}, and \pkg{ODEsensitivity} include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network (CRAN).
+The \pkg{FME} package contains a basic method for performing local sensitivity analysis for dynamic models \citep{JSSv033i03}.
+The \pkg{multisensi} package provides sensitivity analysis for models with multivariate output \citep{R-multisensi}.
+\pkg{ODEsensitivity} can perform sensitivity analysis in ordinary differential equation (ODE) models \citep{R-ODEsensitivity}.
 It uses the ODE interface from the deSolve package, connecting it with the sensitivity analysis from the sensitivity package, and can perform sensitivity analysis on models with multivariate output.
+Other related packages such as \pkg{mtk} \citep{RJ-2015-031} and \pkg{fast} \citep{R-fast} are designed to invetigate the influence of model parameter.
 Although these tools provide various approaches to perform sensitivity analysis, we have not find a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.
-
-Here we develop an R package, called pksensi, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
+Here we develop an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK, and advanced compartment absorption and transit models with multivariate model output.
-The design concepts of pksensi are:
+The design concepts of \pkg{pksensi} are:
 
@@
 *	Freely available: All related packages are free and open source
-*	Integrated application: Users can run pharmacokinetic models in R with script that were written in C or GNU MCSim 
-*	Decision making: The output results and visualization tools can be used to easily determine which parameters have “non-influential” effects on the model output and can be fixed in model calibration.
+*	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \proglang{GNU MCSim} 
+*	Decision making: The output results and visualization tools can be used to easily determine which parameters have "non-influential" effects on the model output and can be fixed in model calibration.
 
@@
 
-We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package (Saltelli et al., 1999).
-Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the R \pkg{sensitivity} package (Pujol et al., 2017).
+We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package \citep{saltelli1999quantitative}.
+Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package \citep{R-sensitivity}.
 To test the convergence and robustness of the sensitivity measurement, we included a random phase-shift approach to replicate sampling from random starting points across parameter space.
@@
 where $x_i$ is the nominal value of the i-th parameter, and $\omega_i$  is a vector giving the set of frequencies, one frequency for each parameter.
-The default set of frequencies is based on the study from Saltelli et al.
-(1999).
+The default set of frequencies is based on the study from \cite{saltelli1999quantitative}.
 The first frequency of the vector $\omega$ is assigned to each factor $x_i$ in turn (corresponding to the estimation of Sobol indices of first and total order, respectively).
@@
 
-To investigate the convergence of sensitivity indices, we adopted the exam approach proposed by Sarrazin et al.
-(2016).
+To investigate the convergence of sensitivity indices, we adopted the exam approach proposed by \citep{sarrazin2016global}.
 This method quantitatively assesses convergence by computing the width of 95% confidence intervals for all parameters across all time-points and output variables.
-To integrate the random phase-shift with eFAST, we built the \proglang{rfast99} function, which can be used to perform the sensitivity test for the whole pharmacokinetic modeling process.
-The function \proglang{rfast99} in pksensi can sample and generate the testing parameter matrix based on a given argument of sample size n and probability distribution (e.g., mean/s.d.
+To integrate the random phase-shift with eFAST, we built the \code{rfast99} function, which can be used to perform the sensitivity test for the whole pharmacokinetic modeling process.
+The function \code{rfast99} in \pkg{pksensi} can sample and generate the testing parameter matrix based on a given argument of sample size n and probability distribution (e.g., mean/s.d.
 of normal distribution, meanlog/sdlog of log-normal distribution, and minimum/maximum of uniform distribution).
@@
 The \pkg{pksensi} provides a useful function to solve ordinary differential equations in pharmacokinetic models using analytical and numerical approaches.
-The solution can be performed under the pure R coding environment by linking pksensi with the deSolve package (Soetaert et al., 2018).
-However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as FORTRAN, C, or C++.
+The solution can be performed under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package \citep{JSSv033i09}.
+However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as \proglang{FORTRAN}, \proglang{C}, or \proglang{C++}.
 The \pkg{pksensi} includes a function to compile and create dynamic-link libraries (.dll) on Windows and shared objects (.so) on Unix-liked systems (e.g., Linux and MacOS).
-This compiled file can load and execute through the build-in function in R.
+This compiled file can load and execute through the build-in function in \proglang{R}.
 However, Windows users need to install Rtools or MinGW to compile the source code by using the GNU GCC compiler.
-More requirements are detailed in the study from Soetaert et al.
-(2010).
-The C implementation of the example model can be found as an example within MCSim v6.0.1 (Bois, 2009).
-The \pkg{pksensi} can also link with MCSim to compile the model code, used in solving each system of equations.
+More requirements are detailed in the study from \citep{soetaertr}.
+The \proglang{C} implementation of the example model can be found as an example within \proglang{GNU MCSim} v6.0.1 \citep{JSSv002i09}.
+The \pkg{pksensi} can also link with \proglang{GNU MCSim} to compile the model code, used in solving each system of equations.
 
@@
 The output of sensitivity analysis returns a list that contains the given time-points and values of all state variables (e.g., chemical concentrations in blood).
-This format is particularly suited for further graphical routines within pkseni.
+This format is particularly suited for further graphical routines within \pkg{pksensi}.
 Also, a plot can can create time-dependent sensitivity measurements of first and total order effect for all parameters in one figure.
-Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
-We also developed a “cut-off”-based approach in this package to accurately distinguish between “influential” and “non-influential” parameters.
-Finally, pksensi includes numerous visualization tools for the effective investigation and communication of results in decision making.
+Furthermore, \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
+We also developed a "cut-off"-based approach in this package to accurately distinguish between "influential" and "non-influential" parameters.
+Finally, \pkg{pksensi} includes numerous visualization tools for the effective investigation and communication of results in decision making.
 
@@
 
-## One-compartment taxicokinetic model
+## One-compartment toxicokinetic model
 
@@
 library(pksensi)
-#mcsim_install()
-
+options(xtable.comment = FALSE)
 knitr::opts_chunk$set(
@@
 )
-
-fn = local({ # not used function for fig caption
-  i = 0
-  function(x) {
-    i <<- i + 1
-    paste('Figure ', i, ': ', x, sep = '')
-  }
-})
-
 ```
@@
 
-In this example, We use a simple, one-compartment PK model from `httk` package [@JSSv079i04] to demonstrate how `pksensi` can be applied to pharmacokinetic studies.
+In this example, We use a simple, one-compartment PK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to pharmacokinetic studies.
 The differential equations for the one-compartment pharmacokinetic model can be written as:
@@
 
-In the beginning, we need to pre-install GNU MCSim [@JSSv002i09] and related compiler to let us generate .c file and executable file.
-The GNU MCSim can be installed by following the instruction in GNU MCSim’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function `mcsim_install()` in `pksensi`.
+In the beginning, we need to pre-install \pkg{GNU MCSim} \citep{JSSv002i09} and related compiler to let us generate .c file and executable file.
+The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_install()} in \pkg{pksensi}.
 The GNU compiler is necessary for users that use Linux or MacOS.
-For Windows users, you should install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and use `Sys.setenv()` to set the working path of compiler.
-The  `Sys.which("gcc")` and `system('g++ -v')` can check whether we can run compiler correctly.
+For Windows users, you should install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and use \code{Sys.setenv()} to set the working path of compiler.
+In \proglang{R} environment, The  \code{Sys.which("gcc")} and \code{system('g++ -v')} can check whether we can run compiler correctly.
 
-We first implemented this model in R by compiling the file written in C.
-pksensi allows users to select the preferred method to solve the pharmacokinetic model, either with the `deSolve` package or with GNU MCSim through the compile function.
-However, running model under GNU MCSim native code can have faster speed to obtain the model outputs.
+We first implemented this model in \proglang{R} by compiling the file written in \proglang{C}.
+\pkg{pksensi} allows users to select the preferred method to solve the pharmacokinetic model, either with the pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
+However, running model under \pkg{GNU MCSim} native code can have faster speed to obtain the model outputs.
 
-The following R script can download and compile the model description file (`pbtk1cpt.model`) and use `deSolve` package [@JSSv033i09] to solve ordinary differential equations in our model.
-The example model code of one-compartment PBTK model is available with pksensi package:
+The following \proglang{R} script can download and compile the model description file \code{pbtk1cpt.model} and use code{ode} function in \pkg{deSolve} package \citep{JSSv033i09} to solve ordinary differential equations in our model.
+The example model code of one-compartment PBTK model is available with \pkg{pksensi} package:
 
@@
 
-Then, use `compile_model()` to generate the executable files (`pbtk1cpt.dll` on Windows or `pbtk1cpt.so` on other systems) and R file (`pbtk1cpt_inits.R`) with default input parameters and initial state settings with the definition of `application = "R"`.
+Then, use \code{compile_model()} to generate the executable files (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
 
@@
 In the current setting, we assumed the initial condition of the intake chemical to be 10 mol.
-The `initParms` and `initStates` functions were used to customize the parameter values and the initial state that will be used in the `solve_fun` function.
-These parameter value can be adopted from the `httk` package, which includes physico-chemical and drug biological properties for 553 chemicals.
+The \code{initParms} and \code{initStates} functions were used to customize the parameter values and the initial state that will be used in the \code{solve_fun()} function.
+These parameter value can be adopted from the \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
 In this case, we used the parameter value of theophylline in this example.
-The given `vdist`, `ke`, and `kgutabs` are 0.74, 0.28, and 2.18, respectively.
+The given \code{vdist}, \code{ke}, and \code{kgutabs} are 0.74, 0.28, and 2.18, respectively.
 
-Through `ode` function in `deSolve` package, we can visualize the pharmacokinetic according to the given parameter conditions such as time points (`times`):
+Through \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic according to the given parameter conditions such as time points (\code{times}):
 
 ```{r, eval=T}
+library(deSolve)
 times <- seq(from = 0.01, to = 24.01, by = 1)
-y <- deSolve::ode(initState, times, func = "derivs", parms = parms, 
+y <- ode(initState, times, func = "derivs", parms = parms, 
          dllname = mName, initfunc = "initmod", nout = 1, outnames = Outputs)
@@
 We assume a uniform distribution for the estimate for each parameter with the coefficient of uncertainty within 50%.
-The parameter ranges are assumed to be (0.37, 1.12) for `vdist`, (0.0058, 0.0174) for `ke`, and (0.045, 0.136) for `kgutabs`.
+The parameter ranges are assumed to be (0.37, 1.12) for \code{vdist}, (0.0058, 0.0174) for \code{ke}, and (0.045, 0.136) for \code{kgutabs}.
 The sample number determines the robustness of the result of sensitivity analysis.
@@
 To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours in the example.
-The `initParmsfun` is used to generate the sampling value for each parameter.
-The `outnames`, `dllname`, `func`, `initfunc` are based on the arguments from the ode function in `deSolve` package.
-The details of the model structure and these arguments are defined in `pbtk1comp.c`.
-and `pbtk1comp_inits.R`.
-Finally, the `tell2` function is used to integrate the parameter values and the output results of numerical analysis that were generated and stored in variables x and y.
-The result of object x is an object of rfast99, which has specific `print`, `plot`, and `check` method.
+The \code{initParmsfun} is used to generate the sampling value for each parameter.
+The \code{outnames}, \code{dllname}, \code{func}, \code{initfunc} are based on the arguments from the ode function in \pkg{deSolve} package.
+The details of the model structure and these arguments are defined in \code{pbtk1comp.c}.
+and \code{pbtk1comp_inits.R}.
+Finally, the \code{tell2()} function is used to integrate the parameter values and the output results of numerical analysis that were generated and stored in variables x and y.
+The result of object x is an object of rfast99, which has specific \code{print()}, \code{plot()}, and \code{check()} method.
 The print function gives the sensitivity and convergence indices for main, interaction, and total order at each time point.
@@
 
-```{r, eval=T, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24 hour time period intake.'}
+```{r, eval=T, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24-hour time period intake.'}
 plot(x)
@@
 
-Here, we can find that `vdist` and `ke` are dominating the plasma concentration in the before and after 5-hour post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
-Besides, the `kgutabs` only plays a crucial role to determine the plasma concentration in the first hour.
+Here, we can find that \code{vdist} and \code{ke} are dominating the plasma concentration in the before and after 5-hour post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
+Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
 The relationship between concentration and the parameters can be plotted as follow (Figure 3):
@@
 
-The x is a list of class "rfast99", containing all the input arguments detailed before and the calculated sensitivity indices of first order (`mSI`), interaction (`iSI`), and total order (`tSI`).
-The convergence indices are also stored in the list named `mCI`, `iCI`, and `tCI`.
-The parameter values are stored in an array `x$a` with c(model evaluation, replication, parameters).
+The \code{x} is a list of class \code{rfast99}, containing all the input arguments detailed before and the calculated sensitivity indices of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
+The convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
+The parameter values are stored in an array \code{x$a} with c(model evaluation, replication, parameters).
 
@@
 
-The `check()` is a useful function to determine which parameters have relative lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
-The argument of `SI.cutoff` is setting at 0.5 to detect the relative non-influential parameters in this case.
+The \code{check()} is a useful function to determine which parameters have relative lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
+The argument of \code{SI.cutoff} is setting at 0.5 to detect the relative non-influential parameters in this case.
 
@@
 
-Based on the sensitivity measurement of the total order, the result shows that `kgutabs` has relative lower measurement of sensitivity index.
+Based on the sensitivity measurement of the total order, the result shows that \code{kgutabs} has relative lower measurement of sensitivity index.
 
@@
 
-Alternatively, to solve ODE by using GNU MCSim, we need to change the argument to `application = mcsim` in `compile_model()`.
-Rather than apply R `deSolve` to solve differential equations, the GNU MCSim can provide higher computational speed in global sensitivity analysis.
+Alternatively, to solve ODE by using \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model()}.
+Rather than apply \pkg{deSolve} to solve differential equations, the \pkg{GNU MCSim} can provide higher computational speed in global sensitivity analysis.
 
@@
 
-Sililiar to `solve_fun` that can define the initial parameter and state values through input function, the `solve_mcsim` has a `condition` argument that is used to givien the specific input value such as oral dose or fixing parameter value or initial state variable.
+Similar to \code{solve_fun()} function that can define the initial parameter and state values through input function, the \code{solve_mcsim()} has a `condition` argument that is used to givien the specific input value such as oral dose or fixing parameter value or initial state variable.
 
@@
 
-Under the same given condition, it takes 5-6 (`deSolve`) and 1-2 (GNU MCSim) seconds to solve model. The `solve_mcsim()` shows the better computational performance than `solve_fun()` in `pksensi`.
+Under the same given condition, it takes 5-6 seconds (\pkg{deSolve}) and 1-2 seconds (\pkg{GNU MCSim})  to solve model.
+The \code{solve_mcsim()} shows the better computational performance than \code{solve_fun()} in \pkg{pksensi}.
 
@@
 
-The aim of this section is to reproduce our previous published [@fphar201800588] result of global sensitivity analysis for acetaminophen PBPK model through `pksensi`.
-The model codes are included in this package and can be generated through `pbpk_apap_model()`.
+The aim of this section is to reproduce our previous published result \citep{fphar201800588} of global sensitivity analysis for acetaminophen PBPK model through \pkg{pksensi}.
+The model codes are included in this package and can be generated through \code{pbpk_apap_model()}.
 We applied the global sensitivity analysis workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
@@
 
-<Table 1>
+```{r, results='asis',message=FALSE,warning=FALSE, echo = F}
+library(xtable)
+
+#Nominal value
+Tg <- log(0.23)
+Tp <- log(0.033)
+CYP_Km <- log(130)
+SULT_Km_apap <- log(300)
+SULT_Ki <- log(526)
+SULT_Km_paps <- log(0.5)
+UGT_Km <- log(6.0e3)
+UGT_Ki <- log(5.8e4)
+UGT_Km_GA <-log(0.5)
+Km_AG <- log(1.99e4)
+Km_AS <- log(2.29e4)
+
+r <- 1.96 # exp(1.96)/exp(-1.96) ~ 50
+
+x <- c("Tg", "Tp", "CYP\\_Km", "CYP\\_VmaxC",
+       "SULT\\_Km\\_apap","SULT\\_Ki","SULT\\_Km\\_paps","SULT\\_VmaxC",
+       "UGT\\_Km","UGT\\_Ki","UGT\\_Km\\_GA","UGT\\_VmaxC",
+       "Km\\_AG","Vmax\\_AG","Km\\_AS","Vmax\\_AS",
+       "kGA\\_syn","PAPS\\_syn", "CLC\\_APAP","CLC\\_AG","CLC\\_AS"
+       )
+
+y <- c("Gatric emptying time constant",
+       "GI perfusion time constant",
+       "Cytochrome P450 metabolism, Km",
+       "Cytochrome P450 metabolism, VMax",
+       "Sulfation pathway acetaminophen, Km",
+       "Sulfation pathway substrate inhibition, Ki",
+       "Sulfation pathway PAPS, Km",
+       "Sulfation pathway acetaminophen, Vmax",
+       "Glucronidation pathway acetaminophen, Km",
+       "Glucronidation pathway substrate inhibition, Ki",
+       "Glucronidation pathway GA, Km",
+       "Glucronidation pathway acetaminophen, Vmax",
+       "APAP-G hepatic transporter, Km",
+       "APAP-G hepatic transporter, Vmax",
+       "APAP-S hepatic transporter, Km",
+       "APAP-S hepatic transporter, Vmax",
+       "UDPGA synthesis",
+       "PAPS synthesis",
+       "APAP clearance",
+       "APAP-G clearance",
+       "APAP-S clearance"
+       )
+
+z <- c("$h$", "$h$", "$\\mu{M}$", "$\\mu{mole}/h\\cdot{BW}^{0.75}$",
+       "$\\mu{M}$", "$\\mu{M}$", "$-$", "$\\mu{mole}/h\\cdot{BW}^{0.75}$",
+       "$\\mu{M}$", "$\\mu{M}$", "$-$", "$\\mu{mole}/h\\cdot{BW}^{0.75}$",
+       "$\\mu{M}$", "$\\mu{mole}/h$", "$\\mu{M}$", "$\\mu{mole}/h$",
+       "$1/h$", "$1/h$",
+       "$L/h\\cdot{BW}^{0.75}$", "$L/h\\cdot{BW}^{0.75}$", "$L/h\\cdot{BW}^{0.75}$"
+       )
+
+min <- c(round(Tg-r, 3), round(Tp-r, 3), round(CYP_Km-r), round(log(0.14), 3),
+         round(SULT_Km_apap-r, 3), round(SULT_Ki-r, 3), round(SULT_Km_paps-r), log(1),
+         round(UGT_Km-r, 3), round(UGT_Ki-r, 3), round(UGT_Km_GA-r), log(1),
+         round(Km_AG-r, 3), round(log(1.09e3), 3), round(Km_AS-r), round(log(1.09e3), 3),
+         log(1), log(1), round(log(2.48e-3), 3), round(log(2.48e-3), 3), round(log(2.48e-3), 3)
+         )
+
+max <- c(round(Tg+r, 3), round(Tp+r, 3), round(CYP_Km+r), round(log(2900), 3),
+         round(SULT_Km_apap+r, 3), round(SULT_Ki+r, 3), round(SULT_Km_paps+r), round(log(22026), 3),
+         round(UGT_Km+r, 3), round(UGT_Ki+r, 3), round(UGT_Km_GA+r), round(log(22026), 3),
+         round(Km_AG+r, 3), round(log(3.26e6), 3), round(Km_AS+r), round(log(3.26e6), 3),
+         round(log(4.43e5), 3), round(log(4.43e5),3), 
+         round(log(2.718), 3), round(log(2.718), 3), round(log(2.718), 3)
+         )
+
+df <- data.frame(x, y, z, min, max)
+names(df) <- c("Parameter","Description", "Unit", "Min", "Max")
+
+print(xtable(df, caption = "Description of sampling range of model parameter"),
+      type = "latex", 
+      caption.placement = "top", include.rownames = FALSE,
+      sanitize.text.function = function(x) {x})
+```
 
@@
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
-The `generate_infile()` provide the arguments of `rtol` and `atol` that can be adjusted to prevent the unwanted error.
+The \code{generate_infile()} provide the arguments of \code{rtol} and \code{atol} that can be adjusted to prevent the unwanted error.
 However, the modification will slow down the computational speed.
@@
 
-In this test case, we adjusted the range of `SULT_VmaxC` and `UGT_VmaxC` from U(0, 15) to U(0, 10).
-The relative and absolute error tolerance were set to 1e-7 and 1e-9, respectively, to prevent the computational error in MCSim,
+In this test case, we adjusted the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} from U(0, 15) to U(0, 10).
+The relative and absolute error tolerance were set to 1e-7 and 1e-9, respectively, to prevent the computational error in \pkg{GNU MCSim},
 
@@
 
-After creating the `pbpk_apap.model` in the working directory, the next step is to generate the executable files (`mcsim.pbpk_apap`) through `compile_model()`.
+After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable files (\code{mcsim.pbpk_apap}) through \code{compile_model()}.
 
@@
 
-To improve the computational speed, this case only uses MCSim to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
+To improve the computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
 The setting oral dose of APAP is 20 mg/kg in this example.
-Generally, the input dosing method can be defined through the `condition` argument.
-Since the unit of the given dose is mg/kg, the `mgkg_flag` is set to 1 to declare the statement.
-More definition of input can be found in the section of input functions in GNU MCSim User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
+Generally, the input dosing method can be defined through the \code{condition} argument.
+Since the unit of the given dose is mg/kg, the \code{mgkg_flag} is set to 1 to declare the statement.
+More definition of input can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
 
@@
 
-The plotting function can output the result of time-dependent sensitivity measurement to determine the parameter impact on model output over time (Figure 1).
+The plotting function can output the result of time-dependent sensitivity measurement to determine the parameter impact on model output over time (Figure 4).
 
-```{r, eval=T, fig.height=8, fig.width=8, fig.cap = 'Figure 4. '}
+```{r, eval=T, fig.height=8, fig.width=8, fig.cap = ' '}
 plot(x, vars = "lnCPL_AG_mcgL")
@@
 The uncertainty analysis is a crucial step before model calibration.
-We can apply uncertainty analysis through the `pksim()` by the given name of output variables (Figure 2).
+We can apply uncertainty analysis through the `pksim()` by the given name of output variables (Figure 5).
 Through this visualization approach, we can recognize whether the simulated outputs can accurately simulate the same concentration profile as the in-vivo experiment under the setting of parameter ranges.
-In Figure 2, all experiment points are included in the intervals, representing the acceptable set of the parameter range.
+In Figure 5, all experiment points are included in the intervals, representing the acceptable set of the parameter range.
 
-```{r, eval=T, fig.cap = 'Figure 5. '}
+```{r, eval=T, fig.cap = ' '}
 par(mfrow = c(2,2), mar = c(2,2,1,1), oma = c(2,2,1,1))
@@
 
-In addition, through using the `check()`, the parameter with sensitivity and convergence indices over the given condition can be easily detected for all output variables.
+In addition, through using the \code{check()}, the parameter with sensitivity and convergence indices over the given condition can be easily detected for all output variables.
 
@@
 
-The `check()` also provides some feasible argument to specify the target output or change the cut-off value.
+The \code{check()} also provides some feasible argument to specify the target output or change the cut-off value.
 
@@
 Based on our previous study, we proposed the heatmap visualization approach to distinguish "influential" and "non-influential" parameters with a cut-off.
-Through the given argument `order`, we can select the specific order of sensitivity measurement that we're interested in (Figure 3 & 4).
+Through the given argument \code{order}, we can select the specific order of sensitivity measurement that we're interested in (Figure 6 & 7).
 
-```{r, eval=T}
+```{r, eval=T, fig.cap = ' '}
 heat_check(x, order = "interaction")
@@
 
-```{r, eval=T}
+```{r, eval=T, fig.cap = ' '}
 heat_check(x, order = "total order")
@@
 Also, adding the `index = "CI"` in the function can further investigate the convergence of the sensitivity index.
-Based on the current setting of sampling number, most parameters cannot reach the acceptable criteria of convergence.
+Based on the current setting of sampling number, most parameters cannot reach the acceptable criteria of convergence (Figure 8).
 Therefore, the higher number of sampling is necessary.
 
-```{r, eval=T, fig.height=9}
-heat_check(x, index = "CI", CI.cutoff = 0.05)
+```{r, eval=T, fig.cap = ' '}
+heat_check(x, index = "CI", CI.cutoff = 0.05, order = "total order")
 ```
-
-## Code formatting
-
-Don't use markdown, instead use the more precise latex commands:
-
-* \proglang{Java}
-* \pkg{plyr}
-* \code{print("abc")}
 
@@
 # References
+
+```{r, include=FALSE}
+knitr::write_bib(c('sensitivity','FME', 'multisensi', 'ODEsensitivity', 'fast'))
+```

### 2019-02-19 manuscript/manuscript.Rmd
Add Frederic Bois

@@
     affiliation: Texas A&M University
+  - name: Frederic Y. Bois
+    affiliation: INERIS 
   - name: Brad Reisfeld

### 2019-02-24 manuscript/manuscript.Rmd
update Introduction

@@
 An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions \citep{ferretti2016trends}.
-Also, this technique has been applied in pharmacology and toxicology research \citep{loizou2015application, mcnally2012reconstruction}.
+This technique has also been applied in pharmacology and toxicology research \citep{loizou2015application, mcnally2012reconstruction}.
 Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time.
@@
 
-Most of sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} \citep{pianosi2015matlab}, \proglang{python} \citep{herman2017salib}, and \proglang{R} \citep{iooss2017introduction}.
+Most sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} \citep{pianosi2015matlab}, \proglang{python} \citep{herman2017salib}, and \proglang{R} \citep{iooss2017introduction}.
 Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package \citep{R-sensitivity}, which is a practical tool to conduct local and global sensitivity analysis.
@@
 This computational approach can effectively solve the heavy computing burden of numerical solutions within the pure \proglang{R} environment.
-Several \proglang{R} packages, such as \pkg{FME}, \pkg{multisensi}, and \pkg{ODEsensitivity} include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network (CRAN).
+Several other \proglang{R} packages, such as \pkg{FME}, \pkg{multisensi}, and \pkg{ODEsensitivity} include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network (CRAN).
 The \pkg{FME} package contains a basic method for performing local sensitivity analysis for dynamic models \citep{JSSv033i03}.
@@
 It uses the ODE interface from the deSolve package, connecting it with the sensitivity analysis from the sensitivity package, and can perform sensitivity analysis on models with multivariate output.
-Other related packages such as \pkg{mtk} \citep{RJ-2015-031} and \pkg{fast} \citep{R-fast} are designed to invetigate the influence of model parameter.
-Although these tools provide various approaches to perform sensitivity analysis, we have not find a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.
-Here we develop an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
+Other related packages such as \pkg{mtk} \citep{RJ-2015-031} and \pkg{fast} \citep{R-fast} are designed to investigate the influence of model parameter.
+Although these tools provide various approaches to perform sensitivity analysis, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.
+Wepresent here an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK, and advanced compartment absorption and transit models with multivariate model output.

### 2019-06-21 manuscript/manuscript.Rmd
manuscript

@@
     affiliation: Texas A&M University
-  - name: Frederic Y. Bois
-    affiliation: INERIS 
   - name: Brad Reisfeld
@@
 
-In this example, We use a simple, one-compartment PK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to pharmacokinetic studies.
-The differential equations for the one-compartment pharmacokinetic model can be written as:
+In this example, we use a simple, one-compartment PBTK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to physiologically based kinetic studies.
+The differential equations for this model is as:
 
@@
 
-In the beginning, we need to pre-install \pkg{GNU MCSim} \citep{JSSv002i09} and related compiler to let us generate .c file and executable file.
+In the beginning, we need to pre-install \pkg{GNU MCSim} \citep{JSSv002i09} and related standard C compiler.
 The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_install()} in \pkg{pksensi}.
-The GNU compiler is necessary for users that use Linux or MacOS.
-For Windows users, you should install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and use \code{Sys.setenv()} to set the working path of compiler.
-In \proglang{R} environment, The  \code{Sys.which("gcc")} and \code{system('g++ -v')} can check whether we can run compiler correctly.
+The C compiler had installed in Linux or MacOS.
+The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler before using this function.
 
-We first implemented this model in \proglang{R} by compiling the file written in \proglang{C}.
-\pkg{pksensi} allows users to select the preferred method to solve the pharmacokinetic model, either with the pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
-However, running model under \pkg{GNU MCSim} native code can have faster speed to obtain the model outputs.
+We first implemented one-compartment TK model in \proglang{R} by compiling the file written under the \pkg{GNU MCSim}'s model format.
+\pkg{pksensi} allows users select the preferred method to solve the pharmacokinetic model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
+Running model under \pkg{GNU MCSim} native code can have faster speed to obtain the model outputs.
+This section demo how to conduct global sensitivity analysis with \pkg{deSolve} package.
+Then compare the computer time with \pkg{MCSim}.
 
-The following \proglang{R} script can download and compile the model description file \code{pbtk1cpt.model} and use code{ode} function in \pkg{deSolve} package \citep{JSSv033i09} to solve ordinary differential equations in our model.
-The example model code of one-compartment PBTK model is available with \pkg{pksensi} package:
-
+The following \proglang{R} script will download the one-compartment PBTK model from https://github.com/nanhung/pksensi/blob/master/tests/1cpt.model:
 ```{r, eval=T}
 pbtk1cpt_model()
-cat(readLines("pbtk1cpt.model"), sep = "\n")
 ```
 
-Then, use \code{compile_model()} to generate the executable files (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
+After create the model file, use \code{compile_model()} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an R session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
 
@@
 parms <- initParms()
-parms["vdist"] <- 0.74
-parms["ke"] <- 0.28
+parms["vdist"] <- 6.13
+parms["ke"] <- 0.0228
 parms["kgutabs"] <- 2.18
@@
 ```
+
+Here shows the giveb parameter value (\code{parms}), initial state conditions (\code{initStat}), and the output variable (\code{Outputs}) that need to specify in model solving.
 
@@
 The \code{initParms} and \code{initStates} functions were used to customize the parameter values and the initial state that will be used in the \code{solve_fun()} function.
-These parameter value can be adopted from the \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
-In this case, we used the parameter value of theophylline in this example.
+These parameter value can be refered from the \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
+In this case, we used the parameter value of Bisphenol A in this example.
 The given \code{vdist}, \code{ke}, and \code{kgutabs} are 0.74, 0.28, and 2.18, respectively.
@@
 
-```{r, eval=T, fig.cap = 'Figure 1. Simulation results of one-compartment PBTK model.'}
+```{r, eval=T, fig.cap = 'Simulation results of one-compartment PBTK model.'}
 plot(y)
@@
 
-To conduct sensitivity analysis for the parameters in one-compartment pharmacokinetic model in this case, we want to quantify the impact of these three parameters on the chemical concentration in plasma during 24-hour time period post intake.
+To conduct sensitivity analysis for the parameters in one-compartment pharmacokinetic model in this case, we aim to quantify the impact of these three parameters on the chemical concentration in plasma during 24-hour time period post intake.
 We assume a uniform distribution for the estimate for each parameter with the coefficient of uncertainty within 50%.
@@
 ```{r, eval=T}
-y <- solve_fun(x, times, initState = initState,
+out <- solve_fun(x, times, initState = initState,
                outnames = Outputs, dllname = mName)
-tell2(x,y)
 ```
@@
 ```{r, eval=T, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24-hour time period intake.'}
-plot(x)
+plot(out)
 ```
@@
 par(mfrow = c(3,3), mar = c(2,2,2,2), oma = c(2,2,1,1))
-plot(x$a[,1,"vdist"], y[,1,"0.01",], main = "vdist")
+plot(x$a[,1,"vdist"], out$y[,1,"0.01",], main = "vdist")
 text(1, .7, "t=0.01",cex = 1.2)
-plot(x$a[,1,"ke"], y[,1,"0.01",], main = "ke")
-plot(x$a[,1,"kgutabs"], y[,1,"0.01",], main = "kgutabs")
-plot(x$a[,1,"vdist"], y[,1,"2.01",])
+plot(x$a[,1,"ke"], out$y[,1,"0.01",], main = "ke")
+plot(x$a[,1,"kgutabs"], out$y[,1,"0.01",], main = "kgutabs")
+plot(x$a[,1,"vdist"], out$y[,1,"2.01",])
 text(1, 18, "t=2.01",cex = 1.2)
-plot(x$a[,1,"ke"], y[,1,"2.01",])
-plot(x$a[,1,"kgutabs"], y[,1,"2.01",])
-plot(x$a[,1,"vdist"], y[,1,"24.01",])
+plot(x$a[,1,"ke"], out$y[,1,"2.01",])
+plot(x$a[,1,"kgutabs"], out$y[,1,"2.01",])
+plot(x$a[,1,"vdist"], out$y[,1,"24.01",])
 text(1, .7, "t=24.01",cex = 1.2)
-plot(x$a[,1,"ke"], y[,1,"24.01",])
-plot(x$a[,1,"kgutabs"], y[,1,"24.01",])
+plot(x$a[,1,"ke"], out$y[,1,"24.01",])
+plot(x$a[,1,"kgutabs"], out$y[,1,"24.01",])
 mtext("parameter", SOUTH<-1, line=0.4, outer=TRUE)
@@
 ```{r, eval=T}
-dim(y)
+dim(out$y)
 ```
@@
 ```{r, eval=T}
-check(x, SI.cutoff = 0.5)
+check(out, SI.cutoff = 0.5)
 ```
@@
 ```{r, eval=T}
-system.time(y<-solve_fun(x, times, initState = initState, 
+system.time(out<-solve_fun(x, times, initState = initState, 
                          outnames = Outputs, dllname = mName))
@@
 conditions <- c("Agutlument = 10") # Set the initial state of Agutlument = 10 
-system.time(y<-solve_mcsim(x, mName = mName, 
+system.time(out <- solve_mcsim(x, mName = mName, 
                            params = params,
@@
                 rtol = 1e-7, atol = 1e-9)
-system.time(y <- solve_mcsim(x, mName = mName, 
+system.time(out <- solve_mcsim(x, mName = mName, 
                              params = params,
@@
                              generate.infile = F))
-tell2(x,y)
 ```
@@
 ```{r, eval=T, fig.height=8, fig.width=8, fig.cap = ' '}
-plot(x, vars = "lnCPL_AG_mcgL")
+plot(out, vars = "lnCPL_AG_mcgL")
 ```
@@
 par(mfrow = c(2,2), mar = c(2,2,1,1), oma = c(2,2,1,1))
-pksim(y, vars = "lnCPL_APAP_mcgL")
+pksim(out$y, vars = "lnCPL_APAP_mcgL")
 text(1, 15, "APAP",cex = 1.2)
 points(APAP$Time, log(APAP$APAP * 1000))
-pksim(y, vars = "lnCPL_AG_mcgL", legend = F)
+pksim(out$y, vars = "lnCPL_AG_mcgL", legend = F)
 text(1, 15, "AG",cex = 1.2)
 points(APAP$Time, log(APAP$AG * 1000))
-pksim(y, vars = "lnCPL_AS_mcgL", legend = F)
+pksim(out$y, vars = "lnCPL_AS_mcgL", legend = F)
 text(1, 15, "AS",cex = 1.2)
@@
 ```{r, eval=T}
-check(x)
+check(out)
 ```
@@
 ```{r, eval=T}
-check(x, vars = "lnCPL_APAP_mcgL", SI.cutoff = 0.1, CI.cutoff = 0.1)
+check(out, vars = "lnCPL_APAP_mcgL", SI.cutoff = 0.1, CI.cutoff = 0.1)
 ```
@@
 ```{r, eval=T, fig.cap = ' '}
-heat_check(x, order = "interaction")
+heat_check(out, order = "interaction")
 ```
@@
 ```{r, eval=T, fig.cap = ' '}
-heat_check(x, order = "total order")
+heat_check(out, order = "total order")
 ```
@@
 ```{r, eval=T, fig.cap = ' '}
-heat_check(x, index = "CI", CI.cutoff = 0.05, order = "total order")
+heat_check(out, index = "CI", CI.cutoff = 0.05, order = "total order")
 ```

### 2019-06-23 manuscript/manuscript.Rmd
manuscript

@@
 bibliography: references.bib 
+geometry: margin=0.9in
 output: rticles::jss_article
@@
 
-Most sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} \citep{pianosi2015matlab}, \proglang{python} \citep{herman2017salib}, and \proglang{R} \citep{iooss2017introduction}.
+Most of sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} \citep{pianosi2015matlab}, \proglang{python} \citep{herman2017salib}, and \proglang{R} \citep{iooss2017introduction}.
 Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package \citep{R-sensitivity}, which is a practical tool to conduct local and global sensitivity analysis.
-The sensitivity package includes some functions to generate the parameter sequences by using different computing algorithms.
+The \pkg{sensitivity} package includes some functions to generate the parameter sequences by using different computing algorithms.
 Also, it provides a feasible way to integrate external modeling results.
@@
 \pkg{ODEsensitivity} can perform sensitivity analysis in ordinary differential equation (ODE) models \citep{R-ODEsensitivity}.
-It uses the ODE interface from the deSolve package, connecting it with the sensitivity analysis from the sensitivity package, and can perform sensitivity analysis on models with multivariate output.
+It uses the ODE interface from the \pkg{deSolve} package, connecting it with the sensitivity analysis from the \pkg{sensitivity} package, and can perform sensitivity analysis on models with multivariate output.
 Other related packages such as \pkg{mtk} \citep{RJ-2015-031} and \pkg{fast} \citep{R-fast} are designed to investigate the influence of model parameter.
 Although these tools provide various approaches to perform sensitivity analysis, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.
-Wepresent here an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
+We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK, and advanced compartment absorption and transit models with multivariate model output.
@@
 *	Freely available: All related packages are free and open source
-*	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \proglang{GNU MCSim} 
+*	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \pkg{GNU MCSim} 
 *	Decision making: The output results and visualization tools can be used to easily determine which parameters have "non-influential" effects on the model output and can be fixed in model calibration.
@@
 
-\pkg{pksensi} is available from the comprehensive R Archive Network at https://CRAN.R-project.org/package=pksensi.
+The R package \pkg{pksensi} is available from the comprehensive R Archive Network at https://CRAN.R-project.org/package=pksensi.
 The latest up-to-date version is available through GitHub at https://github.com/nanhung/pksensi.
+
+In the beginning, the user need to pre-install \pkg{GNU MCSim} \citep{JSSv002i09} and related standard C compiler to solve the ODE function.
+In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
+The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_install()} in \pkg{pksensi}.
+The C compiler had originated in Linux or MacOS.
+The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as,
+
+```{r}
+Rtools.path <- "c:/Rtools/bin; c:/Rtools/mingw_64/bin"
+Sys.setenv(PATH = paste(Rtools.path, Sys.getenv("PATH"), sep=";"))
+```
 
@@
 
-To investigate the convergence of sensitivity indices, we adopted the exam approach proposed by \citep{sarrazin2016global}.
+To investigate the convergence of sensitivity indices, we adopted the "exam" approach proposed by \citep{sarrazin2016global}.
 This method quantitatively assesses convergence by computing the width of 95% confidence intervals for all parameters across all time-points and output variables.
@@
 
-The \pkg{pksensi} provides a useful function to solve ordinary differential equations in pharmacokinetic models using analytical and numerical approaches.
+The R package \pkg{pksensi} provides a useful function to solve ordinary differential equations in pharmacokinetic models using analytical and numerical approaches.
 The solution can be performed under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package \citep{JSSv033i09}.
@@
 This compiled file can load and execute through the build-in function in \proglang{R}.
-However, Windows users need to install Rtools or MinGW to compile the source code by using the GNU GCC compiler.
+However, Windows users need to install Rtools to compile the source code by using the GNU GCC compiler.
 More requirements are detailed in the study from \citep{soetaertr}.
@@
 
-## Output visualization and decision making
+## Output visualization and decision support
 
@@
 This format is particularly suited for further graphical routines within \pkg{pksensi}.
-Also, a plot can can create time-dependent sensitivity measurements of first and total order effect for all parameters in one figure.
-Furthermore, \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
+Also, time-dependent sensitivity measurements of first and total order effect for all parameters can be plotted on one figure.The sensitivity index might have different estimation across replications under the same sample size.
+The inadequate sample number can create the unrobust result and bias.
+It's essential to use the adequate sample size to prevent the incorrect judgement in parameter fixing after senstivity analysis.
+Therefore, \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
 We also developed a "cut-off"-based approach in this package to accurately distinguish between "influential" and "non-influential" parameters.
-Finally, \pkg{pksensi} includes numerous visualization tools for the effective investigation and communication of results in decision making.
+Finally, \pkg{pksensi} includes numerous visualization tools for the effective investigation and communication of results in decision support.
 
 # Examples
-
-## One-compartment toxicokinetic model
 
@@
   collapse = TRUE,
-  comment = "#>",
+  comment = " ",
   fig.height=4, 
-  fig.width=6
+  fig.width=7
 )
@@
 
+Two example physicollogically based kinetic model were applied here to demonstrate how \pkg{pksensi} work.
+The following code shows the version of platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current demo.
+The version of \pkg{GNU MCSim} is 6.1.0.
+
+```{r}
+sessionInfo()$R.version$version.string
+```
+
+```{r}
+sessionInfo()$otherPkgs$pksensi$Version
+```
+
+## One-compartment model
+
 ### Equations
 
-In this example, we use a simple, one-compartment PBTK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to physiologically based kinetic studies.
+In the first example, we use a simple, one-compartment PBTK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to physiologically based kinetic studies.
 The differential equations for this model is as:
@@
 $$\frac{dA_{gutlumen}}{dt} = -k_{gutabs} \cdot A_{gutlumen} + g(t)$$
-$$\frac{dC_{rest}}{dt} = \frac{k_{gutabs}}{V_{dist}}-k_{elim} \cdot C_{rest}$$
+$$\frac{dA_{rest}}{dt} = \frac{dA_{gutlumen}}{dt}-k_{elim} \cdot A_{rest}$$
 
-where $A_{gutlumen}$ is the state variable that describes the quantity of compound in gut lumen (mol), $k_{gutabs}$ is the absorption rate constant that describes the chemical absorption from the gut lumen into gut tissue through first-order processes (/h), $V_{dist}$ is the volume of distribution (L), and $k_{elim}$ is the elimination rate constant (/h), which is equal to the total clearance divided by the volume of distribution.
+where $A_{gutlumen}$ is the state variable that describes the quantity of compound in gut lumen (mol) and $A_{rest}$ is the quantity of compound in rest of body and body blood (mol).
+The parameter $k_{gutabs}$ is the absorption rate constant that describes the chemical absorption from the gut lumen into gut tissue through first-order processes (/h) and $k_{elim}$ is the elimination rate constant (/h), which is equal to the total clearance divided by the volume of distribution.
 The time-dependent function $g(t)$ is used to describe the oral dosing schedule.
-$C_{rest}$ is the chemical concentration in plasma that can be used to compare with observed results in a pharmacokinetic experiment (mol/L).
+
+The concentration of chemical in the rest of body blood can be calculate as
+
+$$ C_{rest} = A_{rest} / V_{dist} \cdot BW$$
+
+In addition, $C_{rest}$ is the chemical concentration in plasma that can be used to compare with observed results in a pharmacokinetic experiment (mol/L).
+
+, $V_{dist}$ is the volume of distribution (L),
 
@@
 
-In the beginning, we need to pre-install \pkg{GNU MCSim} \citep{JSSv002i09} and related standard C compiler.
-The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_install()} in \pkg{pksensi}.
-The C compiler had installed in Linux or MacOS.
-The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler before using this function.
-
-We first implemented one-compartment TK model in \proglang{R} by compiling the file written under the \pkg{GNU MCSim}'s model format.
+To start, we implemented one-compartment TK model in \proglang{R} by compiling the file written under the \pkg{GNU MCSim}'s model format.
 \pkg{pksensi} allows users select the preferred method to solve the pharmacokinetic model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
@@
 
-The following \proglang{R} script will download the one-compartment PBTK model from https://github.com/nanhung/pksensi/blob/master/tests/1cpt.model:
+The following \proglang{R} script will download the one-compartment PBTK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.model.
+The model mainly includes two state variable that are the quantity of compound in gut lumen (\code{Agutlument}) and plasma (\code{Acompartmant}).
+The
+
 ```{r, eval=T}
 pbtk1cpt_model()
+cat(readLines("pbtk1cpt.model"), sep = "\n")
 ```
@@
 
-```{r, eval=T}
+```{r}
 mName <- "pbtk1cpt"
-compile_model(mName, application = "R")
+```
+
+```{r, eval=T, message=F, warning=F}
+compile_model(mName, application = "R", version = "6.1.0")
+```
+
+```{r}
 source(paste0(mName, "_inits.R"))
@@
 The parameter values and initial states can be customized to specify the properties and schedule for the given dosing scenario.
+In the current setting, we assumed the initial condition of the intake chemical to be 10 mol.
+We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the the following modeling and sensitivity analysis.
+These additional functions are generated when compiling the model file.
+
+In this case, we used the corresponding parameter value of theophylline in this example.
+The parameter value can  be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
+The given value of \code{vdist}, \code{ke}, and \code{kgutabs} in \pkg{httk} are 0.78 (L/kg BW), 0.265 (/h), and 2.18 (/h), respectively.
+The body weight is assumed to be 60 (kg).
 
@@
 parms <- initParms()
-parms["vdist"] <- 6.13
-parms["ke"] <- 0.0228
+parms["vdist"] <- 0.78
+parms["ke"] <- 0.265
 parms["kgutabs"] <- 2.18
+parms["BW"] <- 60
 initState <- initStates(parms=parms)
@@
 
-Here shows the giveb parameter value (\code{parms}), initial state conditions (\code{initStat}), and the output variable (\code{Outputs}) that need to specify in model solving.
+Here shows the given parameter value (\code{parms}), initial state condition (\code{initStat}), and the output variable (\code{Outputs}) that need to specify in model solving.
 
@@
 
-In the current setting, we assumed the initial condition of the intake chemical to be 10 mol.
-The \code{initParms} and \code{initStates} functions were used to customize the parameter values and the initial state that will be used in the \code{solve_fun()} function.
-These parameter value can be refered from the \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
-In this case, we used the parameter value of Bisphenol A in this example.
-The given \code{vdist}, \code{ke}, and \code{kgutabs} are 0.74, 0.28, and 2.18, respectively.
-
-Through \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic according to the given parameter conditions such as time points (\code{times}):
+Using the \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic profile (Figure \ref{fig:ode-PK}) according to the given parameter conditions and the time points (\code{t}):
 
 ```{r, eval=T}
-library(deSolve)
-times <- seq(from = 0.01, to = 24.01, by = 1)
-y <- ode(initState, times, func = "derivs", parms = parms, 
-         dllname = mName, initfunc = "initmod", nout = 1, outnames = Outputs)
+t <- seq(from = 0.01, to = 24.01, by = 1)
+y <- deSolve::ode(initState, t, func = "derivs", parms = parms, 
+         dllname = mName, initfunc = "initmod", 
+         nout = 1, outnames = Outputs)
 ```
 
-```{r, eval=T, fig.cap = 'Simulation results of one-compartment PBTK model.'}
+```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PBTK model.'}
 plot(y)
@@
 
-To conduct sensitivity analysis for the parameters in one-compartment pharmacokinetic model in this case, we aim to quantify the impact of these three parameters on the chemical concentration in plasma during 24-hour time period post intake.
-We assume a uniform distribution for the estimate for each parameter with the coefficient of uncertainty within 50%.
-The parameter ranges are assumed to be (0.37, 1.12) for \code{vdist}, (0.0058, 0.0174) for \code{ke}, and (0.045, 0.136) for \code{kgutabs}.
-The sample number determines the robustness of the result of sensitivity analysis.
-Higher sample numbers can generate narrow confidence intervals for sensitivity measurements across different replications.
-However, they might cause heavy computational burden for complex models.
-Here we use a sample number of 400 with 20 replications:
+To conduct a sensitivity analysis for the four parameters in one-compartment TK model, we quantify their impact on the chemical concentration in plasma during 24-hour time period post intake.
+The distribution of parameter is taken to be uniform with bounds corresponding to 50% and 200% of the nominal value.
+Therefore, the parameter ranges are assumed to be 0.39 and 1.56 L/kg BW for \code{vdist}.
+The \code{ke} are ranged 0.133 to 0.53 /h, corresponding to half-times of 1.3 and 5.2 hr.
+The \code{ka} are ranged 1.09 to 4.36 /h, corresponding to half-times of 0.32 and 0.64 hr.
+The Body weight are assumed to normal distribution with mean = 60 kg and sd = 5 kg.
+The number of samples determines the robustness of the result of sensitivity analysis.
+Higher sample numbers sizes lead to narrower confidence intervals for sensitivity measurements across different replications.
+However, it will take longer time in computation.
+Here we use a sample size of 400 with 20 replications:
 
 ```{r, eval=T}
-LL <- 0.5 
-UL <- 1.5
-q <- "qunif"
-q.arg <- list(list(min = parms["vdist"] * LL, max = parms["vdist"] * UL),
-             list(min = parms["ke"] * LL, max = parms["ke"] * UL), 
-             list(min = parms["kgutabs"] * LL, max = parms["kgutabs"] * UL)) 
+q <- c("qunif", "qunif", "qunif", "qnorm")
+q.arg <- list(list(min = parms["vdist"] / 2, max = parms["vdist"] * 2),
+             list(min = parms["ke"] / 2, max = parms["ke"] * 2), 
+             list(min = parms["kgutabs"] / 2, max = parms["kgutabs"] * 2),
+             list(mean = parms["BW"], sd = 5)) 
+params <- c("vdist", "ke", "kgutabs", "BW")
 set.seed(1234)
-params <- c("vdist", "ke", "kgutabs")
 x <- rfast99(params, n = 200, q = q, q.arg = q.arg, replicate = 20)
@@
 
-Because the pharmacokinetic model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements can also show the time-dependent relationships for each model parameter.
-Here we define the output time points to examine the change of the parameter sensitivity over time.
-To solve the pharmacokinetic model through deSolve, we need to provide the details of the argument:
+Through \code{rfast99} function, S3 object for class 'rfast99' will be created.
+The generated parameter will be stored as a 3-D array under the named \code{a} with the dimension of sample number, replication, and parameter number.
+
+```{r}
+dim(x$a)
+```
+
+The sample number was 800 with 4 model parameters, which generated 3,200 model evaluations. The replication was set to 20. Figure \ref{fig:rep-sampl} only plotted for 3 replications.
+
+```{r, eval=T, fig.cap = '\\label{fig:rep-sampl}The parameter search curve generated by eFAST method.
+The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.'}
+par(mfrow=c(4,4),mar=c(0,0,0,0),oma=c(3,3,2,1))
+cex <- 0.1
+for (j in c("vdist", "ke", "kgutabs", "BW")) {
+  if ( j == "BW") {
+    plot(x$a[,1,j], ylab = "", cex = cex)
+  } else plot(x$a[,1,j], xaxt="n", ylab = "", cex = cex)
+  for (i in 2:3) {
+  if ( j == "BW") {
+    plot(x$a[,i,j], ylab = "", yaxt="n", cex = cex)  
+    } else plot(x$a[,i,j], xaxt="n", yaxt="n", ylab = "", cex = cex)
+  } 
+  hist <- hist(x$a[,,j], plot=FALSE, 
+               breaks=seq(from=min(x$a[,,j]), to=max(x$a[,,j]), length.out=20))
+  barplot(hist$density, axes=FALSE, space=0, horiz = T) 
+}
+```
+
+Because the physiologically based kinetic model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements therefore have the time-dependent relationships for each model parameter.
+Here we define the output time points (\code{t}) to examine the change of the parameter sensitivity over time.
+To solve the model through \pkg{deSolve}, we need to provide the details of the argument:
 
 ```{r, eval=T}
-out <- solve_fun(x, times, initState = initState,
-               outnames = Outputs, dllname = mName)
+outputs <- c("Ccompartment", "Ametabolized")
+out <- solve_fun(x, time = t, initState = initState, 
+                 outnames = outputs, dllname = mName)
 ```
 
-To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours in the example.
-The \code{initParmsfun} is used to generate the sampling value for each parameter.
-The \code{outnames}, \code{dllname}, \code{func}, \code{initfunc} are based on the arguments from the ode function in \pkg{deSolve} package.
+To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours as above definition in \code{ode} function in this example.
+The \code{outnames}, \code{dllname}, are based on the arguments from the \code{ode} function in \pkg{deSolve} package.
 The details of the model structure and these arguments are defined in \code{pbtk1comp.c}.
 and \code{pbtk1comp_inits.R}.
-Finally, the \code{tell2()} function is used to integrate the parameter values and the output results of numerical analysis that were generated and stored in variables x and y.
 The result of object x is an object of rfast99, which has specific \code{print()}, \code{plot()}, and \code{check()} method.
@@
 In addition to print out the result of sensitivity analysis, the more efficient way to distinguish the influence of model parameter is to visualize them.
-The time-dependent sensitivity indices are shown in Figure 2.
+The time-dependent sensitivity indices are shown in Figure \ref{fig:pk-solvefun}.
 
-```{r, eval=T, fig.cap = 'Figure 2. Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24-hour time period intake.'}
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24-hour time period intake.
+The solid line represent the total (black) and first (red) order sensitivity index with 95\\% confidence interval (polygon).
+The dash line is the cut-off with default value of 0.05'}
 plot(out)
@@
 
-Here, we can find that \code{vdist} and \code{ke} are dominating the plasma concentration in the before and after 5-hour post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
+The
+
+```{r, eval=T}
+plot(out, vars = "Ametabolized")
+```
+
+Here, we can see (Figure \ref{fig:pk-solvefun}) that \code{vdist} and \code{ke} are dominating the plasma concentration in the before and after 5-hour post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
 Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-The relationship between concentration and the parameters can be plotted as follow (Figure 3):
+The relationship between concentration and the parameters can be plotted as follow:
 
-```{r, eval=T, fig.cap = 'Figure 3. The relationship between model parameter and estimated concentration under the time-point of 0.01, 2.01, and 24.01 hr'}
+```{r, eval=T, fig.cap = 'The relationship between model parameter and estimated concentration at times 0.01, 2.01, and 24.01 hr'}
 par(mfrow = c(3,3), mar = c(2,2,2,2), oma = c(2,2,1,1))
-plot(x$a[,1,"vdist"], out$y[,1,"0.01",], main = "vdist")
+plot(x$a[,1,"vdist"], out$y[,1,"0.01",1], main = "vdist")
 text(1, .7, "t=0.01",cex = 1.2)
-plot(x$a[,1,"ke"], out$y[,1,"0.01",], main = "ke")
-plot(x$a[,1,"kgutabs"], out$y[,1,"0.01",], main = "kgutabs")
-plot(x$a[,1,"vdist"], out$y[,1,"2.01",])
+plot(x$a[,1,"ke"], out$y[,1,"0.01",1], main = "ke")
+plot(x$a[,1,"kgutabs"], out$y[,1,"0.01",1], main = "kgutabs")
+plot(x$a[,1,"vdist"], out$y[,1,"2.01",1])
 text(1, 18, "t=2.01",cex = 1.2)
-plot(x$a[,1,"ke"], out$y[,1,"2.01",])
-plot(x$a[,1,"kgutabs"], out$y[,1,"2.01",])
-plot(x$a[,1,"vdist"], out$y[,1,"24.01",])
+plot(x$a[,1,"ke"], out$y[,1,"2.01",1])
+plot(x$a[,1,"kgutabs"], out$y[,1,"2.01",1])
+plot(x$a[,1,"vdist"], out$y[,1,"24.01",1])
 text(1, .7, "t=24.01",cex = 1.2)
-plot(x$a[,1,"ke"], out$y[,1,"24.01",])
-plot(x$a[,1,"kgutabs"], out$y[,1,"24.01",])
+plot(x$a[,1,"ke"], out$y[,1,"24.01",1])
+plot(x$a[,1,"kgutabs"], out$y[,1,"24.01",1])
 mtext("parameter", SOUTH<-1, line=0.4, outer=TRUE)
@@
 
-In addition, the output are also formated with c(model evaluation, replication, time, variable).
+In addition, the output are also formated as 4-D array under \code{c(model evaluation, replication, time, variable)}.
 
@@
 ```{r, eval=T}
-check(out, SI.cutoff = 0.5)
+check(out, SI.cutoff = 0.05)
 ```
@@
 
-### Model implementations with GNU MCSim
+### Model implementation with GNU MCSim
 
-Alternatively, to solve ODE by using \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model()}.
-Rather than apply \pkg{deSolve} to solve differential equations, the \pkg{GNU MCSim} can provide higher computational speed in global sensitivity analysis.
+In addtion to apply \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can provide better computational efficiency in global sensitivity analysis.
+To solve ODE by using \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model()}.
 
 ```{r, eval=T}
-system.time(out<-solve_fun(x, times, initState = initState, 
+system.time(out<-solve_fun(x, t, initState = initState, 
                            outnames = Outputs, dllname = mName))
@@
 ```{r, eval=T}
-compile_model(mName, application = "mcsim")
+compile_model(mName, application = "mcsim", version = "6.1.0")
 ```
 
-Similar to \code{solve_fun()} function that can define the initial parameter and state values through input function, the \code{solve_mcsim()} has a `condition` argument that is used to givien the specific input value such as oral dose or fixing parameter value or initial state variable.
+Similar to \code{solve_fun()} function that can define the initial parameter and state values through input function, the \code{solve_mcsim()} has a `condition` argument that is used to give the specific input value such as oral dose, fixing parameter value or initial state variable.
 
@@
 conditions <- c("Agutlument = 10") # Set the initial state of Agutlument = 10 
-system.time(out <- solve_mcsim(x, mName = mName, 
-                           params = params,
-                           vars = Outputs,
-                           time = times,
+system.time(out <- solve_mcsim(x, mName = mName, params = params, 
+                               vars = Outputs, time = t, 
                                condition = conditions))
@@
 
-Under the same given condition, it takes 5-6 seconds (\pkg{deSolve}) and 1-2 seconds (\pkg{GNU MCSim})  to solve model.
-The \code{solve_mcsim()} shows the better computational performance than \code{solve_fun()} in \pkg{pksensi}.
+Under the same given condition, it takes about 3 times computer time in (\pkg{deSolve}) than in (\pkg{GNU MCSim}) to solve this model.
+
+---
+title: "R Notebook"
+output: html_notebook
+---
 
@@
 
-Generally, The wide range of parameter value might cause the computational error in the solver.
+Generally, wide ranges for parameter values might cause the computational error in the solver.
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
 The \code{generate_infile()} provide the arguments of \code{rtol} and \code{atol} that can be adjusted to prevent the unwanted error.
-However, the modification will slow down the computational speed.
-Therefore, the alternative method to prevent the computational error is to detect the crucial parameter range that causes the problem.
+However, the modification will lower computational speed.
+Therefore, the alternative method to prevent the computational errors is to detect the crucial parameter range that causes the problem.
 Also, setting the maximum number of (internally defined) steps to higher value instead of using the default value (500) can prevent this problem.
@@
 
-To improve the computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
+To improve computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
 The setting oral dose of APAP is 20 mg/kg in this example.
@@
 Since the unit of the given dose is mg/kg, the \code{mgkg_flag} is set to 1 to declare the statement.
-More definition of input can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
+More definition of input functions can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
 
@@
 FDA.
-This abstract reflects the views of the author and should not be construed to represent FDA’s views or policies.
+This article reflects the views of the author and should not be construed to represent FDA’s views or policies.
 We thank Dr.
 Yasha Hartberg and Dr.
-Barbara Gastel at the Texas A&M University for reviewing the manuscript and consultation.
+Barbara Gastel at the Texas A&M University and Dr.
+Frederic Bois from Certara for reviewing the manuscript and consultation.
 

### 2019-06-23 manuscript/manuscript.Rmd
manuscript

@@
 
-In our previous work \citep{fphar201800588}, we developed an approach to apply global sensitivity analysis workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+In our previous work \citep{fphar201800588}, we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
 We used GNU MCSim \citep{JSSv002i09}, an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
-We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of global sensitivity analysis algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
+We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
+It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as Jansen or OWen method.
 Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
@@
 Although these tools provide various approaches to perform sensitivity analysis, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.
+
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
@@
 In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
-The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_install()} in \pkg{pksensi}.
+The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_instal} in \pkg{pksensi}.
 The C compiler had originated in Linux or MacOS.
@@
 
-We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package \citep{saltelli1999quantitative}.
+We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package \citep{saltelli1999quantitative}.
 Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package \citep{R-sensitivity}.
-To test the convergence and robustness of the sensitivity measurement, we included a random phase-shift approach to replicate sampling from random starting points across parameter space.
+The eFAST method compute the sensitivity index (SI) via ANOVA-like decomposition of the function for analysis.
+This type of GSA approach and SI also know as Sobol method and Sobol index.
+Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
 This can be written as:
@@
 To investigate the convergence of sensitivity indices, we adopted the "exam" approach proposed by \citep{sarrazin2016global}.
-This method quantitatively assesses convergence by computing the width of 95% confidence intervals for all parameters across all time-points and output variables.
+This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed sensitivity index for all parameters across all time-points and output variables.
+Through random phase shift procedure, the sensitivity index can repeatly generate with low and high variation based on the given sampling size.
+Therefore, they can be used to investigate the convergence of sensitivity indices as
+
+$$CI_{i,t} = max(SI_{i,t}^{ub}-SI_{i,t}^{lb})$$
+
+where $SI_{i,t}^{ub}$ and $SI_{i,t}^{ub}$ are the upper and lower bound of the sensitivity index from the specific model output ($i$) at time ($t$). $CI_{i,t}$ is the corresponding convergence index.
+
 To integrate the random phase-shift with eFAST, we built the \code{rfast99} function, which can be used to perform the sensitivity test for the whole pharmacokinetic modeling process.
@@
 The number of model evaluations is equal to the sample size times the number of model parameters.
+
+## Monte Carlo
+
+The uncertainty analysis is a crucial modeling process with sensitivity analysis.
+Sometimes, it's difficult to have informative information to set up the parameter distribution.
+The wider range of parameter distribution might cause the computation error at the extreme value.
+The narrower range might cause the incorrect calibrated result of the model parameter.
+To prevent this problem.
+The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process.
+Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
 
@@
 
-In the first example, we use a simple, one-compartment PBTK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to physiologically based kinetic studies.
+In the first example, we use a simple, one-compartment PK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to physiologically based kinetic studies.
 The differential equations for this model is as:
@@
 
-The concentration of chemical in the rest of body blood can be calculate as
+The concentration of chemical in the rest of body blood $C_{rest}$ can be calculate as
 
@@
 
-In addition, $C_{rest}$ is the chemical concentration in plasma that can be used to compare with observed results in a pharmacokinetic experiment (mol/L).
-
-, $V_{dist}$ is the volume of distribution (L),
+where $V_{dist}$ is the volume of distribution (L/Kg BW) and $BW$ is the body weight (Kg).
+The $C_{rest}$ can also be seen as the chemical concentration in plasma that can be used to compare with observed results in a pharmacokinetic experiment (mol/L).
 
@@
 
-To start, we implemented one-compartment TK model in \proglang{R} by compiling the file written under the \pkg{GNU MCSim}'s model format.
+To start, we implemented the one-compartment PK model in \proglang{R} by compiling the file written under the \pkg{GNU MCSim}'s model format.
 \pkg{pksensi} allows users select the preferred method to solve the pharmacokinetic model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
@@
 
-The following \proglang{R} script will download the one-compartment PBTK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.model.
-The model mainly includes two state variable that are the quantity of compound in gut lumen (\code{Agutlument}) and plasma (\code{Acompartmant}).
-The
+The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.model.
+The model mainly includes two state variable that are the quantity of compound in gut lumen (\code{Agutlument}) and rest of body (\code{Acompartmant}).
+The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.
+The \code{AUC} is the calcualated area under the curve of rest body ($mol \cdot h/L$).
 
@@
 
-After create the model file, use \code{compile_model()} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an R session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
+After create the model file, use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an R session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
 
@@
 
-```{r, eval=T, message=F, warning=F}
+```{r eval=T, message=FALSE, warning=FALSE}
 compile_model(mName, application = "R", version = "6.1.0")
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PBTK model.'}
+```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model.'}
 plot(y)
@@
 
-To conduct a sensitivity analysis for the four parameters in one-compartment TK model, we quantify their impact on the chemical concentration in plasma during 24-hour time period post intake.
+To conduct a sensitivity analysis for the four parameters in one-compartment PK model, we quantify their impact on the chemical concentration in plasma during 24-hr time period post intake.
 The distribution of parameter is taken to be uniform with bounds corresponding to 50% and 200% of the nominal value.
@@
 Through \code{rfast99} function, S3 object for class 'rfast99' will be created.
-The generated parameter will be stored as a 3-D array under the named \code{a} with the dimension of sample number, replication, and parameter number.
+The generated parameters are stored as a 3-D array under the named \code{a} with the dimension of sample number, replication, and parameter number, respectively.
 
@@
 
-The sample number was 800 with 4 model parameters, which generated 3,200 model evaluations. The replication was set to 20. Figure \ref{fig:rep-sampl} only plotted for 3 replications.
+The sample number was 800 with 4 model parameters, which generate 3,200 model evaluations. The replication was set to 20. Figure \ref{fig:rep-sampl} only plotted for 3 replications.
 
@@
 The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.'}
-par(mfrow=c(4,4),mar=c(0,0,0,0),oma=c(3,3,2,1))
-cex <- 0.1
+par(mfrow=c(4,4),mar=c(0,0,0,0),oma=c(3,3,2,1), pch =".")
 for (j in c("vdist", "ke", "kgutabs", "BW")) {
   if ( j == "BW") {
-    plot(x$a[,1,j], ylab = "", cex = cex)
-  } else plot(x$a[,1,j], xaxt="n", ylab = "", cex = cex)
+    plot(x$a[,1,j], ylab = "")
+  } else plot(x$a[,1,j], xaxt="n", ylab = "")
   for (i in 2:3) {
   if ( j == "BW") {
-    plot(x$a[,i,j], ylab = "", yaxt="n", cex = cex)  
-    } else plot(x$a[,i,j], xaxt="n", yaxt="n", ylab = "", cex = cex)
+    plot(x$a[,i,j], ylab = "", yaxt="n")  
+    } else plot(x$a[,i,j], xaxt="n", yaxt="n", ylab = "")
   } 
@@
 
-To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours as above definition in \code{ode} function in this example.
+To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours with time segment of 1 hour as above definition in \code{ode} function in this example.
+The initial time point should avoid 0 to prevent the computational error in misconduction.
 The \code{outnames}, \code{dllname}, are based on the arguments from the \code{ode} function in \pkg{deSolve} package.
 The details of the model structure and these arguments are defined in \code{pbtk1comp.c}.
-and \code{pbtk1comp_inits.R}.
-The result of object x is an object of rfast99, which has specific \code{print()}, \code{plot()}, and \code{check()} method.
-The print function gives the sensitivity and convergence indices for main, interaction, and total order at each time point.
+and \code{pbtk1comp_inits.R} files.
+The \code{out} is an S3 object of \code{rfast99} as well, which can link with \code{print}, \code{plot}, and \code{check} method to exame the sensitivity measurments.
+The \code{print} function gives the sensitivity and convergence indices for main, interaction, and total order at each time point.
 In addition to print out the result of sensitivity analysis, the more efficient way to distinguish the influence of model parameter is to visualize them.
-The time-dependent sensitivity indices are shown in Figure \ref{fig:pk-solvefun}.
+The time-dependent sensitivity indices are shown in Figure \ref{fig:pk-solvefun-rest}.
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PBTK model during 24-hour time period intake.
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-rest}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.
 The solid line represent the total (black) and first (red) order sensitivity index with 95\\% confidence interval (polygon).
@@
 
-The
+Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} are dominating the plasma concentration in the before and after 5-hr post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
+Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
+The current result is only based on the seting distribution of model parameters for theophylline.
+The different given conditions will change the result (result not shown).
 
-```{r, eval=T}
+The default output in the plotting is setting at the first variable.
+To exam the time-sependent sensitivity index of other variables, such as \code{Ametabolized} in this case, we need to assign the variable name \code{vars = "Ametabolized"} in \code{plot} function (Figure \ref{fig:pk-solvefun-Amet}).
+
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.'}
 plot(out, vars = "Ametabolized")
@@
 
-Here, we can see (Figure \ref{fig:pk-solvefun}) that \code{vdist} and \code{ke} are dominating the plasma concentration in the before and after 5-hour post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
-Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-The relationship between concentration and the parameters can be plotted as follow:
+The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is mainly determined by \code{ke}.
+Same as \code{Ccompartment}, the \code{kgutabs} contribute abour 30 - 40% variation of model output in the first hour.
 
-```{r, eval=T, fig.cap = 'The relationship between model parameter and estimated concentration at times 0.01, 2.01, and 24.01 hr'}
-par(mfrow = c(3,3), mar = c(2,2,2,2), oma = c(2,2,1,1))
-plot(x$a[,1,"vdist"], out$y[,1,"0.01",1], main = "vdist")
-text(1, .7, "t=0.01",cex = 1.2)
-plot(x$a[,1,"ke"], out$y[,1,"0.01",1], main = "ke")
-plot(x$a[,1,"kgutabs"], out$y[,1,"0.01",1], main = "kgutabs")
-plot(x$a[,1,"vdist"], out$y[,1,"2.01",1])
-text(1, 18, "t=2.01",cex = 1.2)
-plot(x$a[,1,"ke"], out$y[,1,"2.01",1])
-plot(x$a[,1,"kgutabs"], out$y[,1,"2.01",1])
+In addtion to create the time-sI profile to investigate the parameter impact on model output, the relationship between parameter and corresponding value of model output can be used to perform examination.
+Figure \ref{fig:SI-output} show the concentration of rest body and the model parameters:
+
+```{r, eval=T, fig.cap = '\\label{fig:SI-output}The relationship between model parameter and estimated concentration at times 0.01, 2.01, 6.01, and 24.01 hr'}
+par(mfcol=c(4,4),mar=c(0.5,0.5,0,0),oma=c(3,3,2,1), pch = ".")
+plot(x$a[,1,"vdist"], out$y[,1,"0.01",1], xaxt="n", main = "\nvdist")
+plot(x$a[,1,"vdist"], out$y[,1,"2.01",1], xaxt="n")
+plot(x$a[,1,"vdist"], out$y[,1,"6.01",1], xaxt="n")
 plot(x$a[,1,"vdist"], out$y[,1,"24.01",1])
-text(1, .7, "t=24.01",cex = 1.2)
-plot(x$a[,1,"ke"], out$y[,1,"24.01",1])
-plot(x$a[,1,"kgutabs"], out$y[,1,"24.01",1])
-mtext("parameter", SOUTH<-1, line=0.4, outer=TRUE)
-mtext("Ccompartment", WEST<-2, line=0.4, outer=TRUE)
+for (j in c("ke", "kgutabs", "BW")){
+  for (k in c("0.01", "2.01", "6.01", "24.01")){
+    if (k == "0.01") {
+      plot(x$a[,1,j], out$y[,1,k,1], yaxt = "n", xaxt="n", main = paste0("\n", j))
+    } else if (k == "24.01") {
+      plot(x$a[,1,j], out$y[,1,k,1], yaxt = "n")
+      } else plot(x$a[,1,j], out$y[,1,k,1], xaxt = "n", yaxt = "n")
+  }
+}
+mtext("Parameter", SOUTH<-1, line=1.8, outer=TRUE)
+mtext("Ccompartment", WEST<-2, line=1.8, outer=TRUE)
 ```
 
-The \code{x} is a list of class \code{rfast99}, containing all the input arguments detailed before and the calculated sensitivity indices of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
+We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameter with significant relationship with model output have hight impact to dominate the modeling result.
+The \code{ke} shows the high correlation at the lating time (t = 24.01 h).
+The \code{ke} doesn't show the obvious relationship with \code{Ccompartment}.
+
+The output variable \code{out} containing all the input arguments detailed before and the calculated sensitivity indices of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
 The convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
-The parameter values are stored in an array \code{x$a} with c(model evaluation, replication, parameters).
-
-```{r, eval=T}
-dim(x$a)
-```
-
-In addition, the output are also formated as 4-D array under \code{c(model evaluation, replication, time, variable)}.
+The output are formated as 4-D array in \code{y} with the dimension name of model evaluation, replication, time, and variable, respectively.
 
@@
 
-The \code{check()} is a useful function to determine which parameters have relative lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
-The argument of \code{SI.cutoff} is setting at 0.5 to detect the relative non-influential parameters in this case.
+The \code{check} is a useful function to determine which parameters have relative lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
+The argument of \code{SI.cutoff} is setting at 0.05 to detect the relative non-influential parameters as default.
 
@@
 
-Based on the sensitivity measurement of the total order, the result shows that \code{kgutabs} has relative lower measurement of sensitivity index.
+Based on the sensitivity measurement of the total order, the result shows that \code{BW} has a relative lower measurement of sensitivity index.
+However, all parameters do not converge to the setting cut-off, which means the larger sample size is required in further sensitivity testing.
+Same as \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use this assignment to exam the given output.
 
@@
 
-In addtion to apply \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can provide better computational efficiency in global sensitivity analysis.
-To solve ODE by using \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model()}.
+In addtion to apply \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can provide better computational efficiency.
+To solve ODE through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
+The computing time of using \code{solve_fun} in sensitivity analysis is estimated as,
 
@@
 
-```{r, eval=T}
+```{r, eval=T, message=F, warning=F}
 compile_model(mName, application = "mcsim", version = "6.1.0")
@@
 
-Similar to \code{solve_fun()} function that can define the initial parameter and state values through input function, the \code{solve_mcsim()} has a `condition` argument that is used to give the specific input value such as oral dose, fixing parameter value or initial state variable.
+Similar to \code{solve_fun} function that can define the initial parameter and state values through input function, the \code{solve_mcsim} has a \code{condition} argument that is used to give the specific input value such as oral dose, fixing parameter value or initial state variable.
 
 ```{r, eval=T}
-conditions <- c("Agutlument = 10") # Set the initial state of Agutlument = 10 
+conditions <- c("Agutlument = 10") 
 system.time(out <- solve_mcsim(x, mName = mName, params = params, 
@@
 
-Under the same given condition, it takes about 3 times computer time in (\pkg{deSolve}) than in (\pkg{GNU MCSim}) to solve this model.
-
----
-title: "R Notebook"
-output: html_notebook
----
+After soving the equations under the same given condition, we can find that (\pkg{GNU MCSim}) has about 3 times fatser in computing performance than using \pkg{deSolve}.
 

### 2019-06-24 manuscript/manuscript.Rmd
APAP

@@
 The aim of this section is to reproduce our previous published result \citep{fphar201800588} of global sensitivity analysis for acetaminophen PBPK model through \pkg{pksensi}.
-The model codes are included in this package and can be generated through \code{pbpk_apap_model()}.
+The model codes are included in this package and can be generated through \code{pbpk_apap_model} function.
 We applied the global sensitivity analysis workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
@@
 pbpk_apap_model()
-compile_model(mName, application = "mcsim")
+compile_model(mName, application = "mcsim", version = "6.1.0")
 ```

### 2019-06-24 manuscript/manuscript.Rmd
manuscript

@@
 
-Two example physicollogically based kinetic model were applied here to demonstrate how \pkg{pksensi} work.
-The following code shows the version of platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current demo.
+Two examples of physiologically based kinetic model were applied here to demonstrate how \pkg{pksensi} work.
+The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current demo.
 The version of \pkg{GNU MCSim} is 6.1.0.
@@
 
-where $A_{gutlumen}$ is the state variable that describes the quantity of compound in gut lumen (mol) and $A_{rest}$ is the quantity of compound in rest of body and body blood (mol).
+where $A_{gutlumen}$ is the state variable that describes the quantity of compound in the gut lumen (mol) and $A_{rest}$ is the quantity of compound in rest of body and body blood (mol).
 The parameter $k_{gutabs}$ is the absorption rate constant that describes the chemical absorption from the gut lumen into gut tissue through first-order processes (/h) and $k_{elim}$ is the elimination rate constant (/h), which is equal to the total clearance divided by the volume of distribution.
@@
 
-The concentration of chemical in the rest of body blood $C_{rest}$ can be calculate as
+The concentration of the chemical in the rest of body blood $C_{rest}$ can be calculated as
 
@@
 \pkg{pksensi} allows users select the preferred method to solve the pharmacokinetic model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
-Running model under \pkg{GNU MCSim} native code can have faster speed to obtain the model outputs.
+Running model under \pkg{GNU MCSim} native code can have a faster speed to obtain the model outputs.
 This section demo how to conduct global sensitivity analysis with \pkg{deSolve} package.
@@
 The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.model.
-The model mainly includes two state variable that are the quantity of compound in gut lumen (\code{Agutlument}) and rest of body (\code{Acompartmant}).
+The model mainly includes two state variables that are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
 The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.
-The \code{AUC} is the calcualated area under the curve of rest body ($mol \cdot h/L$).
+The \code{AUC} is the calculated area under the curve of the rest body ($mol \cdot h/L$).
 
@@
 
-After create the model file, use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an R session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
+After create the model file, use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
 
-```{r}
+```{r eval=T, message=FALSE, warning=FALSE}
 mName <- "pbtk1cpt"
+compile_model(mName, application = "R", version = "6.1.0")
 ```
 
-```{r eval=T, message=FALSE, warning=FALSE}
-compile_model(mName, application = "R", version = "6.1.0")
-```
+The \code{pbtk1cpt_inits.R} file includes \code{initParms} and \code{initStates} functions and \code{Outputs} variable.
+The created function have default value of model parameters and initial conditions that can further use to customize in the simulation.
 
@@
 
-The parameter values and initial states can be customized to specify the properties and schedule for the given dosing scenario.
+The parameter values and initial states can be customized to the specific condition.
+It can also schedule for the given dosing scenario.
 In the current setting, we assumed the initial condition of the intake chemical to be 10 mol.
-We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the the following modeling and sensitivity analysis.
+We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the following modeling and sensitivity analysis.
 These additional functions are generated when compiling the model file.
 
-In this case, we used the corresponding parameter value of theophylline in this example.
+We used the corresponding parameter value of theophylline in this example.
 The parameter value can  be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
@@
 
-To conduct a sensitivity analysis for the four parameters in one-compartment PK model, we quantify their impact on the chemical concentration in plasma during 24-hr time period post intake.
+To conduct a sensitivity analysis for the four parameters in one-compartment PK model, we quantify their impact on the chemical concentration in plasma during the 24-hr time period post intake.
 The distribution of parameter is taken to be uniform with bounds corresponding to 50% and 200% of the nominal value.
 Therefore, the parameter ranges are assumed to be 0.39 and 1.56 L/kg BW for \code{vdist}.
-The \code{ke} are ranged 0.133 to 0.53 /h, corresponding to half-times of 1.3 and 5.2 hr.
+The \code{ke} are ranged from 0.133 to 0.53 /h, corresponding to half-times of 1.3 and 5.2 hr.
 The \code{ka} are ranged 1.09 to 4.36 /h, corresponding to half-times of 0.32 and 0.64 hr.
-The Body weight are assumed to normal distribution with mean = 60 kg and sd = 5 kg.
+The Bodyweight is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
 The number of samples determines the robustness of the result of sensitivity analysis.
 Higher sample numbers sizes lead to narrower confidence intervals for sensitivity measurements across different replications.
-However, it will take longer time in computation.
+However, it will take a longer time in computation.
 Here we use a sample size of 400 with 20 replications:
@@
 
-Through \code{rfast99} function, S3 object for class 'rfast99' will be created.
-The generated parameters are stored as a 3-D array under the named \code{a} with the dimension of sample number, replication, and parameter number, respectively.
+Through \code{rfast99} function, S3 object with class 'rfast99' will be created.
+The generated parameters are stored as a 3-D array under the named \code{a} with the dimension of sample size, number of replication, and number of parameter, respectively.
 
@@
 ```{r, eval=T, fig.cap = '\\label{fig:rep-sampl}The parameter search curve generated by eFAST method.
-The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.'}
+The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.
+The bar plot shows the distribution of sampling parameter'}
 par(mfrow=c(4,4),mar=c(0,0,0,0),oma=c(3,3,2,1), pch =".")
@@
 
-Because the physiologically based kinetic model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements therefore have the time-dependent relationships for each model parameter.
+Because the physiologically based kinetic model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements, therefore, have the time-dependent relationships for each model parameter.
 Here we define the output time points (\code{t}) to examine the change of the parameter sensitivity over time.
@@
 
-To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours with time segment of 1 hour as above definition in \code{ode} function in this example.
-The initial time point should avoid 0 to prevent the computational error in misconduction.
+To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours with the time segment of 1 hour as the above definition in \code{ode} function in this example.
+The initial time point should avoid 0 to prevent computational error in misconduct.
 The \code{outnames}, \code{dllname}, are based on the arguments from the \code{ode} function in \pkg{deSolve} package.
@@
 and \code{pbtk1comp_inits.R} files.
-The \code{out} is an S3 object of \code{rfast99} as well, which can link with \code{print}, \code{plot}, and \code{check} method to exame the sensitivity measurments.
+The \code{out} is an S3 object of \code{rfast99} as well, which can link with \code{print}, \code{plot}, and \code{check} method to examine the sensitivity measurements.
 The \code{print} function gives the sensitivity and convergence indices for main, interaction, and total order at each time point.
@@
 Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-The current result is only based on the seting distribution of model parameters for theophylline.
+The current result is only based on the setting distribution of model parameters for theophylline.
 The different given conditions will change the result (result not shown).
@@
 The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is mainly determined by \code{ke}.
-Same as \code{Ccompartment}, the \code{kgutabs} contribute abour 30 - 40% variation of model output in the first hour.
+Same as \code{Ccompartment}, the \code{kgutabs} contribute about 30 - 40% variation of model output in the first hour.
 
-In addtion to create the time-sI profile to investigate the parameter impact on model output, the relationship between parameter and corresponding value of model output can be used to perform examination.
+In addition to use time-sI profile to investigate the parameter impact on model output, we can construct the relationship between parameter and the corresponding value of model output in the examination.
 Figure \ref{fig:SI-output} show the concentration of rest body and the model parameters:
@@
 We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameter with significant relationship with model output have hight impact to dominate the modeling result.
-The \code{ke} shows the high correlation at the lating time (t = 24.01 h).
+The \code{ke} shows the high correlation at the lateing time (t = 24.01 h).
 The \code{ke} doesn't show the obvious relationship with \code{Ccompartment}.
@@
 The convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
-The output are formated as 4-D array in \code{y} with the dimension name of model evaluation, replication, time, and variable, respectively.
+The output are formated as 4-D array in \code{y} with the dimension name of model evaluation, number of replication, number of time points, and number of variable, respectively.
 
@@
 
-The \code{check} is a useful function to determine which parameters have relative lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
+The \code{check} is a useful function to determine which parameters have relatively lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
+The \code{check} also provides some feasible argument to specify the target output or change the cut-off value.
 The argument of \code{SI.cutoff} is setting at 0.05 to detect the relative non-influential parameters as default.
@@
 However, all parameters do not converge to the setting cut-off, which means the larger sample size is required in further sensitivity testing.
-Same as \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use this assignment to exam the given output.
+Same as \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to exam the given output.
 
@@
 
+The following code is used to compile the \pkg{GNU MCSim} model code to the exexutable program.
+
 ```{r, eval=T, message=F, warning=F}
@@
 
-Similar to \code{solve_fun} function that can define the initial parameter and state values through input function, the \code{solve_mcsim} has a \code{condition} argument that is used to give the specific input value such as oral dose, fixing parameter value or initial state variable.
+Similar to \code{solve_fun} function that can define the initial parameter and state values through generated functions, the \code{solve_mcsim} has also a \code{condition} argument that can be used to give the specific input value such as oral dose, fixing parameter value or initial state variable.
 
@@
 
-After soving the equations under the same given condition, we can find that (\pkg{GNU MCSim}) has about 3 times fatser in computing performance than using \pkg{deSolve}.
+After solving the equations under the same given condition, we can find that \pkg{GNU MCSim} has about 4 times faster in computing performance than using \pkg{deSolve}.
 
@@
 
-### Uncertainty and sensitivity analysis
-
-The aim of this section is to reproduce our previous published result \citep{fphar201800588} of global sensitivity analysis for acetaminophen PBPK model through \pkg{pksensi}.
+The aim of this section is to reproduce our previous published result [@fphar201800588] of global sensitivity analysis for acetaminophen PBPK model through \pkg{pksensi}.
 The model codes are included in this package and can be generated through \code{pbpk_apap_model} function.
 We applied the global sensitivity analysis workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
-The descriptions of each parameter and the sampling ranges are list in Table 1.
+The descriptions of each parameter and the sampling ranges are list in Table \ref{tab:data}.
 
@@
 
-print(xtable(df, caption = "Description of sampling range of model parameter"),
+print(xtable(df, caption = "\\label{tab:data}Description of sampling range of model parameter"),
       type = "latex", 
@@
 
-Same as the example of one-compartment PBTK model.
+Same as the example of one-compartment PK model.
 The model parameter and the corresponding sampling range should be defined to create the parameter matrix.
@@
 To rapidly reach the acceptance convergence, we apply uniform distribution for all testing parameters.
-The ranges of informative parameters are set to 1.96-times difference for single side (approximate 54.6 times difference between minimum and maximum) under log-scaled.
+The ranges of informative parameters are set to 1.96-times difference for single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
 The nominal values of informative model parameters were defined as:
@@
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
-The \code{generate_infile()} provide the arguments of \code{rtol} and \code{atol} that can be adjusted to prevent the unwanted error.
+The \code{generate_infile} and \code{solve_mcsim} provide the arguments of \code{rtol} and \code{atol} that can be adjusted to prevent the unwanted error.
 However, the modification will lower computational speed.
@@
 The maximum number of step is set to 5000 in this case.
+Here we seperate the GSA of APAP-PBPK model process to several steps.
 
-In this test case, we adjusted the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} from U(0, 15) to U(0, 10).
-The relative and absolute error tolerance were set to 1e-7 and 1e-9, respectively, to prevent the computational error in \pkg{GNU MCSim},
+### Construct and compile the model file
+
+The model code needs to be prepared in the following GSA workflow.
+After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through \code{compile_model}.
+
+```{r, eval=T}
+mName <- "pbpk_apap"
+pbpk_apap_model()
+compile_model(mName, application = "mcsim", version = "6.1.0")
+```
+
+### Define the parameter and its distribution
+
+The 21 testing model parameters are defined in this part, including parameter name, probability distribution, and distribution parameter.
+To prevent the computational error, we adjusted the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} from $U(0, 15)$ to $U(0, 10)$ in this case.
+The objects \code{q} and \code{dist} are used to set the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for sensitivity analysis).
 
@@
            "lnkGA_syn","lnkPAPS_syn", "lnCLC_APAP","lnCLC_AG","lnCLC_AS")
-q <- "qunif"
-q.arg <-list(list(Tg-rng, Tg+rng),
-             list(Tp-rng, Tp+rng),
-             list(CYP_Km-rng, CYP_Km+rng),
-             list(-2., 5.),
+dist <- rep("Uniform", 21)
+q <- rep("qunif", 21)
+q.arg <-list(list(Tg-rng, Tg+rng), list(Tp-rng, Tp+rng), 
+             list(CYP_Km-rng, CYP_Km+rng), list(-2., 5.),
              list(SULT_Km_apap-rng, SULT_Km_apap+rng),
@@
              list(SULT_Km_paps-rng, SULT_Km_paps+rng),
-             list(0, 10),
-             list(UGT_Km-rng, UGT_Km+rng),
+             list(0, 10), list(UGT_Km-rng, UGT_Km+rng),
              list(UGT_Ki-rng, UGT_Ki+rng),
              list(UGT_Km_GA-rng, UGT_Km_GA+rng),
-             list(0, 10),
-             list(Km_AG-rng, Km_AG+rng),
-             list(7., 15),
-             list(Km_AS-rng, Km_AS+rng),
-             list(7., 15),
-             list(0., 13),
-             list(0., 13),
-             list(-6., 1),
-             list(-6., 1),
-             list(-6., 1))
+             list(0, 10), list(Km_AG-rng, Km_AG+rng),
+             list(7., 15), list(Km_AS-rng, Km_AS+rng),
+             list(7., 15), list(0., 13), list(0., 13),
+             list(-6., 1), list(-6., 1), list(-6., 1))
+```
 
+### Define additional input condition and output time and variables
+
+To improve computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
+The setting oral dose of APAP is 20 mg/kg in this example.
+Generally, the input dosing method can be defined through the \code{condition} argument.
+Since the unit of the given dose is mg/kg, the \code{mgkg_flag} is set to 1.
+More definition of input schedule functions can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
+
+```{r, eval=T}
+conditions <- c("mgkg_flag = 1",
+                "OralExp_APAP = NDoses(2, 1, 0, 0, 0.001)",
+                "OralDose_APAP_mgkg = 20.0")
+vars <- c("lnCPL_APAP_mcgL", "lnCPL_AG_mcgL", "lnCPL_AS_mcgL")
 times <- seq(from = 0.1, to = 12.1, by = 0.2)
+```
+
+### Uncertainty analysis
+
+We apply uncertainty analysis through the \code{solve_mcsim} and visualize the result by \code{pksim}.
+Some example APAP PK data are included in the \pkg{pksensi} with experiment time (h) and concentration (mg/L).
+
+```{r}
+head(APAP)
+```
+
+In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the computational error.
+The Monte Carlo simulation is run fun for 1000 iteration as the assignment of \code{monte_carlo}.
+The input file ('sim.in') and output file ('simmc.out') will be generated under the standard ASCII format.
+The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
+
+```{r}
+set.seed(1111)
+out <- solve_mcsim(mName = mName, params = params, vars = vars,
+                   monte_carlo = 1000, dist = dist, q.arg = q.arg, 
+                   time = times, condition = conditions, 
+                   rtol = 1e-7, atol = 1e-9)
+```
+
+Figure \ref{fig:pksim-MC} shows the result of the uncertainty analysis of APAP data and modeling results.
+For parent compound, all data points are located in the simulated interval of 25-75%.
+Through this result, we can determine that the simulated outputs can accurately generate the same concentration profile as the in-vivo experiment under the setting of parameter ranges for APAP.
+The simulated result of metabolites APAP-G shows the different PK profile with experiment data.
+However, all data points are located in the simulated interval.
+
+```{r, fig.cap = '\\label{fig:pksim-MC}Coverage checks of prior PBPK model predictions with calibrated APAP data', fig.height=3}
+par(mfrow = c(1,3), mar = c(4,4,1,1))
+pksim(out, xlab = "Time (h)", ylab = "Conc. (ug/L)", main = "APAP")
+points(APAP$Time, log(APAP$APAP * 1000))
+pksim(out, vars = "lnCPL_AG_mcgL", xlab = "Time (h)", main = "APAP-G", 
+      ylab = " ", legend = FALSE)
+points(APAP$Time, log(APAP$AG * 1000))
+pksim(out, vars = "lnCPL_AS_mcgL", xlab = "Time (h)", main = "APAP-S", 
+      ylab = " ", legend = FALSE)
+points(APAP$Time, log(APAP$AS * 1000))
+```
+
+### Generate parameter matrix
+
+In GSA, we have to additionally generate the parameter matrix from the eFAST method. The current setting uses 512 sample size with 10 replication.
+
+```{r, eval=T}
 set.seed(1234)
@@
 
-After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable files (\code{mcsim.pbpk_apap}) through \code{compile_model()}.
+### Conduct the GSA
+
+To conduct the GSA with \pkg{GNU MCsim} and \pkg{pksensi}, the input file with given "Setpoint" condition should be generated before modeling.
+The file can create by \code{generate_infile} function.
+The \code{solve_mcsim} can also automatically the input file and compute the output.
 
 ```{r, eval=T}
-mName <- "pbpk_apap"
-pbpk_apap_model()
-compile_model(mName, application = "mcsim", version = "6.1.0")
+system.time(out <- solve_mcsim(x, mName = mName,
+                               params = params, 
+                               time = times, 
+                               vars = vars,
+                               condition = conditions, 
+                               rtol = 1e-7, atol = 1e-9))
 ```
 
-To improve computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
-The setting oral dose of APAP is 20 mg/kg in this example.
-Generally, the input dosing method can be defined through the \code{condition} argument.
-Since the unit of the given dose is mg/kg, the \code{mgkg_flag} is set to 1 to declare the statement.
-More definition of input functions can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
+### Visualization and decision
 
-```{r, eval=T}
-vars <- c("lnCPL_APAP_mcgL", "lnCPL_AG_mcgL", "lnCPL_AS_mcgL")
-conditions <- c("mgkg_flag = 1",
-                "OralExp_APAP = NDoses(2, 1, 0, 0, 0.001)",
-                "OralDose_APAP_mgkg = 20.0")
-generate_infile(params = params,
-                vars = vars,
-                time = times, 
-                condition = conditions,
-                rtol = 1e-7, atol = 1e-9)
-system.time(out <- solve_mcsim(x, mName = mName, 
-                             params = params,
-                             vars = vars,
-                             time = times,
-                             condition = conditions,
-                             generate.infile = F))
+The plotting function can create the result of time-dependent sensitivity measurement to determine the parameter impact on model output over time (Figure \ref{fig:APAP-sensi}).
+
+```{r, eval=T, fig.height=8, fig.width=8, fig.cap = '\\label{fig:APAP-sensi}Time-dependent sensitivity indices of the plasma APAP concentration estimated from APAP PBPK model during 12-hr time period post intake.'}
+plot(out, vars = "lnCPL_APAP_mcgL")
 ```
 
-The plotting function can output the result of time-dependent sensitivity measurement to determine the parameter impact on model output over time (Figure 4).
+In addition, through using the \code{check}, the parameter with sensitivity and convergence indices over the given condition can be preliminary detected for all output variables.
+Based on our previous study, we proposed the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.
+Through the given argument \code{order}, we can select the specific order of sensitivity measurement that we're interested in (Figure \ref{fig:Heat-SI}).
 
-```{r, eval=T, fig.height=8, fig.width=8, fig.cap = ' '}
-plot(out, vars = "lnCPL_AG_mcgL")
-```
-
-The uncertainty analysis is a crucial step before model calibration.
-We can apply uncertainty analysis through the `pksim()` by the given name of output variables (Figure 5).
-Through this visualization approach, we can recognize whether the simulated outputs can accurately simulate the same concentration profile as the in-vivo experiment under the setting of parameter ranges.
-In Figure 5, all experiment points are included in the intervals, representing the acceptable set of the parameter range.
-
-```{r, eval=T, fig.cap = ' '}
-par(mfrow = c(2,2), mar = c(2,2,1,1), oma = c(2,2,1,1))
-pksim(out$y, vars = "lnCPL_APAP_mcgL")
-text(1, 15, "APAP",cex = 1.2)
-points(APAP$Time, log(APAP$APAP * 1000))
-pksim(out$y, vars = "lnCPL_AG_mcgL", legend = F)
-text(1, 15, "AG",cex = 1.2)
-points(APAP$Time, log(APAP$AG * 1000))
-pksim(out$y, vars = "lnCPL_AS_mcgL", legend = F)
-text(1, 15, "AS",cex = 1.2)
-points(APAP$Time, log(APAP$AS * 1000))
-mtext("Time", SOUTH<-1, line=0.4, cex=1.2, outer=TRUE)
-mtext("Conc.", WEST<-2, line=0.4, cex=1.2, outer=TRUE)
-```
-
-In addition, through using the \code{check()}, the parameter with sensitivity and convergence indices over the given condition can be easily detected for all output variables.
-
-```{r, eval=T}
-check(out)
-```
-
-The \code{check()} also provides some feasible argument to specify the target output or change the cut-off value.
-
-```{r, eval=T}
-check(out, vars = "lnCPL_APAP_mcgL", SI.cutoff = 0.1, CI.cutoff = 0.1)
-```
-
-### Heatmap visualization combined with an index “cut-off”
-
-Based on our previous study, we proposed the heatmap visualization approach to distinguish "influential" and "non-influential" parameters with a cut-off.
-Through the given argument \code{order}, we can select the specific order of sensitivity measurement that we're interested in (Figure 6 & 7).
-
-```{r, eval=T, fig.cap = ' '}
-heat_check(out, order = "interaction")
-```
-
-```{r, eval=T, fig.cap = ' '}
+```{r, eval=T, fig.cap = '\\label{fig:Heat-SI}Heat map representation of time-dependent total SI computed for GSA method', fig.height=5}
 heat_check(out, order = "total order")
@@
 
-Also, adding the `index = "CI"` in the function can further investigate the convergence of the sensitivity index.
-Based on the current setting of sampling number, most parameters cannot reach the acceptable criteria of convergence (Figure 8).
-Therefore, the higher number of sampling is necessary.
+Adding the `index = "CI"` in the function can further investigate the convergence of the sensitivity index.
+Based on the current setting of sampling size, most parameters cannot reach the acceptable criteria of convergence (Figure \ref{fig:Heat-CI}).
+Therefore, a higher number of sampling is necessary.
+The sample size of convergence in the current PBPK model is 8,192 [@fphar201800588].
 
-```{r, eval=T, fig.cap = ' '}
-heat_check(out, index = "CI", CI.cutoff = 0.05, order = "total order")
+```{r, eval=T, fig.cap = '\\label{fig:Heat-CI}Heat map representation of time-dependent total CI computed for GSA method', fig.height=5}
+heat_check(out, index = "CI", order = "total order")
 ```

### 2019-06-25 manuscript/manuscript.Rmd
manuscript

@@
 In our previous work, we applied a global sensitivity analysis workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-Although several sensitivity analysis algorithms are available, no comprehensive package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize sensitivity analysis results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
+Although several sensitivity analysis algorithms are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize sensitivity analysis results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
 Therefore, we developed an open-source R package, called **pksensi**, to fill this need and make sensitivity analysis more accessible in pharmacological and toxicological research.
@@
   # at least one keyword must be supplied
-  formatted: [Pharmacokinetics, Physiologically based pharmacokinetics, Sensitivity analysis]
-  plain:     [Pharmacokinetics, Physiologically based pharmacokinetics, Sensitivity analysis]
+  formatted: [Pharmacokinetics, Physiologically based kinetics, Global sensitivity analysis]
+  plain:     [Pharmacokinetics, Physiologically based kinetics, Global sensitivity analysis]
 preamble: >
   \usepackage{amsmath}
+  \usepackage{booktabs}
 bibliography: references.bib 
-geometry: margin=0.9in
 output: rticles::jss_article
@@
 \pkg{ODEsensitivity} can perform sensitivity analysis in ordinary differential equation (ODE) models \citep{R-ODEsensitivity}.
-It uses the ODE interface from the \pkg{deSolve} package, connecting it with the sensitivity analysis from the \pkg{sensitivity} package, and can perform sensitivity analysis on models with multivariate output.
+It uses the ordinary differential equation (ODE) interface from the \pkg{deSolve} package, connecting it with the sensitivity analysis from the \pkg{sensitivity} package, and can perform sensitivity analysis on models with multivariate output.
 Other related packages such as \pkg{mtk} \citep{RJ-2015-031} and \pkg{fast} \citep{R-fast} are designed to investigate the influence of model parameter.
@@
 
-## Monte Carlo
-
-The uncertainty analysis is a crucial modeling process with sensitivity analysis.
-Sometimes, it's difficult to have informative information to set up the parameter distribution.
-The wider range of parameter distribution might cause the computation error at the extreme value.
-The narrower range might cause the incorrect calibrated result of the model parameter.
-To prevent this problem.
-The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process.
-Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
-
 ## Pharmacokinetic modeling
@@
 The \pkg{pksensi} can also link with \proglang{GNU MCSim} to compile the model code, used in solving each system of equations.
+
+The uncertainty analysis is a crucial modeling process with sensitivity analysis.
+Sometimes, it's difficult to have informative information to set up the parameter distribution.
+The wider range of parameter distribution might cause the computation error at the extreme value.
+The narrower range might cause the incorrect calibrated result of the model parameter.
+To prevent this problem.
+The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process.
+Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
 
@@
 
+The code must first be compiled to run the model.
 After create the model file, use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
@@
 params <- c("vdist", "ke", "kgutabs", "BW")
+```
+
+The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
+
+```{r, eval=T}
 set.seed(1234)
@@
 The input file ('sim.in') and output file ('simmc.out') will be generated under the standard ASCII format.
-The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
 
@@
 
+The \pkg{GNU MCSim} is the powerful simulation package, that can conduct various dynamical simulations for biological modeling, and is mainly build for pharmacokinetic research with probabiliistic approaches [@bois2009gnu].
+The \Proglang{R} is a comprehensive statistical computing language that can conduct the probabilistic programming and have tons of packages in statitical analysis and data visualization.
+In dynamic simulation, .
+
+The dosing schedule
+
+Some studies applied another GSA algorithm in the physiologically and biologically based ODE model, such as compartmental absorption and transit (also known as CAT) and other system biology model [@melillo2019variance; @sumner2012methodology].
+In future version, we expect to add additional GSA approach that had been applied in advance study.
+
+Principle component
+
+parallel
+
 # Acknowledgments
@@
 
+# Session information
+
+```{r results = 'asis'}
+toLatex(sessionInfo())
+```
+
 # References

### 2019-06-25 manuscript/manuscript.Rmd
Concluding

@@
 
-The \pkg{GNU MCSim} is the powerful simulation package, that can conduct various dynamical simulations for biological modeling, and is mainly build for pharmacokinetic research with probabiliistic approaches [@bois2009gnu].
-The \Proglang{R} is a comprehensive statistical computing language that can conduct the probabilistic programming and have tons of packages in statitical analysis and data visualization.
-In dynamic simulation, .
+The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct various dynamical simulations for biological-based modeling, and it's mainly built for pharmacokinetic research with probabilistic approaches [@bois2009gnu].
+The \Proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
+Some fundamental packages on the CRAN repository and other software platforms hasd been built to perform different types of global sensitivity analysis.
+The development of \pkg{pksensi} is mainly for physiologically based kinetic models that are constructed by differential equations.
+It provides a straightforward application to investigate the impact of parameters on model output across time and variables.
+We adopted the eFSAT method, a common variance-based approach and had better efficiency and accuracy in the model assessment.
+
+in \pkg{pksensi}
+
+  have the developed solver to solve the differential equation in dynamic simulation.
 
@@
 Principle component
-
-parallel
 

### 2019-06-26 manuscript/manuscript.Rmd
Concluding

@@
 We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
-It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as Jansen or OWen method.
+It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from \citep{jansen1999analysis} and \citep{owen2013better}.
 Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
@@
 *	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \pkg{GNU MCSim} 
-*	Decision making: The output results and visualization tools can be used to easily determine which parameters have "non-influential" effects on the model output and can be fixed in model calibration.
+*	Decision support: The output results and visualization tools can be used to easily determine which parameters have "non-influential" effects on the model output and can be fixed in model calibration.
 
@@
 The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct various dynamical simulations for biological-based modeling, and it's mainly built for pharmacokinetic research with probabilistic approaches [@bois2009gnu].
-The \Proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
-Some fundamental packages on the CRAN repository and other software platforms hasd been built to perform different types of global sensitivity analysis.
+The \proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
+The current version had included \pkg{GNU MCSim} and \pkg{deSolve} as the main solver in our model assessment.
+Compare with \pkg{deSolve} package, we found that The \pkg{GNU MCSim} can provide better efficiency to solve the ODE in PK model.
+Some developed package in \proglang{R} had also provide the feature to solve ODE such as \pkg{RxODE} and \pkg{mrgsolve} [@wang2016tutorial; @R-mrgsolve].
+Therefore, we hope to incorporate this package in our GSA workflow to provide user more options in the modeling.
+
+Moreover, some fundamental packages on the CRAN repository and other software platforms had been built to perform different types of GSA.
 The development of \pkg{pksensi} is mainly for physiologically based kinetic models that are constructed by differential equations.
 It provides a straightforward application to investigate the impact of parameters on model output across time and variables.
-We adopted the eFSAT method, a common variance-based approach and had better efficiency and accuracy in the model assessment.
+We adopted the eFSAT method, a common variance-based sensitivity analysis that had better efficiency and accuracy in the model assessment.
+It can exam the parameter effect from first order to the total order.
+In addition, the package can not only conduct GSA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most GSA studies and its software package.
+We also develop the function to visualize the output result that can help us determine the "influential" and "noninfluential" parameters that can be applied in parameter fixing in the model Calibration.
+In addition to eFAST, Some studies applied another GSA algorithm in the physiologically and biologically based ODE model, such as compartmental absorption and transit (also known as CAT) and other system biology model [@melillo2019variance; @sumner2012methodology].
+In future version, we expect to add additional GSA approach that had been applied in advance study.
+The other features such as principle component analysis [@lamboni2011multivariate] were considered to include in \pkg{pksensi}.
 
-in \pkg{pksensi}
-
-  have the developed solver to solve the differential equation in dynamic simulation.
-
-The dosing schedule
-
-Some studies applied another GSA algorithm in the physiologically and biologically based ODE model, such as compartmental absorption and transit (also known as CAT) and other system biology model [@melillo2019variance; @sumner2012methodology].
-In future version, we expect to add additional GSA approach that had been applied in advance study.
-
-Principle component
+It is worth noting that the GSA results are limited to the specific model and the given range of the model parameter.
+The provided conditions, such as the number of examing parameter, parameter distribution, parameter range, and the given dosing level in PK model might have a potential impact on the level of importance of each parameter.
+The misconduct of the setting condition might lead to the incorrect estimated result.
+In other words, if we have various types of experiment data (e.g.
+different exposure routes and biological measurements) and will use them in model calibration, it will enhance the difficulties in GSA.
+The complex model with a high dimension of the parameter is also a challenge in GSA, which will need higher computer demand in the model assessment.
+Overall, we have developed \pkg{pksensi} that can apply the comprehensive GSA workflow for most PK models.
+More additional applications still need to be investigated and strengthen.
 

### 2019-06-26 manuscript/manuscript.Rmd
manuscript

@@
 title:
-  formatted: "\\pkg{pksensi}: An R package to apply sensitivity analysis in pharmacokinetic models"
+  formatted: "\\pkg{pksensi}: An R package to apply global sensitivity analysis in physiologically based kinetic models"
   # If you use tex in the formatted title, also supply version without
-  plain:     "pksensi: An R package an R package to apply sensitivity analysis in pharmacokinetic models"
+  plain:     "pksensi: An R package an R package to apply global sensitivity analysis in physiologically based kinetic models"
   # For running headers, if needed
@@
 It is also increasingly used in developing and assessing pharmacokinetic models.
-In our previous work, we applied a global sensitivity analysis workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+In our earlier work, we applied a global sensitivity analysis workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
 Although several sensitivity analysis algorithms are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize sensitivity analysis results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
@@
 Utilizing pksensi, we successfully reproduced our previously published sensitivity analysis results of human PBPK modeling for acetaminophen and its two primary metabolites.
-Overall, pksensi improves the user experience of performing sensitivity analysis and can create robust and reproducible results for decision making in pharmacokinetic model calibration.
+Overall, **pksensi** improves the user experience of performing sensitivity analysis and can create robust and reproducible results for decision making in pharmacokinetic model calibration.
 keywords:
@@
 Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs.
-An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions \citep{ferretti2016trends}.
-This technique has also been applied in pharmacology and toxicology research \citep{loizou2015application, mcnally2012reconstruction}.
+An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions [@ferretti2016trends].
+This technique has also been applied in pharmacology and toxicology research [@loizou2015application; @mcnally2012reconstruction].
 Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time.
 The goal of sensitivity analysis in pharmacokinetic research is to examine the sensitivity of output variables (e.g.
-compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants \citep{mcnally2011workflow}.
-It can be further applied to parameter prioritization and parameter fixing before model calibration \citep{fphar201800588}.
+compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
+It can be further applied to parameter prioritization and parameter fixing before model calibration [@fphar201800588].
 
-In our previous work \citep{fphar201800588}, we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-We used GNU MCSim \citep{JSSv002i09}, an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
+In our earlier work [@fphar201800588], we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+We used GNU MCSim [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
 We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
-It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from \citep{jansen1999analysis} and \citep{owen2013better}.
+It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from @jansen1999analysis and @owen2013better.
 Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
@@
 
-Most of sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} \citep{pianosi2015matlab}, \proglang{python} \citep{herman2017salib}, and \proglang{R} \citep{iooss2017introduction}.
-Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package \citep{R-sensitivity}, which is a practical tool to conduct local and global sensitivity analysis.
+Most of sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} [@pianosi2015matlab], \proglang{python} [@herman2017salib], and \proglang{R} [@iooss2017introduction].
+Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package [@R-sensitivity], which is a practical tool to conduct local and global sensitivity analysis.
 The \pkg{sensitivity} package includes some functions to generate the parameter sequences by using different computing algorithms.
@@
 Several other \proglang{R} packages, such as \pkg{FME}, \pkg{multisensi}, and \pkg{ODEsensitivity} include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network (CRAN).
-The \pkg{FME} package contains a basic method for performing local sensitivity analysis for dynamic models \citep{JSSv033i03}.
-The \pkg{multisensi} package provides sensitivity analysis for models with multivariate output \citep{R-multisensi}.
-\pkg{ODEsensitivity} can perform sensitivity analysis in ordinary differential equation (ODE) models \citep{R-ODEsensitivity}.
-It uses the ordinary differential equation (ODE) interface from the \pkg{deSolve} package, connecting it with the sensitivity analysis from the \pkg{sensitivity} package, and can perform sensitivity analysis on models with multivariate output.
-Other related packages such as \pkg{mtk} \citep{RJ-2015-031} and \pkg{fast} \citep{R-fast} are designed to investigate the influence of model parameter.
-Although these tools provide various approaches to perform sensitivity analysis, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic modeling.
+The \pkg{FME} package contains a basic method for performing local sensitivity analysis for dynamic models [@JSSv033i03].
+The \pkg{multisensi} package provides sensitivity analysis for models with multivariate output [@R-multisensi].
+\pkg{ODEsensitivity} can perform sensitivity analysis in ordinary differential equation (ODE) models [@R-ODEsensitivity].
+It utilizes the ODE interface from the \pkg{deSolve} package and connect it with the sensitivity analysis from the \pkg{sensitivity} package.
+Other related packages such as \pkg{mtk} [@RJ-2015-031] and \pkg{fast} [@R-fast] are designed to investigate the importance of model parameter.
+Although these tools provide various approaches to perform sensitivity analysis, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic and physiologically based kinetic (PK) modeling.
 
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
-This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK, and advanced compartment absorption and transit models with multivariate model output.
+This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic and PK model with multivariate model outputs.
 The design concepts of \pkg{pksensi} are:
@@
 
-In the beginning, the user need to pre-install \pkg{GNU MCSim} \citep{JSSv002i09} and related standard C compiler to solve the ODE function.
+In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09} and related standard \proglang{C} compiler to solve the ODE function.
 In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
 The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_instal} in \pkg{pksensi}.
-The C compiler had originated in Linux or MacOS.
+The \proglang{C} compiler had originated in Linux or MacOS.
 The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as,
 
-```{r}
+```{r, eval=F}
 Rtools.path <- "c:/Rtools/bin; c:/Rtools/mingw_64/bin"
@@
 
-We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ordinary differential equation-based dynamical models, in this package \citep{saltelli1999quantitative}.
-Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package \citep{R-sensitivity}.
+We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ODE-based dynamical models, in this package [@saltelli1999quantitative].
+Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity].
 The eFAST method compute the sensitivity index (SI) via ANOVA-like decomposition of the function for analysis.
 This type of GSA approach and SI also know as Sobol method and Sobol index.
-Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
+Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
 This can be written as:
@@
 
-where $x_i$ is the nominal value of the i-th parameter, and $\omega_i$  is a vector giving the set of frequencies, one frequency for each parameter.
+where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$  is a vector giving the set of frequencies, one frequency for each parameter.
 The default set of frequencies is based on the study from \cite{saltelli1999quantitative}.
-The first frequency of the vector $\omega$ is assigned to each factor $x_i$ in turn (corresponding to the estimation of Sobol indices of first and total order, respectively).
 Through the random phase-shift, the robust result of the sensitivity measurement should be similar across replications under the same sample size.
 
-To investigate the convergence of sensitivity indices, we adopted the "exam" approach proposed by \citep{sarrazin2016global}.
+To investigate the convergence of sensitivity indices, we adopted the "exam" approach proposed by [@sarrazin2016global].
 This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed sensitivity index for all parameters across all time-points and output variables.
@@
 
-To integrate the random phase-shift with eFAST, we built the \code{rfast99} function, which can be used to perform the sensitivity test for the whole pharmacokinetic modeling process.
+To integrate the random phase-shift with eFAST, we built the \code{rfast99} function, which can be used to perform the sensitivity test for the whole modeling process.
 The function \code{rfast99} in \pkg{pksensi} can sample and generate the testing parameter matrix based on a given argument of sample size n and probability distribution (e.g., mean/s.d.
@@
 
-The R package \pkg{pksensi} provides a useful function to solve ordinary differential equations in pharmacokinetic models using analytical and numerical approaches.
-The solution can be performed under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package \citep{JSSv033i09}.
+The R package \pkg{pksensi} provides a useful function to solve ODE in pharmacokinetic and PK models using analytical and numerical approaches.
+The solution can be performed under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package [@JSSv033i09].
 However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as \proglang{FORTRAN}, \proglang{C}, or \proglang{C++}.
@@
 However, Windows users need to install Rtools to compile the source code by using the GNU GCC compiler.
-More requirements are detailed in the study from \citep{soetaertr}.
-The \proglang{C} implementation of the example model can be found as an example within \proglang{GNU MCSim} v6.0.1 \citep{JSSv002i09}.
+More requirements are detailed in the study from [@soetaertr].
+The \proglang{C} implementation of the example model can be found as an example within \pkg{GNU MCSim} v6.0.1 [@JSSv002i09].
 The \pkg{pksensi} can also link with \proglang{GNU MCSim} to compile the model code, used in solving each system of equations.
@@
 This format is particularly suited for further graphical routines within \pkg{pksensi}.
-Also, time-dependent sensitivity measurements of first and total order effect for all parameters can be plotted on one figure.The sensitivity index might have different estimation across replications under the same sample size.
-The inadequate sample number can create the unrobust result and bias.
-It's essential to use the adequate sample size to prevent the incorrect judgement in parameter fixing after senstivity analysis.
+Also, time-dependent sensitivity measurements of first and total order effect for all parameters can be plotted simultaneously.
+The sensitivity index might have different estimation across replications under the same sample size.
+The inadequate sample number can create an unreliable result and bias.
+It's essential to use the sufficient sample size to prevent incorrect judgment in parameter fixing.
 Therefore, \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
@@
 
-Two examples of physiologically based kinetic model were applied here to demonstrate how \pkg{pksensi} work.
-The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current demo.
+Two examples of PK model were applied here to demonstrate how \pkg{pksensi} work.
+The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current example.
 The version of \pkg{GNU MCSim} is 6.1.0.
@@
 
-In the first example, we use a simple, one-compartment PK model from \pkg{httk} package \citep{JSSv079i04} to demonstrate how \pkg{pksensi} can be applied to physiologically based kinetic studies.
+In the first example, we use a simple, one-compartment PK model from \pkg{httk} package [@JSSv079i04] to demonstrate how \pkg{pksensi} can be applied to PK studies.
 The differential equations for this model is as:
@@
 
-Because the physiologically based kinetic model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements, therefore, have the time-dependent relationships for each model parameter.
+Because the PK model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements, therefore, have the time-dependent relationships for each model parameter.
 Here we define the output time points (\code{t}) to examine the change of the parameter sensitivity over time.
@@
 
-Generally, wide ranges for parameter values might cause the computational error in the solver.
+Generally, wide ranges of parameter values might cause the computational error in the solver.
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
@@
 Moreover, some fundamental packages on the CRAN repository and other software platforms had been built to perform different types of GSA.
-The development of \pkg{pksensi} is mainly for physiologically based kinetic models that are constructed by differential equations.
+The development of \pkg{pksensi} is mainly for PK models that are constructed by differential equations.
 It provides a straightforward application to investigate the impact of parameters on model output across time and variables.
@@
 In addition, the package can not only conduct GSA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most GSA studies and its software package.
-We also develop the function to visualize the output result that can help us determine the "influential" and "noninfluential" parameters that can be applied in parameter fixing in the model Calibration.
-In addition to eFAST, Some studies applied another GSA algorithm in the physiologically and biologically based ODE model, such as compartmental absorption and transit (also known as CAT) and other system biology model [@melillo2019variance; @sumner2012methodology].
+We also develop the function to visualize the output result that can help us determine the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model Calibration.
+In addition to eFAST, Some studies applied another GSA algorithm in the physiologically and biologically based ODE models, such as compartmental absorption and transit (also known as CAT) and other system biology models [@melillo2019variance; @sumner2012methodology].
 In future version, we expect to add additional GSA approach that had been applied in advance study.
@@
 The complex model with a high dimension of the parameter is also a challenge in GSA, which will need higher computer demand in the model assessment.
-Overall, we have developed \pkg{pksensi} that can apply the comprehensive GSA workflow for most PK models.
+Overall,
+Overall, \pkg{pksensi} improves the user experience of performing GSA and can create robust and reproducible results for decision making in pharmacokinetic model calibration.
 More additional applications still need to be investigated and strengthen.

### 2019-06-26 manuscript.Rmd
move files

--- 2019-06-26 manuscript/manuscript.Rmd
@@
 In our earlier work [@fphar201800588], we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-We used GNU MCSim [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
+We used \pkg{GNU MCSim} [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
 We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
@@
 
-In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09} and related standard \proglang{C} compiler to solve the ODE function.
+In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
 In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
@@
 
-```{r, eval=F}
+```{r, eval=T}
 Rtools.path <- "c:/Rtools/bin; c:/Rtools/mingw_64/bin"
@@
 
-We adopted eFAST (extended Fourier Amplitude Sensitivity Test), a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ODE-based dynamical models, in this package [@saltelli1999quantitative].
+We adopted eFAST, a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ODE-based dynamical models, in this package [@saltelli1999quantitative].
 Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity].
@@
 
-## Pharmacokinetic modeling
+## Modeling
 
@@
 The uncertainty analysis is a crucial modeling process with sensitivity analysis.
-Sometimes, it's difficult to have informative information to set up the parameter distribution.
+Sometimes, it's difficult to have informative data to set up the parameter distribution.
 The wider range of parameter distribution might cause the computation error at the extreme value.
@@
 The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process.
-Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
+Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.
+\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
+The general workflow of uncertainty and sensitivityanalysis is illustrated in Figure \ref{fig:Workflow}.
+
+```{r workflow, echo = FALSE, out.width = "100%", fig.cap = '\\label{fig:Workflow}Workflow for performing uncertainty and sensitivity analysis in \\pkg{pksensi}.'}
+knitr::include_graphics("diagram/workflow.png")
+```
 
@@
 
-The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.model.
+The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.
 The model mainly includes two state variables that are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).

### 2019-06-26 manuscript.Rmd
workflow

@@
 Although several sensitivity analysis algorithms are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize sensitivity analysis results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
-Therefore, we developed an open-source R package, called **pksensi**, to fill this need and make sensitivity analysis more accessible in pharmacological and toxicological research.
+Therefore, we developed an open-source \proglang{R} package, called **pksensi**, to fill this need and make sensitivity analysis more accessible in pharmacological and toxicological research.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK model with multivariate model outputs (multiple time-points and tissue compartments).

### 2019-06-28 manuscript.Rmd
manuscript

--- 2019-06-26 manuscript.Rmd
@@
   # If you use tex in the formatted title, also supply version without
-  plain:     "pksensi: An R package an R package to apply global sensitivity analysis in physiologically based kinetic models"
+  plain:     "pksensi: An R package to apply global sensitivity analysis in physiologically based kinetic models"
   # For running headers, if needed
@@
 abstract: >
-  Sensitivity analysis is an essential tool for modelers to understand the influence of model parameters on model outputs.
+  Sensitivity analysis (SA) is an essential tool for modelers to understand the influence of model parameters on model outputs.
 It is also increasingly used in developing and assessing pharmacokinetic models.
-In our earlier work, we applied a global sensitivity analysis workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-Although several sensitivity analysis algorithms are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize sensitivity analysis results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
-Therefore, we developed an open-source R package, called **pksensi**, to fill this need and make sensitivity analysis more accessible in pharmacological and toxicological research.
+In our earlier work, we applied a global SA workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+Although several SA algorithms are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize SA results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
+Therefore, we developed an open-source \proglang{R} package, called **pksensi**, to fill this need and make SA more accessible in pharmacological and toxicological research.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK model with multivariate model outputs (multiple time-points and tissue compartments).
-It also refined the extended Fourier Amplitude Sensitivity Test method for sensitivity analysis by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
+It also refined the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
 Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
-Utilizing pksensi, we successfully reproduced our previously published sensitivity analysis results of human PBPK modeling for acetaminophen and its two primary metabolites.
-Overall, **pksensi** improves the user experience of performing sensitivity analysis and can create robust and reproducible results for decision making in pharmacokinetic model calibration.
+Utilizing pksensi, we successfully reproduced our previously published SA results of human PBPK modeling for acetaminophen and its two primary metabolites.
+Overall, **pksensi** improves the user experience of performing SA and can create robust and reproducible results for decision making in pharmacokinetic model calibration.
 keywords:
@@
 
-Sensitivity analysis is a mathematical technique to investigate how variations in model parameters affect model outputs.
-An increasing number of studies use sensitivity analysis to determine which model parameters contribute to high variation in model predictions [@ferretti2016trends].
+Sensitivity analysis (SA) is a mathematical technique to investigate how variations in model parameters affect model outputs.
+An increasing number of studies use SA approach to determine which model parameters contribute to high variation in model predictions [@ferretti2016trends].
 This technique has also been applied in pharmacology and toxicology research [@loizou2015application; @mcnally2012reconstruction].
-Pharmacokinetic modeling describes the changes in the concentrations or amounts of a substance within model compartments over time.
-The goal of sensitivity analysis in pharmacokinetic research is to examine the sensitivity of output variables (e.g.
-compound concentration in blood or tissues) in pharmacokinetic models responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
+Physiologically based kinetic (PK) modeling describes the changes in the concentrations or amounts of a substance within model compartments over time.
+The goal of SA in PK research is to examine the sensitivity of output variables (e.g.
+chemical concentration in blood or tissues) in model responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
+This technique can be guidance of experimental design, parameter estimation during the model establishment, and qualification processes [@zhang2015sobol].
+Through SA, it can optimal experimental design and distinguish the "unidentifiable" parameters that can lead to problems with numerical procedures used in parameter estimation [@chu2010quantitative].
 It can be further applied to parameter prioritization and parameter fixing before model calibration [@fphar201800588].
 
-In our earlier work [@fphar201800588], we developed an approach to apply global sensitivity analysis (GSA) workflow to reduce the computational burden in the Bayesian,  Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+In our earlier work [@fphar201800588], we developed an approach to apply global SA workflow to reduce the computational burden in the Bayesian, Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
 We used \pkg{GNU MCSim} [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
-We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based GSA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
-It's more comptational efficient in calculate the influence of model parameter than other GSA methods via Monte Carlo integration, such as the GSA approach from @jansen1999analysis and @owen2013better.
+We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based global SA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
+This method was first introduced in @cukier1973study (the classic FAST method with first-order effect only) that study of the sensitivity of coupled reaction systems that are constructed by differential equations with the numerous rate coefficients and then evolved to Sobol' sensitivity measure in @saltelli1999quantitative (the eFAST method with first- and total-order effects).
+The FAST algorithm creates the multidimensional space of input parameters through "search-curve".
+It's more computationally efficient in calculating the influence of model parameter than other global SA methods such as the Monte Carlo based approaches from @jansen1999analysis and @owen2013better.
 Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
+This approach can solve identifiability issue of PBPK model parameter that increases the computational burden of Bayesian analysis, and therefore reduce the computational efficiency.
 We also found a useful approach for communicating the parameter sensitivity in decision making.
 
-Most of sensitivity analysis tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} [@pianosi2015matlab], \proglang{python} [@herman2017salib], and \proglang{R} [@iooss2017introduction].
-Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package [@R-sensitivity], which is a practical tool to conduct local and global sensitivity analysis.
+Most of SA tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} [@pianosi2015matlab], \proglang{python} [@herman2017salib], and \proglang{R} [@iooss2017introduction].
+Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package [@R-sensitivity], which is a practical tool to conduct local and global SA.
 The \pkg{sensitivity} package includes some functions to generate the parameter sequences by using different computing algorithms.
@@
 This computational approach can effectively solve the heavy computing burden of numerical solutions within the pure \proglang{R} environment.
-Several other \proglang{R} packages, such as \pkg{FME}, \pkg{multisensi}, and \pkg{ODEsensitivity} include sensitivity analysis tools and are freely available in the Comprehensive R Archive Network (CRAN).
-The \pkg{FME} package contains a basic method for performing local sensitivity analysis for dynamic models [@JSSv033i03].
-The \pkg{multisensi} package provides sensitivity analysis for models with multivariate output [@R-multisensi].
-\pkg{ODEsensitivity} can perform sensitivity analysis in ordinary differential equation (ODE) models [@R-ODEsensitivity].
-It utilizes the ODE interface from the \pkg{deSolve} package and connect it with the sensitivity analysis from the \pkg{sensitivity} package.
+Several other \proglang{R} packages, such as \pkg{FME}, \pkg{multisensi}, and \pkg{ODEsensitivity} include SA tools and are freely available in the Comprehensive R Archive Network (CRAN).
+The \pkg{FME} package contains a basic method for performing local SA for dynamic models [@JSSv033i03].
+The \pkg{multisensi} package provides SA for models with multivariate output [@R-multisensi].
+\pkg{ODEsensitivity} can perform SA in ordinary differential equation (ODE) models [@R-ODEsensitivity].
+It utilizes the ODE interface from the \pkg{deSolve} package and connect it with the SA from the \pkg{sensitivity} package.
 Other related packages such as \pkg{mtk} [@RJ-2015-031] and \pkg{fast} [@R-fast] are designed to investigate the importance of model parameter.
-Although these tools provide various approaches to perform sensitivity analysis, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic and physiologically based kinetic (PK) modeling.
+Although these tools provide various approaches to perform SA, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic and PK modeling.
+The types of the kinetic model include the terms "pharmacokinetic", "toxicokinetic", and "biokinetic" model that are exchangeable in the model description.
 
-We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make sensitivity analysis more accessible and reproducible in pharmacological and toxicological researches.
+We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make SA more accessible and reproducible in pharmacological and toxicological researches.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic and PK model with multivariate model outputs.
+The analytical framework in \pkg{pkseni} not only include global SA but the uncertainty analysis and diagnostic tool to support the modeling process.
 The design concepts of \pkg{pksensi} are:
@@
 In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
-In addition, \pkg{pksensi} can link with R \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
+In addition, \pkg{pksensi} can link with \proglang{R} \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
 The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_instal} in \pkg{pksensi}.
@@
 
-```{r, eval=T}
+```{r, eval=F}
 Rtools.path <- "c:/Rtools/bin; c:/Rtools/mingw_64/bin"
@@
 
-We adopted eFAST, a widely used variance-based global sensitivity analysis approach in biomathematical modeling, especially for ODE-based dynamical models, in this package [@saltelli1999quantitative].
+We adopted eFAST, a widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimentional parameters in the numerical models, especially for ODE-based dynamical models, in this package.
 Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity].
-The eFAST method compute the sensitivity index (SI) via ANOVA-like decomposition of the function for analysis.
-This type of GSA approach and SI also know as Sobol method and Sobol index.
+The eFAST method computes the sensitivity index (SI) via ANOVA-like decomposition of the function for analysis.
+This type of global SA approach and SI also know as Sobol' method and Sobol' index.
 Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
-This can be written as:
+Considering the model that is built under $y=f(x_{i})$.
+The sampling scheme can be written as:
 
@@
 
-where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$  is a vector giving the set of frequencies, one frequency for each parameter.
-The default set of frequencies is based on the study from \cite{saltelli1999quantitative}.
+where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$ is a vector giving the set of frequencies, one frequency for each parameter, and $\varphi$ is a random phasei-shift coefficient ranged 0 to $2\pi$.
+The default set of frequencies is based on the study from @saltelli1999quantitative.
 Through the random phase-shift, the robust result of the sensitivity measurement should be similar across replications under the same sample size.
@@
 
-The uncertainty analysis is a crucial modeling process with sensitivity analysis.
+The uncertainty analysis is a crucial modeling process with SA.
 Sometimes, it's difficult to have informative data to set up the parameter distribution.
@@
 
-```{r workflow, echo = FALSE, out.width = "100%", fig.cap = '\\label{fig:Workflow}Workflow for performing uncertainty and sensitivity analysis in \\pkg{pksensi}.'}
+```{r workflow, echo = FALSE, out.width = "100%", fig.cap = '\\label{fig:Workflow}Workflow for performing uncertainty and sensitivity analysis.'}
 knitr::include_graphics("diagram/workflow.png")
@@
 
-The output of sensitivity analysis returns a list that contains the given time-points and values of all state variables (e.g., chemical concentrations in blood).
+The output of SA returns a list that contains the given time-points and values of all state variables (e.g., chemical concentrations in blood).
 This format is particularly suited for further graphical routines within \pkg{pksensi}.
@@
 
-Two examples of PK model were applied here to demonstrate how \pkg{pksensi} work.
+In this section, two types of PK model were applied here to demonstrate how \pkg{pksensi} work.
+The first one is generic model and the second is the chemical-specific model.
 The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current example.
@@
 
-In the first example, we use a simple, one-compartment PK model from \pkg{httk} package [@JSSv079i04] to demonstrate how \pkg{pksensi} can be applied to PK studies.
+In the first example, we use a generic, one-compartment PK model from \pkg{httk} package [@JSSv079i04] to demonstrate how \pkg{pksensi} can be applied to PK studies.
 The differential equations for this model is as:
@@
 
-where $A_{gutlumen}$ is the state variable that describes the quantity of compound in the gut lumen (mol) and $A_{rest}$ is the quantity of compound in rest of body and body blood (mol).
+where $A_{gutlumen}$ is the state variable that describes the quantity of compound in the gut lumen (mg) and $A_{rest}$ is the quantity of compound in rest of body and body blood (mg).
 The parameter $k_{gutabs}$ is the absorption rate constant that describes the chemical absorption from the gut lumen into gut tissue through first-order processes (/h) and $k_{elim}$ is the elimination rate constant (/h), which is equal to the total clearance divided by the volume of distribution.
@@
 
-where $V_{dist}$ is the volume of distribution (L/Kg BW) and $BW$ is the body weight (Kg).
-The $C_{rest}$ can also be seen as the chemical concentration in plasma that can be used to compare with observed results in a pharmacokinetic experiment (mol/L).
+where $V_{dist}$ is the volume of distribution (L/kg BW) and $BW$ is the body weight (kg).
+The $C_{rest}$ can also be seen as the chemical concentration in plasma (mg/L) that can be further used to compare with observed results in a pharmacokinetic experiment.
+The bioavailability is assumed to be 100% in this model.
 
@@
 Running model under \pkg{GNU MCSim} native code can have a faster speed to obtain the model outputs.
-This section demo how to conduct global sensitivity analysis with \pkg{deSolve} package.
+This section demo how to conduct global SA with \pkg{deSolve} package.
 Then compare the computer time with \pkg{MCSim}.
@@
 The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.
-The \code{AUC} is the calculated area under the curve of the rest body ($mol \cdot h/L$).
+The \code{AUC} is the calculated area under the curve of the rest body ($mg \cdot h/L$).
 
@@
 The code must first be compiled to run the model.
-After create the model file, use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
+After create the model file, we can use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
 
@@
 It can also schedule for the given dosing scenario.
-In the current setting, we assumed the initial condition of the intake chemical to be 10 mol.
-We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the following modeling and sensitivity analysis.
+In the current setting, we assumed the initial condition of the intake chemical to be 1000 mg.
+We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the following modeling and SA.
 These additional functions are generated when compiling the model file.
 
-We used the corresponding parameter value of theophylline in this example.
+We used the corresponding parameter value of acetaminophen in this example.
 The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
-The given value of \code{vdist}, \code{ke}, and \code{kgutabs} in \pkg{httk} are 0.78 (L/kg BW), 0.265 (/h), and 2.18 (/h), respectively.
-The body weight is assumed to be 60 (kg).
+The given value of \code{vdist}, \code{ke}, and \code{kgutabs} in \pkg{httk} are 1.1 (L/kg BW), 0.23 (/h), and 2.18 (/h), respectively.
+The body weight is assumed to be 70 (kg).
 
@@
 parms <- initParms()
-parms["vdist"] <- 0.78
-parms["ke"] <- 0.265
+parms["vdist"] <- 1.1
+parms["ke"] <- 0.23
 parms["kgutabs"] <- 2.18
-parms["BW"] <- 60
+parms["BW"] <- 70
 initState <- initStates(parms=parms)
@@
 
-To conduct a sensitivity analysis for the four parameters in one-compartment PK model, we quantify their impact on the chemical concentration in plasma during the 24-hr time period post intake.
+We conduct global SA for the four parameters in one-compartment PK model to investigate the parameter impact on the plasma concentration during the 24-hr time period post intake.
 The distribution of parameter is taken to be uniform with bounds corresponding to 50% and 200% of the nominal value.
-Therefore, the parameter ranges are assumed to be 0.39 and 1.56 L/kg BW for \code{vdist}.
-The \code{ke} are ranged from 0.133 to 0.53 /h, corresponding to half-times of 1.3 and 5.2 hr.
+Therefore, the parameter ranges are assumed to be 0.55 and 2.2 L/kg BW for \code{vdist}.
+The \code{ke} are ranged from 0.12 to 0.46 /h, corresponding to half-times of 1.5 and 5.8 hr.
 The \code{ka} are ranged 1.09 to 4.36 /h, corresponding to half-times of 0.32 and 0.64 hr.
-The Bodyweight is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
-The number of samples determines the robustness of the result of sensitivity analysis.
+The $BW$ is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
+The number of samples determines the robustness of the result of SA.
 Higher sample numbers sizes lead to narrower confidence intervals for sensitivity measurements across different replications.
 However, it will take a longer time in computation.
-Here we use a sample size of 400 with 20 replications:
+Here we use a sample size of 200 with 20 replications:
 
@@
 
-The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
+Through \code{rfast99} function, a S3 object with class 'rfast99' will be created. The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
 
@@
 
-Through \code{rfast99} function, S3 object with class 'rfast99' will be created.
 The generated parameters are stored as a 3-D array under the named \code{a} with the dimension of sample size, number of replication, and number of parameter, respectively.
@@
 
-The sample number was 800 with 4 model parameters, which generate 3,200 model evaluations. The replication was set to 20. Figure \ref{fig:rep-sampl} only plotted for 3 replications.
+The sample number was 200 with 4 model parameters, which generate 800 model evaluations.
+The replication was set to 20.
+Therefore, the totol of 16,000 parameter sequence will be used to compute the corresponding outputs.
+Figure \ref{fig:rep-sampl} plotted the sampling process for each parameter from the first 3 replications.
+The search curves show the different intensity of sampling patterns in each segment.
 
@@
 The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.
-The bar plot shows the distribution of sampling parameter'}
+The bar plot shows the distribution of sampling parameter.'}
 par(mfrow=c(4,4),mar=c(0,0,0,0),oma=c(3,3,2,1), pch =".")
@@
 The \code{print} function gives the sensitivity and convergence indices for main, interaction, and total order at each time point.
-In addition to print out the result of sensitivity analysis, the more efficient way to distinguish the influence of model parameter is to visualize them.
+In addition to print out the result of SA, the more efficient way to distinguish the influence of model parameter is to visualize them.
 The time-dependent sensitivity indices are shown in Figure \ref{fig:pk-solvefun-rest}.
@@
 The solid line represent the total (black) and first (red) order sensitivity index with 95\\% confidence interval (polygon).
-The dash line is the cut-off with default value of 0.05'}
+The dash line is the cut-off with default value of 0.05.'}
 plot(out)
@@
 
-Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} are dominating the plasma concentration in the before and after 5-hr post chemical intake, respectively, representing that the elimination is a key parameter to dominate the plasma concentration.
+Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} are dominating the plasma concentration before and after 5-hr post chemical intake, respectively.
 Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-The current result is only based on the setting distribution of model parameters for theophylline.
-The different given conditions will change the result (result not shown).
+However, the current result is only based on the setting distribution of model parameters for acetaminophen.
+The different given input conditions (e.g., range of parameter uncertainty, chemical dependent parameter value) can change the result (result not shown).
 
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.'}
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake..'}
 plot(out, vars = "Ametabolized")
@@
 
-The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is mainly determined by \code{ke}.
+The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is also determined by \code{ke}.
 Same as \code{Ccompartment}, the \code{kgutabs} contribute about 30 - 40% variation of model output in the first hour.
+The \code{BW} is the least important parameter in the current analysis, and therefore can be fixed in the model calibration and applications.
 
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:SI-output}The relationship between model parameter and estimated concentration at times 0.01, 2.01, 6.01, and 24.01 hr'}
+```{r, eval=T, fig.cap = '\\label{fig:SI-output}The relationship between model parameter and estimated concentration at times 0.01, 2.01, 6.01, and 24.01 hr (top to bottom), respectively.'}
 par(mfcol=c(4,4),mar=c(0.5,0.5,0,0),oma=c(3,3,2,1), pch = ".")
@@
 
-We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameter with significant relationship with model output have hight impact to dominate the modeling result.
+We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameters have high impact on the modeling result.
 The \code{ke} shows the high correlation at the lateing time (t = 24.01 h).
-The \code{ke} doesn't show the obvious relationship with \code{Ccompartment}.
+The \code{BW} doesn't show any obvious relationship with \code{Ccompartment}.
 
@@
 
-The \code{check} is a useful function to determine which parameters have relatively lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
+Some functions in \pkg{pksensi} provide the useful way to check the result from global SA.
+The \code{check} can determine which parameters have relatively lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
 The \code{check} also provides some feasible argument to specify the target output or change the cut-off value.
@@
 
-In addtion to apply \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can provide better computational efficiency.
+In addtion to use \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can provide better computational efficiency.
 To solve ODE through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
-The computing time of using \code{solve_fun} in sensitivity analysis is estimated as,
+The computing time of using \code{solve_fun} in SA is estimated as,
 
@@
 
-The following code is used to compile the \pkg{GNU MCSim} model code to the exexutable program.
+Then, before we conduct the SA through \pkg{GNU MCSim}, The following code is used to compile the \pkg{GNU MCSim} model code to the exexutable program.
 
@@
 
-Similar to \code{solve_fun} function that can define the initial parameter and state values through generated functions, the \code{solve_mcsim} has also a \code{condition} argument that can be used to give the specific input value such as oral dose, fixing parameter value or initial state variable.
+Similar to \code{solve_fun} function that can define the initial parameter and state variables' values through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as oral dose, fixed parameter value or initial condition of state variable.
 
@@
 
-After solving the equations under the same given condition, we can find that \pkg{GNU MCSim} has about 4 times faster in computing performance than using \pkg{deSolve}.
+After solving the equations under the same given condition, we can find that \pkg{GNU MCSim} has about 4 - 5 times faster in computing performance than using \pkg{deSolve}.
+In this case, we only focus on performing the global SA alone for generic PK model without additonal comparison with experiment data.
+The next example will display and reproduce our previous published result [@fphar201800588] with full global SA workflow.
 
@@
 
-The aim of this section is to reproduce our previous published result [@fphar201800588] of global sensitivity analysis for acetaminophen PBPK model through \pkg{pksensi}.
-The model codes are included in this package and can be generated through \code{pbpk_apap_model} function.
-We applied the global sensitivity analysis workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
+The aim of this section is to apply global SA for acetaminophen-PBPK model through \pkg{pksensi}.
+The model code is included in this package in https://github.com/nanhung/pksensi/blob/master/tests/pbpk_apap.model and can be generated through \code{pbpk_apap_model} function.
+We applied the global SA workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
 The descriptions of each parameter and the sampling ranges are list in Table \ref{tab:data}.
@@
 
-print(xtable(df, caption = "\\label{tab:data}Description of sampling range of model parameter"),
+print(xtable(df, caption = "\\label{tab:data}Description of sampling range of model parameter."),
       type = "latex", 
-      caption.placement = "top", include.rownames = FALSE,
+      caption.placement = "bottom", include.rownames = FALSE,
       sanitize.text.function = function(x) {x})
@@
 ```{r, eval=T}
-# Nominal value
 Tg <- log(0.23)
@@
 
-Generally, wide ranges of parameter values might cause the computational error in the solver.
+Generally, wide range of parameter value might cause the computating error when the solving the ODE.
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
-The \code{generate_infile} and \code{solve_mcsim} provide the arguments of \code{rtol} and \code{atol} that can be adjusted to prevent the unwanted error.
+The \code{generate_infile} and \code{solve_mcsim} provide the arguments of \code{rtol} and \code{atol} that adjust the error tolerance to prevent the unwanted error.
 However, the modification will lower computational speed.
 Therefore, the alternative method to prevent the computational errors is to detect the crucial parameter range that causes the problem.
-Also, setting the maximum number of (internally defined) steps to higher value instead of using the default value (500) can prevent this problem.
+Also, setting the maximum number of (internally defined) steps to higher value instead of using the default value (500) in \pkg{GNU MCSim} can prevent this problem.
 The maximum number of step is set to 5000 in this case.
-Here we seperate the GSA of APAP-PBPK model process to several steps.
+Here we seperate the global SA of APAP-PBPK model process to several steps.
 
-### Construct and compile the model file
+### Prepare and compile the model file
 
-The model code needs to be prepared in the following GSA workflow.
+The model code needs to be prepared in the following global SA workflow.
 After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through \code{compile_model}.
@@
 To prevent the computational error, we adjusted the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} from $U(0, 15)$ to $U(0, 10)$ in this case.
-The objects \code{q} and \code{dist} are used to set the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for sensitivity analysis).
+The objects \code{q} and \code{dist} are used to set the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
 
@@
 
-To improve computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (AG) and sulfate (AS) in plasma.
+To improve computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (APAP-G) and sulfate (APAP-S) in plasma.
 The setting oral dose of APAP is 20 mg/kg in this example.
@@
 
-In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the computational error.
-The Monte Carlo simulation is run fun for 1000 iteration as the assignment of \code{monte_carlo}.
+In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the computating error.
+The Monte Carlo simulation is run for 1000 iteration as the assignment of \code{monte_carlo}.
 The input file ('sim.in') and output file ('simmc.out') will be generated under the standard ASCII format.
@@
 
-Figure \ref{fig:pksim-MC} shows the result of the uncertainty analysis of APAP data and modeling results.
+Figure \ref{fig:pksim-MC} shows the uncertainty analysis of APAP data and modeling results.
 For parent compound, all data points are located in the simulated interval of 25-75%.
 Through this result, we can determine that the simulated outputs can accurately generate the same concentration profile as the in-vivo experiment under the setting of parameter ranges for APAP.
-The simulated result of metabolites APAP-G shows the different PK profile with experiment data.
+The simulated result of metabolites APAP-G shows the different pharmacokinetic profile with experiment data.
 However, all data points are located in the simulated interval.
@@
 
-In GSA, we have to additionally generate the parameter matrix from the eFAST method. The current setting uses 512 sample size with 10 replication.
+In global SA, we have to additionally generate the parameter matrix from the eFAST method. The current setting uses 512 sample size with 10 replication.
 
@@
 
-### Conduct the GSA
+### Conduct the global SA
 
-To conduct the GSA with \pkg{GNU MCsim} and \pkg{pksensi}, the input file with given "Setpoint" condition should be generated before modeling.
+To conduct the global SA with \pkg{GNU MCsim} and \pkg{pksensi}, the input file with given "Setpoint" condition should be generated before modeling.
 The file can create by \code{generate_infile} function.
-The \code{solve_mcsim} can also automatically the input file and compute the output.
+The \code{solve_mcsim} can also automatically create the input file and compute the output.
 
 ```{r, eval=T}
-system.time(out <- solve_mcsim(x, mName = mName,
+out <- solve_mcsim(x, mName = mName,
                    params = params, 
@@
                    condition = conditions, 
-                               rtol = 1e-7, atol = 1e-9))
+                   rtol = 1e-7, atol = 1e-9)
 ```
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:Heat-SI}Heat map representation of time-dependent total SI computed for GSA method', fig.height=5}
+```{r, eval=T, fig.cap = '\\label{fig:Heat-SI}Heat map representation of time-dependent total SI computed for global SA method', fig.height=5}
 heat_check(out, order = "total order")
@@
 
-Adding the `index = "CI"` in the function can further investigate the convergence of the sensitivity index.
+Adding the `index = "CI"` in the function can further investigate the convergence index.
 Based on the current setting of sampling size, most parameters cannot reach the acceptable criteria of convergence (Figure \ref{fig:Heat-CI}).
@@
 The sample size of convergence in the current PBPK model is 8,192 [@fphar201800588].
+However, based on the current sample size we still can find 6 parameters that can be an influential parameter for the plasma APAP concentration.
 
-```{r, eval=T, fig.cap = '\\label{fig:Heat-CI}Heat map representation of time-dependent total CI computed for GSA method', fig.height=5}
+```{r, eval=T, fig.cap = '\\label{fig:Heat-CI}Heat map representation of time-dependent total CI computed for global SA method', fig.height=5}
 heat_check(out, index = "CI", order = "total order")
@@
 
-The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct various dynamical simulations for biological-based modeling, and it's mainly built for pharmacokinetic research with probabilistic approaches [@bois2009gnu].
+Computational modeling has become an important technique in modern scientific research to support the hypothesis, experiment, and discoveries.
+Uncertainty and SA is essential knowledge for modelers to conduct the model development and examination.
+More computational tools had been developed for various types of modeling and simulation.
+The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct dynamical simulation for biological-based modeling, and it's mainly built for PK research with probabilistic approaches [@bois2009gnu].
 The \proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
+We initiated \pkg{pksensi} that aim to conduct comprehensive and efficient global SA workflow in PK modeling, which hasn't been developed for open source communities.
 The current version had included \pkg{GNU MCSim} and \pkg{deSolve} as the main solver in our model assessment.
-Compare with \pkg{deSolve} package, we found that The \pkg{GNU MCSim} can provide better efficiency to solve the ODE in PK model.
-Some developed package in \proglang{R} had also provide the feature to solve ODE such as \pkg{RxODE} and \pkg{mrgsolve} [@wang2016tutorial; @R-mrgsolve].
-Therefore, we hope to incorporate this package in our GSA workflow to provide user more options in the modeling.
+Compare with \pkg{deSolve} package, we found that The \pkg{GNU MCSim} can provide better efficiency to solve the ODE in PK models.
+Although \pkg{pksensi} provide the features and functions to install and link with \pkg{GNU MCSim}, conducting the uncertainty and SA in PK modeling, the overall learning curve of this workflow might steep for users that are not familiar with \pkg{GNU MCSim} from model development, debugging, and testing.
+Therefore, the general \proglang{R} package, such as \pkg{deSolve}, can be an alternative option for \proglang{R} user.
+Some developed package in \proglang{R} had also provide the features to solve ODE such as \pkg{RxODE} [@wang2016tutorial] and \pkg{mrgsolve} [@R-mrgsolve].
+We hope to incorporate these packages that can improve the user experience in using \pkg{pksensi} in the future.
 
-Moreover, some fundamental packages on the CRAN repository and other software platforms had been built to perform different types of GSA.
-The development of \pkg{pksensi} is mainly for PK models that are constructed by differential equations.
-It provides a straightforward application to investigate the impact of parameters on model output across time and variables.
-We adopted the eFSAT method, a common variance-based sensitivity analysis that had better efficiency and accuracy in the model assessment.
+Moreover, some fundamental packages on the CRAN repository and other software platforms had been built to perform global SA with different algorithms.
+The development of \pkg{pksensi} is mainly for PK models that are constructed by ODE.
+It provides a straightforward application to investigate the impact of parameters on model output across simulation time-points and variables.
+We adopted the eFSAT method, a common variance-based SA that had better efficiency and accuracy in the model assessment.
 It can exam the parameter effect from first order to the total order.
-In addition, the package can not only conduct GSA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most GSA studies and its software package.
-We also develop the function to visualize the output result that can help us determine the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model Calibration.
-In addition to eFAST, Some studies applied another GSA algorithm in the physiologically and biologically based ODE models, such as compartmental absorption and transit (also known as CAT) and other system biology models [@melillo2019variance; @sumner2012methodology].
-In future version, we expect to add additional GSA approach that had been applied in advance study.
-The other features such as principle component analysis [@lamboni2011multivariate] were considered to include in \pkg{pksensi}.
+In addition, the package can not only conduct global SA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most global SA studies and its software package.
+We also develop the function to visualize the output result that can help us distinguish the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model Calibration.
+In addition to eFAST, Some studies applied another global SA algorithm in the physiologically and biologically based ODE models, such as compartmental absorption and transit (also known as CAT) and other system biology models [@melillo2019variance; @sumner2012methodology].
+Base on our knowledge, the eFAST is a least computational expensive approch than other Monte-Carlo based Sobol' method.
+In future version, we expect to add additional global SA approach that had been applied in advanced studies [@glen2012estimating; @ring2017identifying].
+The other features such as principle component analysis [@lamboni2011multivariate] were considered to further include in our global SA workflow.
 
-It is worth noting that the GSA results are limited to the specific model and the given range of the model parameter.
+It is worth noting that the global SA results are limited to the specific model and the given range of the model parameter.
 The provided conditions, such as the number of examing parameter, parameter distribution, parameter range, and the given dosing level in PK model might have a potential impact on the level of importance of each parameter.
@@
 In other words, if we have various types of experiment data (e.g.
-different exposure routes and biological measurements) and will use them in model calibration, it will enhance the difficulties in GSA.
-The complex model with a high dimension of the parameter is also a challenge in GSA, which will need higher computer demand in the model assessment.
-Overall,
-Overall, \pkg{pksensi} improves the user experience of performing GSA and can create robust and reproducible results for decision making in pharmacokinetic model calibration.
+different exposure routes and biological measurements) and will use them in model calibration, it will enhance the difficulties in global SA.
+The complex model with a high dimension of the parameter is also a challenge in global SA, which will need higher computer demand in the model assessment.
 More additional applications still need to be investigated and strengthen.
+Overall, \pkg{pksensi} provide an additional option of performing global SA for PK model and can create robust and reproducible results for decision making in model calibration.
 

### 2019-06-28 manuscript.Rmd
manuscript

@@
 chemical concentration in blood or tissues) in model responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
-This technique can be guidance of experimental design, parameter estimation during the model establishment, and qualification processes [@zhang2015sobol].
+In addition, this SA approach can be guidance of experimental design, parameter estimation during the model establishment, and qualification processes [@zhang2015sobol].
 Through SA, it can optimal experimental design and distinguish the "unidentifiable" parameters that can lead to problems with numerical procedures used in parameter estimation [@chu2010quantitative].
@@
 
-In our earlier work [@fphar201800588], we developed an approach to apply global SA workflow to reduce the computational burden in the Bayesian, Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+In our earlier work [@fphar201800588], we developed a workflow to apply global SA to reduce the computational burden in the Bayesian, Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
 We used \pkg{GNU MCSim} [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
@@
 This approach can solve identifiability issue of PBPK model parameter that increases the computational burden of Bayesian analysis, and therefore reduce the computational efficiency.
-We also found a useful approach for communicating the parameter sensitivity in decision making.
+We also found a useful features for communicating the parameter sensitivity in decision making.
 
 Most of SA tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} [@pianosi2015matlab], \proglang{python} [@herman2017salib], and \proglang{R} [@iooss2017introduction].
-Previously, we mainly used \proglang{R} with the \pkg{sensitivity} package [@R-sensitivity], which is a practical tool to conduct local and global SA.
+The \proglang{R} software is widely used and had numerious developed packages, such as \pkg{sensitivity} [@R-sensitivity], which is a practical tool to conduct local and global SA.
 The \pkg{sensitivity} package includes some functions to generate the parameter sequences by using different computing algorithms.
@@
 The \pkg{FME} package contains a basic method for performing local SA for dynamic models [@JSSv033i03].
+It can also conduct collinearity and identifiability analysis to evaluate which model parameters can be estimated or exclude based on available observation data.
 The \pkg{multisensi} package provides SA for models with multivariate output [@R-multisensi].

### 2019-06-29 manuscript.Rmd
manuscript

@@
 title:
-  formatted: "\\pkg{pksensi}: An R package to apply global sensitivity analysis in physiologically based kinetic models"
+  formatted: "\\pkg{pksensi}: R package to apply global sensitivity analysis in physiologically based kinetic models"
   # If you use tex in the formatted title, also supply version without
-  plain:     "pksensi: An R package to apply global sensitivity analysis in physiologically based kinetic models"
+  plain:     "pksensi: R package to apply global sensitivity analysis in physiologically based kinetic models"
   # For running headers, if needed
-  short:     "**pksensi**: Sensitivity analysis for pharmacokinetic models in R"
+  short:     "**pksensi**: Global SA for PK models in R"
 abstract: >
@@
 Other related packages such as \pkg{mtk} [@RJ-2015-031] and \pkg{fast} [@R-fast] are designed to investigate the importance of model parameter.
+
 Although these tools provide various approaches to perform SA, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic and PK modeling.
-The types of the kinetic model include the terms "pharmacokinetic", "toxicokinetic", and "biokinetic" model that are exchangeable in the model description.
-
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make SA more accessible and reproducible in pharmacological and toxicological researches.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic and PK model with multivariate model outputs.
+The types of the kinetic model include the terms "pharmacokinetic", "toxicokinetic", and "biokinetic" model that are exchangeable in the model description.
 The analytical framework in \pkg{pkseni} not only include global SA but the uncertainty analysis and diagnostic tool to support the modeling process.
@@
 This type of global SA approach and SI also know as Sobol' method and Sobol' index.
-Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicate sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
+Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicating sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
 Considering the model that is built under $y=f(x_{i})$.
@@
 
-where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$ is a vector giving the set of frequencies, one frequency for each parameter, and $\varphi$ is a random phasei-shift coefficient ranged 0 to $2\pi$.
+where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$ is a vector giving the set of frequencies, one frequency for each parameter, and $\varphi$ is a random phasei-shift coefficient ranged from 0 to $2\pi$.
 The default set of frequencies is based on the study from @saltelli1999quantitative.
@@
 
-To investigate the convergence of sensitivity indices, we adopted the "exam" approach proposed by [@sarrazin2016global].
-This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed sensitivity index for all parameters across all time-points and output variables.
-Through random phase shift procedure, the sensitivity index can repeatly generate with low and high variation based on the given sampling size.
-Therefore, they can be used to investigate the convergence of sensitivity indices as
+To investigate the convergence of SI, we adopted the "exam" approach proposed by [@sarrazin2016global].
+This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed SI for all parameters across all time-points and output variables.
+Through random phase shift procedure, the SI can repeatly generate with low and high variation based on the given sampling size.
+Therefore, they can be used to investigate the convergence of SI as
 
@@
 
-where $SI_{i,t}^{ub}$ and $SI_{i,t}^{ub}$ are the upper and lower bound of the sensitivity index from the specific model output ($i$) at time ($t$). $CI_{i,t}$ is the corresponding convergence index.
+where $SI_{i,t}^{ub}$ and $SI_{i,t}^{ub}$ are the upper and lower bound of the SI from the specific model output ($i$) at time ($t$). $CI_{i,t}$ is the corresponding convergence index.
 
 To integrate the random phase-shift with eFAST, we built the \code{rfast99} function, which can be used to perform the sensitivity test for the whole modeling process.
-The function \code{rfast99} in \pkg{pksensi} can sample and generate the testing parameter matrix based on a given argument of sample size n and probability distribution (e.g., mean/s.d.
+The function \code{rfast99} in \pkg{pksensi} can sample and generate the testing parameter matrix based on a given argument of sample size $n$ and probability distribution (e.g., mean/s.d.
 of normal distribution, meanlog/sdlog of log-normal distribution, and minimum/maximum of uniform distribution).
@@
 The \pkg{pksensi} includes a function to compile and create dynamic-link libraries (.dll) on Windows and shared objects (.so) on Unix-liked systems (e.g., Linux and MacOS).
-This compiled file can load and execute through the build-in function in \proglang{R}.
+This compiled program can load and execute through the build-in function in \proglang{R}.
 However, Windows users need to install Rtools to compile the source code by using the GNU GCC compiler.
-More requirements are detailed in the study from [@soetaertr].
-The \proglang{C} implementation of the example model can be found as an example within \pkg{GNU MCSim} v6.0.1 [@JSSv002i09].
-The \pkg{pksensi} can also link with \proglang{GNU MCSim} to compile the model code, used in solving each system of equations.
+More requirements are detailed in @soetaertr.
+The \proglang{C} implementation of the example model can be found as an example within \pkg{GNU MCSim} [@JSSv002i09].
+The \pkg{pksensi} can also link with \pkg{GNU MCSim} to create the model program, used in solving each system of equations.
 
@@
 Sometimes, it's difficult to have informative data to set up the parameter distribution.
-The wider range of parameter distribution might cause the computation error at the extreme value.
+The wider range of parameter distribution might cause the computing error at the extreme value.
 The narrower range might cause the incorrect calibrated result of the model parameter.
-To prevent this problem.
-The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process.
-Before the model calibration, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.
-\pkg{pksensi} includes the function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
-The general workflow of uncertainty and sensitivityanalysis is illustrated in Figure \ref{fig:Workflow}.
+The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process to prevent this problem.
+Before fitting the data to model, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.
+\pkg{pksensi} also has a function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
+The general workflow of uncertainty and sensitivity analysis is illustrated in Figure \ref{fig:Workflow}.
 
@@
 This format is particularly suited for further graphical routines within \pkg{pksensi}.
-Also, time-dependent sensitivity measurements of first and total order effect for all parameters can be plotted simultaneously.
-The sensitivity index might have different estimation across replications under the same sample size.
-The inadequate sample number can create an unreliable result and bias.
+Also, time-dependent sensitivity measurements of first and total order effects for all parameters can be plotted simultaneously.
+The computed SI might show different estimation across replications under the same sample size.
+Therefore, the inadequate sample number can create an unreliable result and bias.
 It's essential to use the sufficient sample size to prevent incorrect judgment in parameter fixing.
-Therefore, \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
+\pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
 We also developed a "cut-off"-based approach in this package to accurately distinguish between "influential" and "non-influential" parameters.
-Finally, \pkg{pksensi} includes numerous visualization tools for the effective investigation and communication of results in decision support.
+Finally, \pkg{pksensi} provides visualization tools for the effective investigation and communication of results in decision support.
 
@@
 The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current example.
-The version of \pkg{GNU MCSim} is 6.1.0.
+We use \pkg{GNU MCSim} v6.1.0 in these examples.
 
@@
 
-where $A_{gutlumen}$ is the state variable that describes the quantity of compound in the gut lumen (mg) and $A_{rest}$ is the quantity of compound in rest of body and body blood (mg).
+where $A_{gutlumen}$ is the state variable that describes the quantity of compound in the gut lumen (mg) and $A_{rest}$ is the quantity of compound in rest of body and blood (mg).
 The parameter $k_{gutabs}$ is the absorption rate constant that describes the chemical absorption from the gut lumen into gut tissue through first-order processes (/h) and $k_{elim}$ is the elimination rate constant (/h), which is equal to the total clearance divided by the volume of distribution.
@@
 
-The concentration of the chemical in the rest of body blood $C_{rest}$ can be calculated as
+The concentration of the chemical in the rest of body and blood ($C_{rest}$, mg/L) can be calculated as
 
@@
 where $V_{dist}$ is the volume of distribution (L/kg BW) and $BW$ is the body weight (kg).
-The $C_{rest}$ can also be seen as the chemical concentration in plasma (mg/L) that can be further used to compare with observed results in a pharmacokinetic experiment.
+The $C_{rest}$ can also be seen as the chemical concentration in plasma that can be further used to compare with observed results in a pharmacokinetic experiment.
 The bioavailability is assumed to be 100% in this model.
@@
 Running model under \pkg{GNU MCSim} native code can have a faster speed to obtain the model outputs.
-This section demo how to conduct global SA with \pkg{deSolve} package.
-Then compare the computer time with \pkg{MCSim}.
+This section mainly focus on how to conduct global SA with \pkg{deSolve} package.
+Then compare the computing time with \pkg{MCSim}.
 
@@
 mName <- "pbtk1cpt"
-compile_model(mName, application = "R", version = "6.1.0")
+compile_model(mName, application = "R")
 ```
@@
 It can also schedule for the given dosing scenario.
-In the current setting, we assumed the initial condition of the intake chemical to be 1000 mg.
+In the current setting, we assumed the initial condition of the intake dose to be 1000 mg.
 We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the following modeling and SA.
@@
 
-We used the corresponding parameter value of acetaminophen in this example.
-The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical and drug biological properties for over 500 chemicals.
+We used the corresponding parameter value of  (APAP) in this example.
+The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical properties for over 500 chemicals.
 The given value of \code{vdist}, \code{ke}, and \code{kgutabs} in \pkg{httk} are 1.1 (L/kg BW), 0.23 (/h), and 2.18 (/h), respectively.
@@
 
-Using the \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic profile (Figure \ref{fig:ode-PK}) according to the given parameter conditions and the time points (\code{t}):
+Using the \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic profile (Figure \ref{fig:ode-PK}) according to the given parameter baseline and the time points (\code{t}):
 
@@
 The $BW$ is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
-The number of samples determines the robustness of the result of SA.
-Higher sample numbers sizes lead to narrower confidence intervals for sensitivity measurements across different replications.
+The number of sample size determines the robustness of the result of SA.
+Higher number of sample size lead to narrower confidence intervals for sensitivity measurements across different replications.
 However, it will take a longer time in computation.
-Here we use a sample size of 200 with 20 replications:
+Here we use a sample size of 200 with 20 replications.
 
@@
 
-The generated parameters are stored as a 3-D array under the named \code{a} with the dimension of sample size, number of replication, and number of parameter, respectively.
+The generated parameters are stored as a 3-D array under the named \code{a}, with the dimension of sample size, number of replication, and number of parameter, respectively.
 
@@
 
-The sample number was 200 with 4 model parameters, which generate 800 model evaluations.
-The replication was set to 20.
+The sample number is 200 with 4 model parameters, which generates 800 model evaluations.
+The replication is set to 20.
 Therefore, the totol of 16,000 parameter sequence will be used to compute the corresponding outputs.
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:rep-sampl}The parameter search curve generated by eFAST method.
+```{r, eval=T, fig.height=5, fig.cap = '\\label{fig:rep-sampl}The parameter search curve generated by eFAST method.
 The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.
 The bar plot shows the distribution of sampling parameter.'}
-par(mfrow=c(4,4),mar=c(0,0,0,0),oma=c(3,3,2,1), pch =".")
+par(mfrow=c(4,4),mar=c(0.8,0.8,0.8,0),oma=c(4,4,2,1), pch =".")
 for (j in c("vdist", "ke", "kgutabs", "BW")) {
   if ( j == "BW") {
-    plot(x$a[,1,j], ylab = "")
+    plot(x$a[,1,j], ylab = "BW")
   } else plot(x$a[,1,j], xaxt="n", ylab = "")
@@
                breaks=seq(from=min(x$a[,,j]), to=max(x$a[,,j]), length.out=20))
-  barplot(hist$density, axes=FALSE, space=0, horiz = T) 
+  barplot(hist$density, axes=FALSE, space=0, horiz = T, main = j) 
 }
+mtext("Model evaluation", SOUTH<-1, line=2, outer=TRUE)
 ```
@@
 Here we define the output time points (\code{t}) to examine the change of the parameter sensitivity over time.
-To solve the model through \pkg{deSolve}, we need to provide the details of the argument:
+To solve the model through \pkg{deSolve}, we need to provide the details of the argument, which include time (\code{t}), initial  conditions of state variable (\code{initState}), output variables (\code{outnames}), and name of the shared library (\code{mName}, without extension).
+To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours with the time segment of 1 hour as the above definition in \code{ode} function in this example.
+The initial time point should avoid 0 to prevent computational error in misconduct.
+The \code{outnames}, \code{dllname}, are based on the arguments from the \code{ode} function in \pkg{deSolve} package.
+The details of the model structure and these arguments are defined in \code{pbtk1comp.c}.
+and \code{pbtk1comp_inits.R} files.
 
@@
 
-To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours with the time segment of 1 hour as the above definition in \code{ode} function in this example.
-The initial time point should avoid 0 to prevent computational error in misconduct.
-The \code{outnames}, \code{dllname}, are based on the arguments from the \code{ode} function in \pkg{deSolve} package.
-The details of the model structure and these arguments are defined in \code{pbtk1comp.c}.
-and \code{pbtk1comp_inits.R} files.
-The \code{out} is an S3 object of \code{rfast99} as well, which can link with \code{print}, \code{plot}, and \code{check} method to examine the sensitivity measurements.
+The output result \code{out} is an S3 object of \code{rfast99} as well, which can link with \code{print}, \code{plot}, and \code{check} method to examine the sensitivity measurements.
 The \code{print} function gives the sensitivity and convergence indices for main, interaction, and total order at each time point.
-In addition to print out the result of SA, the more efficient way to distinguish the influence of model parameter is to visualize them.
-The time-dependent sensitivity indices are shown in Figure \ref{fig:pk-solvefun-rest}.
+In addition to print out the result of SA, the more efficient way to distinguish the influence of model parameter is to visualize these indices.
+The time-dependent SI are shown in Figure \ref{fig:pk-solvefun-rest}.
+The SI has computed range from 0 (no impact) to 1 (high impact) and represent the contribution percentage of output variance under the given parameter distributions.
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-rest}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.
-The solid line represent the total (black) and first (red) order sensitivity index with 95\\% confidence interval (polygon).
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-rest}Time-dependent SI of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.
+The solid line represent the total (black) and first (red) order SI with 95\\% confidence interval (polygon).
 The dash line is the cut-off with default value of 0.05.'}
@@
 Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-However, the current result is only based on the setting distribution of model parameters for acetaminophen.
+However, the current result is only based on the setting distribution of model parameters for APAP.
 The different given input conditions (e.g., range of parameter uncertainty, chemical dependent parameter value) can change the result (result not shown).
@@
 The default output in the plotting is setting at the first variable.
-To exam the time-sependent sensitivity index of other variables, such as \code{Ametabolized} in this case, we need to assign the variable name \code{vars = "Ametabolized"} in \code{plot} function (Figure \ref{fig:pk-solvefun-Amet}).
+To exam the time-sependent SI of other variables, such as \code{Ametabolized} in this case, we need to assign the variable name \code{vars = "Ametabolized"} in \code{plot} function (Figure \ref{fig:pk-solvefun-Amet}).
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent sensitivity indices of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake..'}
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent SI of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake..'}
 plot(out, vars = "Ametabolized")
@@
 
-The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is also determined by \code{ke}.
+The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is also determined by parameter \code{ke}.
 Same as \code{Ccompartment}, the \code{kgutabs} contribute about 30 - 40% variation of model output in the first hour.
-The \code{BW} is the least important parameter in the current analysis, and therefore can be fixed in the model calibration and applications.
+The \code{BW} is the least important parameter in the current analysis, and therefore can be fixed in the model fitting to data and additional applications.
 
-In addition to use time-sI profile to investigate the parameter impact on model output, we can construct the relationship between parameter and the corresponding value of model output in the examination.
-Figure \ref{fig:SI-output} show the concentration of rest body and the model parameters:
+In addition to use time-SI profile to investigate the parameter impact on model output, we can construct the relationship between parameter and the corresponding value of model output in the examination.
 
 ```{r, eval=T, fig.cap = '\\label{fig:SI-output}The relationship between model parameter and estimated concentration at times 0.01, 2.01, 6.01, and 24.01 hr (top to bottom), respectively.'}
-par(mfcol=c(4,4),mar=c(0.5,0.5,0,0),oma=c(3,3,2,1), pch = ".")
+par(mfcol=c(4,4),mar=c(0.8,0.8,0,0),oma=c(4,4,2,1), pch = ".")
 plot(x$a[,1,"vdist"], out$y[,1,"0.01",1], xaxt="n", main = "\nvdist")
@@
 }
-mtext("Parameter", SOUTH<-1, line=1.8, outer=TRUE)
-mtext("Ccompartment", WEST<-2, line=1.8, outer=TRUE)
+mtext("Parameter", SOUTH<-1, line=2, outer=TRUE)
+mtext("Ccompartment", WEST<-2, line=2, outer=TRUE)
 ```
 
+Figure \ref{fig:SI-output} show the concentration of rest body (\code{Ccompartment}) and the model parameters.
 We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameters have high impact on the modeling result.
-The \code{ke} shows the high correlation at the lateing time (t = 24.01 h).
+The \code{ke} shows the high correlation at the later time period (t = 24.01 h).
 The \code{BW} doesn't show any obvious relationship with \code{Ccompartment}.
 
-The output variable \code{out} containing all the input arguments detailed before and the calculated sensitivity indices of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
+The output variable \code{out} containing all the input arguments detailed before and the calculated SI of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
 The convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
-The output are formated as 4-D array in \code{y} with the dimension name of model evaluation, number of replication, number of time points, and number of variable, respectively.
+The output are formated as 4-D array in \code{y} with the dimension name of model evaluation, number of replications, number of time points, and number of output variables, respectively.
 
@@
 
-Some functions in \pkg{pksensi} provide the useful way to check the result from global SA.
-The \code{check} can determine which parameters have relatively lower sensitivity measurement across the given time interval, and therefore can be applied parameter fixing in model calibration.
+Some functions in \pkg{pksensi} provide a efficient way to check the result from global SA.
+The \code{check} can determine which parameters have relatively lower sensitivity measurement across the given time points and model outputs, and therefore can be applied parameter fixing in model calibration.
 The \code{check} also provides some feasible argument to specify the target output or change the cut-off value.
@@
 
-Based on the sensitivity measurement of the total order, the result shows that \code{BW} has a relative lower measurement of sensitivity index.
+Based on the sensitivity measurement of the total order, the result shows that \code{BW} has a relative lower measurement of SI.
 However, all parameters do not converge to the setting cut-off, which means the larger sample size is required in further sensitivity testing.
@@
 ```{r, eval=T, message=F, warning=F}
-compile_model(mName, application = "mcsim", version = "6.1.0")
+compile_model(mName, application = "mcsim")
 ```
 
-Similar to \code{solve_fun} function that can define the initial parameter and state variables' values through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as oral dose, fixed parameter value or initial condition of state variable.
+Similar to \code{solve_fun} function that can define the initial value of parameter and state variable through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as exposure dose, fixed parameter value or initial condition of state variable.
 
@@
 
-The aim of this section is to apply global SA for acetaminophen-PBPK model through \pkg{pksensi}.
+The aim of this section is to apply global SA for APAP-PBPK model through \pkg{pksensi}.
 The model code is included in this package in https://github.com/nanhung/pksensi/blob/master/tests/pbpk_apap.model and can be generated through \code{pbpk_apap_model} function.
@@
 
-print(xtable(df, caption = "\\label{tab:data}Description of sampling range of model parameter."),
+print(xtable(df, center = "centering", caption = "\\label{tab:data}Description of sampling range of model parameter."), 
       type = "latex", 
@@
 
-Generally, wide range of parameter value might cause the computating error when the solving the ODE.
+Generally, wide range of parameter value might cause the computing error when solving the differentiial equation.
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
 The \code{generate_infile} and \code{solve_mcsim} provide the arguments of \code{rtol} and \code{atol} that adjust the error tolerance to prevent the unwanted error.
-However, the modification will lower computational speed.
-Therefore, the alternative method to prevent the computational errors is to detect the crucial parameter range that causes the problem.
-Also, setting the maximum number of (internally defined) steps to higher value instead of using the default value (500) in \pkg{GNU MCSim} can prevent this problem.
+However, the modification will decrease the computing speed.
+Therefore, the alternative method to prevent this issue is to detect the crucial parameter range that causes the problem.
+Also, setting the maximum number of steps to higher value instead of using the default value (500) in \pkg{GNU MCSim} can prevent this problem (internally defined).
 The maximum number of step is set to 5000 in this case.
 Here we seperate the global SA of APAP-PBPK model process to several steps.
+The value can reset through reinstall the the \pkg{GNU MCSim} by using \code{mcsim_install(mxstp = 5000)}.
 
@@
 The model code needs to be prepared in the following global SA workflow.
-After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through \code{compile_model}.
+After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through \code{compile_model} function.
 
@@
 pbpk_apap_model()
-compile_model(mName, application = "mcsim", version = "6.1.0")
+compile_model(mName, application = "mcsim")
 ```
@@
 
-The 21 testing model parameters are defined in this part, including parameter name, probability distribution, and distribution parameter.
-To prevent the computational error, we adjusted the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} from $U(0, 15)$ to $U(0, 10)$ in this case.
-The objects \code{q} and \code{dist} are used to set the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
+The 21 testing model parameters are defined in this part, including parameter name, probability distribution, and distributed parameter value.
+To prevent the computing error, the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} need to adjust from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
+The objects \code{q} and \code{dist} are set to the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
 
@@
 
-To improve computational speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of acetaminophen (APAP) and its metabolites glucuronide (APAP-G) and sulfate (APAP-S) in plasma.
+To optimize the computing speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of APAP and its metabolites glucuronide (APAP-G) and sulfate (APAP-S) in plasma.
 The setting oral dose of APAP is 20 mg/kg in this example.
@@
 
-We apply uncertainty analysis through the \code{solve_mcsim} and visualize the result by \code{pksim}.
-Some example APAP PK data are included in the \pkg{pksensi} with experiment time (h) and concentration (mg/L).
+We apply uncertainty analysis through the \code{solve_mcsim} and visualize the result by \code{pksim} function.
+Some example data are included in the \pkg{pksensi} with experiment time (h) and concentration (mg/L).
 
@@
 
-```{r, fig.cap = '\\label{fig:pksim-MC}Coverage checks of prior PBPK model predictions with calibrated APAP data', fig.height=3}
+```{r, fig.cap = '\\label{fig:pksim-MC}Coverage checks of prior PBPK model predictions with calibrated APAP data', fig.height=3.5, fig.width=9}
 par(mfrow = c(1,3), mar = c(4,4,1,1))
@@
 
-To conduct the global SA with \pkg{GNU MCsim} and \pkg{pksensi}, the input file with given "Setpoint" condition should be generated before modeling.
+To conduct the global SA with \pkg{GNU MCsim} and \pkg{pksensi}, the input file with given "setpoint" condition should be generated before modeling.
 The file can create by \code{generate_infile} function.
@@
 
-```{r, eval=T, fig.height=8, fig.width=8, fig.cap = '\\label{fig:APAP-sensi}Time-dependent sensitivity indices of the plasma APAP concentration estimated from APAP PBPK model during 12-hr time period post intake.'}
+```{r, eval=T, fig.height=8, fig.width=8, fig.cap = '\\label{fig:APAP-sensi}Time-dependent SI of the plasma APAP concentration estimated from APAP-PBPK model during 12-hr time period post intake.'}
 plot(out, vars = "lnCPL_APAP_mcgL")
@@
 The sample size of convergence in the current PBPK model is 8,192 [@fphar201800588].
-However, based on the current sample size we still can find 6 parameters that can be an influential parameter for the plasma APAP concentration.
+However, based on the current sample size we still can find 6 parameters that can be an important parameter for the plasma APAP concentration.
 
@@
 The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct dynamical simulation for biological-based modeling, and it's mainly built for PK research with probabilistic approaches [@bois2009gnu].
-The \proglang{R} is a comprehensive statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
+The \proglang{R} is a well-developed statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
 We initiated \pkg{pksensi} that aim to conduct comprehensive and efficient global SA workflow in PK modeling, which hasn't been developed for open source communities.
@@
 Compare with \pkg{deSolve} package, we found that The \pkg{GNU MCSim} can provide better efficiency to solve the ODE in PK models.
-Although \pkg{pksensi} provide the features and functions to install and link with \pkg{GNU MCSim}, conducting the uncertainty and SA in PK modeling, the overall learning curve of this workflow might steep for users that are not familiar with \pkg{GNU MCSim} from model development, debugging, and testing.
+Although \pkg{pksensi} provide the essential features and functions to install and link with \pkg{GNU MCSim}, conducting the uncertainty and SA in PK modeling, the overall learning curve of this workflow might steep for users that are not familiar with \pkg{GNU MCSim} from model development, debugging, and testing.
 Therefore, the general \proglang{R} package, such as \pkg{deSolve}, can be an alternative option for \proglang{R} user.
@@
 Overall, \pkg{pksensi} provide an additional option of performing global SA for PK model and can create robust and reproducible results for decision making in model calibration.
+Although \pkg{pksensi} used here is mainly for PK model, it can also work for the other ODE-based dynamic model to investigate the sensitivity of model output from the input parameters.
 

### 2019-06-29 manuscript.Rmd
manuscript

@@
 This section mainly focus on how to conduct global SA with \pkg{deSolve} package.
-Then compare the computing time with \pkg{MCSim}.
+Then compare the computing time with \pkg{GNU MCSim}.
 
@@
 
-Using the \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic profile (Figure \ref{fig:ode-PK}) according to the given parameter baseline and the time points (\code{t}):
+Using the \code{ode} function in \pkg{deSolve} package, we can visualize the pharmacokinetic profile (Figure \ref{fig:ode-PK}) according to the given parameter baseline and the time points (\code{t}).
 
@@
 The \code{ka} are ranged 1.09 to 4.36 /h, corresponding to half-times of 0.32 and 0.64 hr.
-The $BW$ is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
+The \code{BW} is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
 The number of sample size determines the robustness of the result of SA.
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent SI of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake..'}
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent SI of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.'}
 plot(out, vars = "Ametabolized")
@@
 
-In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the computating error.
+In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the computing error.
 The Monte Carlo simulation is run for 1000 iteration as the assignment of \code{monte_carlo}.

### 2019-07-02 manuscript.Rmd
manuscript

@@
 It is also increasingly used in developing and assessing pharmacokinetic models.
-In our earlier work, we applied a global SA workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-Although several SA algorithms are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize SA results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
+For instance, we recently applied a global SA workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+Although several SA algorithms and software packages are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize SA results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
 Therefore, we developed an open-source \proglang{R} package, called **pksensi**, to fill this need and make SA more accessible in pharmacological and toxicological research.
@@
 It also refined the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
-Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
+Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the SA results.
 Utilizing pksensi, we successfully reproduced our previously published SA results of human PBPK modeling for acetaminophen and its two primary metabolites.
@@
 
-Sensitivity analysis (SA) is a mathematical technique to investigate how variations in model parameters affect model outputs.
+Sensitivity analysis (SA) is a mathematical technique to investigate how variations in model parameters affect model outputs [@saltelli2008global].
 An increasing number of studies use SA approach to determine which model parameters contribute to high variation in model predictions [@ferretti2016trends].
 This technique has also been applied in pharmacology and toxicology research [@loizou2015application; @mcnally2012reconstruction].
-Physiologically based kinetic (PK) modeling describes the changes in the concentrations or amounts of a substance within model compartments over time.
-The goal of SA in PK research is to examine the sensitivity of output variables (e.g.
+Pharmacokinetic models describe the changes in the concentrations or amounts of a substance in the body over time (the various terms for the kinetic models, including "pharmacokinetic", "toxicokinetic", and "biokinetic" models, are used exchangeably here).
+The goal of SA in pharmacokinetic research is to examine the sensitivity of output variables (e.g.
 chemical concentration in blood or tissues) in model responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
@@
 
-In our earlier work [@fphar201800588], we developed a workflow to apply global SA to reduce the computational burden in the Bayesian, Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
+In our earlier work [@fphar201800588], we developed an updated workflow to apply global SA to reduce the computational burden in the Bayesian, Markov Chain Monte Carlo (MCMC)-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
 We used \pkg{GNU MCSim} [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
@@
 It's more computationally efficient in calculating the influence of model parameter than other global SA methods such as the Monte Carlo based approaches from @jansen1999analysis and @owen2013better.
-Also, we found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
-This approach can solve identifiability issue of PBPK model parameter that increases the computational burden of Bayesian analysis, and therefore reduce the computational efficiency.
-We also found a useful features for communicating the parameter sensitivity in decision making.
+Our previous work also found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
+This approach address common identifiability issues of PBPK model parameter that increase the computational burden of Bayesian analysis [@garcia2015identifiability], and therefore reduce the computational efficiency.
+We developed approaches for communicating parameter sensitivity in decision making.
 
@@
 
-Although these tools provide various approaches to perform SA, we have not found a suitable package that provides functions with which users can visualize and distinguish the “non-influential” model parameters and can further apply to parameter fixing in pharmacokinetic and PK modeling.
+Although these tools provide various approaches to perform SA, they each individually have limitations for application to pharmacokinetic and PK modeling.
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make SA more accessible and reproducible in pharmacological and toxicological researches.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic and PK model with multivariate model outputs.
-The types of the kinetic model include the terms "pharmacokinetic", "toxicokinetic", and "biokinetic" model that are exchangeable in the model description.
+Moreover, it seamlessly integrates with both general ODE solvers available in R or as external \proglang{C} code, as well as the \pkg{GNU MCSim} software used for conducting sophisticated PK modeling, including MCMC.
 The analytical framework in \pkg{pkseni} not only include global SA but the uncertainty analysis and diagnostic tool to support the modeling process.
@@
 This type of global SA approach and SI also know as Sobol' method and Sobol' index.
-Since the estimated SI is not stable (high variation) when using the lower number of sample size, we included a random phase-shift approach to conduct replicating sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
+Since the estimated SI is not stable (high variation) when using smaller sample sizes, we included a random phase-shift approach to conduct replicating sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
 Considering the model that is built under $y=f(x_{i})$.
@@
 Through the random phase-shift, the robust result of the sensitivity measurement should be similar across replications under the same sample size.
+For most Sobol'-based SA algorithm, the larger sample size is needed to reduce the probability of having negative estimates. However, FAST estimates are always positive according to its construction.
 
@@
 This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed SI for all parameters across all time-points and output variables.
-Through random phase shift procedure, the SI can repeatly generate with low and high variation based on the given sampling size.
+Through random phase shift procedure, the replicate SI values can be repeatedly sampled independently.
 Therefore, they can be used to investigate the convergence of SI as
@@
 
-The R package \pkg{pksensi} provides a useful function to solve ODE in pharmacokinetic and PK models using analytical and numerical approaches.
+The R package \pkg{pksensi} provides a useful function to solve ODEs in pharmacokinetic and PK models using analytical and numerical approaches.
 The solution can be performed under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package [@JSSv033i09].
@@
 
-The uncertainty analysis is a crucial modeling process with SA.
+The uncertainty analysis is a crucial modeling process within SA.
 Sometimes, it's difficult to have informative data to set up the parameter distribution.
-The wider range of parameter distribution might cause the computing error at the extreme value.
-The narrower range might cause the incorrect calibrated result of the model parameter.
-The uncertainty analysis with Monte Carlo method is an appropriate approach in the data-driven modeling process to prevent this problem.
-Before fitting the data to model, the Monte Carlo can be applied with the given distribution of model parameter to exam whether the experiment data can be included in the distribution of model output.
+A wider range for a parameter distribution might cause a numerical error at extreme values.
+On the other hand, narrower ranges might cause the incorrect calibrated result of the model parameter.
+The uncertainty analysis using Monte Carlo is an appropriate approach in the data-driven modeling process to address this problem.
+Before fitting the data to model, the Monte Carlo can be applied with the given distribution of model parameter to examine whether the experiment data within the range of in the distribution of model output.
 \pkg{pksensi} also has a function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
@@
 Therefore, the inadequate sample number can create an unreliable result and bias.
-It's essential to use the sufficient sample size to prevent incorrect judgment in parameter fixing.
+It is essential to use a sufficient sample size to prevent incorrect judgment in parameter fixing.
 \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
@@
 
-We used the corresponding parameter value of  (APAP) in this example.
+We used the corresponding parameter value of acetaminophen (APAP) in this example.
+These model parameters are derived from the *in-vivo* or *in-vitro* experiment results.
 The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical properties for over 500 chemicals.
@@
 The body weight is assumed to be 70 (kg).
-
 ```{r, eval=T}
@@
 However, all parameters do not converge to the setting cut-off, which means the larger sample size is required in further sensitivity testing.
-Same as \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to exam the given output.
+Similiar to the \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to exam the given output.
 
@@
 
-In addtion to use \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can provide better computational efficiency.
-To solve ODE through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
+In addition to using \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can also be used, and provides better computational efficiency.
+To solve ODEs through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
 The computing time of using \code{solve_fun} in SA is estimated as,
@@
 After solving the equations under the same given condition, we can find that \pkg{GNU MCSim} has about 4 - 5 times faster in computing performance than using \pkg{deSolve}.
-In this case, we only focus on performing the global SA alone for generic PK model without additonal comparison with experiment data.
+In this case, we only focus on performing the global SA alone for generic PK model without additional comparison with experimental data.
 The next example will display and reproduce our previous published result [@fphar201800588] with full global SA workflow.
@@
 
-Same as the example of one-compartment PK model.
-The model parameter and the corresponding sampling range should be defined to create the parameter matrix.
-Previously, the probability distributions of model parameters were set to either truncated normal or uniform distribution when the parameters have informative prior information or not.
-To rapidly reach the acceptance convergence, we apply uniform distribution for all testing parameters.
+As in the example of one-compartment PK model, the model parameter and the corresponding sampling range should be defined to create the parameter matrix.
+Previously, the probability distributions of model parameters were set to either truncated normal or uniform distributions when the parameters have informative prior information or not, repectively.
+To rapidly reach the acceptance convergence in  this example, we apply uniform distribution for all testing parameters.
 The ranges of informative parameters are set to 1.96-times difference for single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
@@
 
-Generally, wide range of parameter value might cause the computing error when solving the differentiial equation.
+Generally, wide range of parameter value might cause the computing error when solving the differential equation.
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
@@
 The maximum number of step is set to 5000 in this case.
-Here we seperate the global SA of APAP-PBPK model process to several steps.
+Here we separate the global SA of APAP-PBPK model process to several steps.
 The value can reset through reinstall the the \pkg{GNU MCSim} by using \code{mcsim_install(mxstp = 5000)}.
@@
 
-The 21 testing model parameters are defined in this part, including parameter name, probability distribution, and distributed parameter value.
-To prevent the computing error, the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} need to adjust from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
+The 21 tested model parameters are defined in this step, including parameter name, probability distribution, and distributed parameter value.
+To prevent errors, the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} needed to be adjusted from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
 The objects \code{q} and \code{dist} are set to the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
@@
 To optimize the computing speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of APAP and its metabolites glucuronide (APAP-G) and sulfate (APAP-S) in plasma.
-The setting oral dose of APAP is 20 mg/kg in this example.
+The oral dose of APAP is 20 mg/kg in this example.
 Generally, the input dosing method can be defined through the \code{condition} argument.
@@
 
-In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the computing error.
-The Monte Carlo simulation is run for 1000 iteration as the assignment of \code{monte_carlo}.
+In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the numerical errors.
+The Monte Carlo simulation is run for 1000 iterations in the assignment of \code{monte_carlo}.
 The input file ('sim.in') and output file ('simmc.out') will be generated under the standard ASCII format.
@@
 
-In addition, through using the \code{check}, the parameter with sensitivity and convergence indices over the given condition can be preliminary detected for all output variables.
+In addition, through using \code{check}, the parameter with sensitivity and convergence indices over the given condition can be preliminary detected for all output variables.
 Based on our previous study, we proposed the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.
@@
 
-Computational modeling has become an important technique in modern scientific research to support the hypothesis, experiment, and discoveries.
-Uncertainty and SA is essential knowledge for modelers to conduct the model development and examination.
-More computational tools had been developed for various types of modeling and simulation.
-The \pkg{GNU MCSim} is a powerful simulation software package, that can conduct dynamical simulation for biological-based modeling, and it's mainly built for PK research with probabilistic approaches [@bois2009gnu].
-The \proglang{R} is a well-developed statistical computing language that can conduct probabilistic programming and have tons of packages for statistical analysis and data visualization.
-We initiated \pkg{pksensi} that aim to conduct comprehensive and efficient global SA workflow in PK modeling, which hasn't been developed for open source communities.
-The current version had included \pkg{GNU MCSim} and \pkg{deSolve} as the main solver in our model assessment.
-Compare with \pkg{deSolve} package, we found that The \pkg{GNU MCSim} can provide better efficiency to solve the ODE in PK models.
-Although \pkg{pksensi} provide the essential features and functions to install and link with \pkg{GNU MCSim}, conducting the uncertainty and SA in PK modeling, the overall learning curve of this workflow might steep for users that are not familiar with \pkg{GNU MCSim} from model development, debugging, and testing.
-Therefore, the general \proglang{R} package, such as \pkg{deSolve}, can be an alternative option for \proglang{R} user.
-Some developed package in \proglang{R} had also provide the features to solve ODE such as \pkg{RxODE} [@wang2016tutorial] and \pkg{mrgsolve} [@R-mrgsolve].
-We hope to incorporate these packages that can improve the user experience in using \pkg{pksensi} in the future.
+Computational modeling has become an important technique in modern scientific research to support hypotheses, experiments, and discoveries.
+Uncertainty and SA are essential tools for modelers in the conduct of model development and examination.
+Several packages on the CRAN repository and other software platforms had been built to perform global SA with a variety of algorithms, but none have included integration with ODE-based models, such as PK models.
 
-Moreover, some fundamental packages on the CRAN repository and other software platforms had been built to perform global SA with different algorithms.
-The development of \pkg{pksensi} is mainly for PK models that are constructed by ODE.
-It provides a straightforward application to investigate the impact of parameters on model output across simulation time-points and variables.
+We initiated \pkg{pksensi} that aim to conduct comprehensive and efficient global SA workflow in PK modeling, filling a key gap for open source modeling communities.
+It provides a straightforward application to investigate the impact of parameters on model outputs across simulation time-points and variables.
 We adopted the eFSAT method, a common variance-based SA that had better efficiency and accuracy in the model assessment.
@@
 In addition, the package can not only conduct global SA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most global SA studies and its software package.
-We also develop the function to visualize the output result that can help us distinguish the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model Calibration.
+We also develop the function to visualize the output result that can help us distinguish the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model calibration.
 In addition to eFAST, Some studies applied another global SA algorithm in the physiologically and biologically based ODE models, such as compartmental absorption and transit (also known as CAT) and other system biology models [@melillo2019variance; @sumner2012methodology].
-Base on our knowledge, the eFAST is a least computational expensive approch than other Monte-Carlo based Sobol' method.
-In future version, we expect to add additional global SA approach that had been applied in advanced studies [@glen2012estimating; @ring2017identifying].
+Base on our experience, eFAST is the least computationally expensive approach compared to other Monte-Carlo based Sobol’ method.
+In the future version, we expect to add additional global SA approach that had been applied in advanced studies [@glen2012estimating; @ring2017identifying].
 The other features such as principle component analysis [@lamboni2011multivariate] were considered to further include in our global SA workflow.
 
+We chose to integrate sensitivity analysis with \pkg{GNU MCSim} because it is a powerful simulation software package for dynamical simulation of biological-based models, mainly built for PK research with probabilistic approaches [@bois2009gnu] and had developed over decades.
+In addition, the current version of \pkg{pksensi} includes both \pkg{GNU MCSim} and \pkg{deSolve} as the main ODE solvers.
+Compared with the \pkg{deSolve} package, the \pkg{GNU MCSim} can provide better speed and efficiency to solve the ODEs in PK models.
+Although \pkg{pksensi} provides the essential features and functions to install and link with \pkg{GNU MCSim}, conducting the uncertainty and SA in PK modeling, the overall learning curve of this workflow might steep for users that are not familiar with \pkg{GNU MCSim} from model development, debugging, and testing.
+Therefore, the general \proglang{R} package, such as \pkg{deSolve}, can be an alternative option for R users.
+Some other packages in R are also available to solve ODEs, such as \pkg{RxODE} [@wang2016tutorial] and \pkg{mrgsolve} [@R-mrgsolve].
+We hope to incorporate these packages to expand the flexibility of using \pkg{pksensi} in the future.
+
 It is worth noting that the global SA results are limited to the specific model and the given range of the model parameter.
-The provided conditions, such as the number of examing parameter, parameter distribution, parameter range, and the given dosing level in PK model might have a potential impact on the level of importance of each parameter.
-The misconduct of the setting condition might lead to the incorrect estimated result.
+The provided conditions, such as the number of parameters examined, the parameter distributions/ranges, and the given dosing level in PK model might have a potential impact on the level of importance of each parameter.
 In other words, if we have various types of experiment data (e.g.
@@
 The complex model with a high dimension of the parameter is also a challenge in global SA, which will need higher computer demand in the model assessment.
-More additional applications still need to be investigated and strengthen.
-Overall, \pkg{pksensi} provide an additional option of performing global SA for PK model and can create robust and reproducible results for decision making in model calibration.
-Although \pkg{pksensi} used here is mainly for PK model, it can also work for the other ODE-based dynamic model to investigate the sensitivity of model output from the input parameters.
+As approaches are developed to address these complexities, we hope to incorporate them into \pkg{pksensi}.
+
+Overall, \pkg{pksensi} provides comprehensive features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model calibration.
+Although \pkg{pksensi} used here is mainly for pharmacokinetic modeling, it can also work for the other ODE-based dynamic models to investigate the sensitivity of model output from the input parameters.
 

### 2019-07-02 pbtk1cpt.c
pbtk1cpt model

@@ -1,0 +1,163 @@
+/* pbtk1cpt.c for R deSolve package
+   ___________________________________________________
+
+   Model File:  pbtk1cpt.model
+
+   Date:  Tue Jul  2 13:25:06 2019
+
+   Created by:  "mod v6.1.0"
+    -- a model preprocessor by Don Maszle
+   ___________________________________________________
+
+   Copyright (c) 1993-2019 Free Software Foundation, Inc.
+
+   Model calculations for compartmental model:
+
+   4 States:
+     Agutlument = 0.0,
+     Acompartment = 0.0,
+     Ametabolized = 0.0,
+     AUC = 0.0,
+
+   1 Output:
+    "Ccompartment",
+
+   0 Inputs:
+
+   4 Parameters:
+     kgutabs = 1,
+     vdist = 1,
+     ke = 1,
+     BW = 60,
+*/
+
+#include <R.h>
+#include <Rinternals.h>
+#include <Rdefines.h>
+#include <R_ext/Rdynload.h>
+
+/* Model variables: States */
+#define ID_Agutlument 0x00000
+#define ID_Acompartment 0x00001
+#define ID_Ametabolized 0x00002
+#define ID_AUC 0x00003
+
+/* Model variables: Outputs */
+#define ID_Ccompartment 0x00000
+
+/* Parameters */
+static double parms[4];
+
+#define kgutabs parms[0]
+#define vdist parms[1]
+#define ke parms[2]
+#define BW parms[3]
+
+/* Forcing (Input) functions */
+static double forc[0];
+
+
+/* Function definitions for delay differential equations */
+
+int Nout=1;
+int nr[1]={0};
+double ytau[1] = {0.0};
+
+static double yini[4] = {0.0, 0.0, 0.0, 0.0}; /*Array of initial state variables*/
+
+void lagvalue(double T, int *nr, int N, double *ytau) {
+  static void(*fun)(double, int*, int, double*) = NULL;
+  if (fun == NULL)
+    fun = (void(*)(double, int*, int, double*))R_GetCCallable("deSolve", "lagvalue");
+  return fun(T, nr, N, ytau);
+}
+
+double CalcDelay(int hvar, double dTime, double delay) {
+  double T = dTime-delay;
+  if (dTime > delay){
+    nr[0] = hvar;
+    lagvalue( T, nr, Nout, ytau );
+}
+  else{
[…]

### 2019-07-02 manuscript.Rmd
manuscript

@@
 chemical concentration in blood or tissues) in model responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
-In addition, this SA approach can be guidance of experimental design, parameter estimation during the model establishment, and qualification processes [@zhang2015sobol].
+In addition, this SA approach can be a guidance of experimental design, parameter estimation during the model establishment, and qualification processes [@zhang2015sobol].
 Through SA, it can optimal experimental design and distinguish the "unidentifiable" parameters that can lead to problems with numerical procedures used in parameter estimation [@chu2010quantitative].
@@
 Most of SA tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} [@pianosi2015matlab], \proglang{python} [@herman2017salib], and \proglang{R} [@iooss2017introduction].
-The \proglang{R} software is widely used and had numerious developed packages, such as \pkg{sensitivity} [@R-sensitivity], which is a practical tool to conduct local and global SA.
+The \proglang{R} software is widely used and had numerous developed packages, such as \pkg{sensitivity} [@R-sensitivity], which is a practical tool to conduct local and global SA.
 The \pkg{sensitivity} package includes some functions to generate the parameter sequences by using different computing algorithms.
@@
 Moreover, it seamlessly integrates with both general ODE solvers available in R or as external \proglang{C} code, as well as the \pkg{GNU MCSim} software used for conducting sophisticated PK modeling, including MCMC.
-The analytical framework in \pkg{pkseni} not only include global SA but the uncertainty analysis and diagnostic tool to support the modeling process.
+The analytical framework in \pkg{pksensi} not only include global SA but the uncertainty analysis and diagnostic tool to support the modeling process.
 The design concepts of \pkg{pksensi} are:
@@
 
-In the beginning, the user need to pre-install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
+In the beginning, the user need to install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
 In addition, \pkg{pksensi} can link with \proglang{R} \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
@@
 
-We adopted eFAST, a widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimentional parameters in the numerical models, especially for ODE-based dynamical models, in this package.
+We adopted eFAST, a widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimensional parameters in the numerical models, especially for ODE-based dynamical models, in this package.
 Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity].
@@
 
-where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$ is a vector giving the set of frequencies, one frequency for each parameter, and $\varphi$ is a random phasei-shift coefficient ranged from 0 to $2\pi$.
+where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$ is a vector giving the set of frequencies, one frequency for each parameter, and $\varphi$ is a random phase-shift coefficient ranged from 0 to $2\pi$.
 The default set of frequencies is based on the study from @saltelli1999quantitative.
@@
 
-To start, we implemented the one-compartment PK model in \proglang{R} by compiling the file written under the \pkg{GNU MCSim}'s model format.
-\pkg{pksensi} allows users select the preferred method to solve the pharmacokinetic model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
+To start, we implemented the one-compartment PK model in \proglang{R} by compiling the file written under the \proglang{C} or \pkg{GNU MCSim} format.
+\pkg{pksensi} allows users select the preferred method to solve the PK model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
 Running model under \pkg{GNU MCSim} native code can have a faster speed to obtain the model outputs.
@@
 The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.
-The model mainly includes two state variables that are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
+The model was developed on \pkg{GNU MCSim}'s format that mainly includes two state variables, which are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
 The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.
@@
 The code must first be compiled to run the model.
-After create the model file, we can use \code{compile_model} to generate the file that has dynamic-link library (DLL) or share object (SO) extention and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
+After create the model file, we can use \code{compile_model} to generate the file that has dynamic-link library (.dll) or share object (.so) extension and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
+
+```{r}
+mName <- "pbtk1cpt"
+```
 
 ```{r eval=T, message=FALSE, warning=FALSE}
-mName <- "pbtk1cpt"
 compile_model(mName, application = "R")
@@
 
-```{r}
+```{r, eval = T}
 source(paste0(mName, "_inits.R"))
+```
+
+On the other side, if the user only have the \proglang{C} version of the model code (\code{pbtk1cpt.c}), it can be compiled and loaded as:
+
+```{r, eval = F}
+system("R CMD SHLIB pbtk1cpt.c")
+dyn.load(paste0("pbtk1cpt", .Platform$dynlib.ext))
 ```
@@
 It can also schedule for the given dosing scenario.
-In the current setting, we assumed the initial condition of the intake dose to be 1000 mg.
+In the current setting, we assumed the initial condition of the intake dose to be 10 mg.
 We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the following modeling and SA.
-These additional functions are generated when compiling the model file.
+These additional functions are generated when compiling the model file through \pkg{GNU MCSim}.
+Therefore, if the model was compiled through \code{R CMD SHLIB}, the user need to assign the baseline value of model parameters and state variable on their own.
 
@@
 These model parameters are derived from the *in-vivo* or *in-vitro* experiment results.
-The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package, which includes physico-chemical properties for over 500 chemicals.
-The given value of \code{vdist}, \code{ke}, and \code{kgutabs} in \pkg{httk} are 1.1 (L/kg BW), 0.23 (/h), and 2.18 (/h), respectively.
-The body weight is assumed to be 70 (kg).
+The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package as,
+
+```{r, warning=F}
+library(httk)
+pars1comp <- (parameterize_1comp(chem.name = "acetaminophen"))
+unlist(pars1comp)
+```
+
+Therefore, the given value of \code{vdist}, \code{ke}, and \code{kgutabs} in \pkg{httk} are 1.1 (L/kg BW), 0.23 (/h), and 2.18 (/h), respectively. The body weight is assumed to be 70 (kg).
+
 ```{r, eval=T}
@@
 
-Here shows the given parameter value (\code{parms}), initial state condition (\code{initStat}), and the output variable (\code{Outputs}) that need to specify in model solving.
+Here shows the given parameter value (\code{parms}), initial state condition (\code{initState}), and the output variable (\code{Outputs}) that need to specify in model solving.
 
@@
 
+Both \code{parms} and \code{initState} are "numeric variables" that contain the value of parameter and initial state condition.
+
 ```{r, eval=T}
@@
 ```
+
+The \code{Outputs} is a "character" that specify the estimated model output of concentration in rest body.
 
@@
 The replication is set to 20.
-Therefore, the totol of 16,000 parameter sequence will be used to compute the corresponding outputs.
+Therefore, the total of 16,000 parameter sequence will be used to compute the corresponding outputs.
 Figure \ref{fig:rep-sampl} plotted the sampling process for each parameter from the first 3 replications.
@@
 The default output in the plotting is setting at the first variable.
-To exam the time-sependent SI of other variables, such as \code{Ametabolized} in this case, we need to assign the variable name \code{vars = "Ametabolized"} in \code{plot} function (Figure \ref{fig:pk-solvefun-Amet}).
+To exam the time-dependent SI of other variables, such as \code{Ametabolized} in this case, we need to assign the variable name \code{vars = "Ametabolized"} in \code{plot} function (Figure \ref{fig:pk-solvefun-Amet}).
 
@@
 The convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
-The output are formated as 4-D array in \code{y} with the dimension name of model evaluation, number of replications, number of time points, and number of output variables, respectively.
+The output are formatted as 4-D array in \code{y} with the dimension name of model evaluation, number of replications, number of time points, and number of output variables, respectively.
 
@@
 However, all parameters do not converge to the setting cut-off, which means the larger sample size is required in further sensitivity testing.
-Similiar to the \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to exam the given output.
+Similar to the \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to exam the given output.
 
@@
 
-Then, before we conduct the SA through \pkg{GNU MCSim}, The following code is used to compile the \pkg{GNU MCSim} model code to the exexutable program.
+Then, before we conduct the SA through \pkg{GNU MCSim}, The following code is used to compile the \pkg{GNU MCSim} model code to the executable program.
 
@@
 
-As in the example of one-compartment PK model, the model parameter and the corresponding sampling range should be defined to create the parameter matrix.
-Previously, the probability distributions of model parameters were set to either truncated normal or uniform distributions when the parameters have informative prior information or not, repectively.
+As in the example of the one-compartment PK model, the model parameter and the corresponding sampling range should be defined to create the parameter matrix.
+Previously, the probability distributions of model parameters were set to either truncated normal or uniform distributions when the parameters have informative prior information or not, respectively.
 To rapidly reach the acceptance convergence in this example, we apply uniform distribution for all testing parameters.
-The ranges of informative parameters are set to 1.96-times difference for single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
+The ranges of informative parameters are set to 1.96-times difference for the single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
 The nominal values of informative model parameters were defined as:
@@
 
-To conduct the global SA with \pkg{GNU MCsim} and \pkg{pksensi}, the input file with given "setpoint" condition should be generated before modeling.
+To conduct the global SA with \pkg{GNU MCSim} and \pkg{pksensi}, the input file with given "setpoint" condition should be generated before modeling.
 The file can create by \code{generate_infile} function.
@@
 It provides a straightforward application to investigate the impact of parameters on model outputs across simulation time-points and variables.
-We adopted the eFSAT method, a common variance-based SA that had better efficiency and accuracy in the model assessment.
+We adopted the eFAST method, a common variance-based SA that had better efficiency and accuracy in the model assessment.
 It can exam the parameter effect from first order to the total order.
@@
 In addition to eFAST, Some studies applied another global SA algorithm in the physiologically and biologically based ODE models, such as compartmental absorption and transit (also known as CAT) and other system biology models [@melillo2019variance; @sumner2012methodology].
-Base on our experience, eFAST is the least computationally expensive approach compared to other Monte-Carlo based Sobol’ method.
+Base on our experience, eFAST is the least computationally expensive approach compared to other Monte Carlo based Sobol’ method.
 In the future version, we expect to add additional global SA approach that had been applied in advanced studies [@glen2012estimating; @ring2017identifying].
@@
 This work was funded, in part, by grant 1U01FD005838 from the U.S.
-FDA.
+Food and Drug Administration and grant P42 ES027704 from the U.S.
+National Institute of Environmental Health Sciences.
 This article reflects the views of the author and should not be construed to represent FDA’s views or policies.

### 2019-07-02 manuscript.Rmd
MANUSCRIPT

@@
 output: rticles::jss_article
-
+#output: word_document
 ---
@@
 
-The R package \pkg{pksensi} is available from the comprehensive R Archive Network at https://CRAN.R-project.org/package=pksensi.
+The R package \pkg{pksensi} is available from the Comprehensive R Archive Network at https://CRAN.R-project.org/package=pksensi.
 The latest up-to-date version is available through GitHub at https://github.com/nanhung/pksensi.
@@
 In addition, \pkg{pksensi} can link with \proglang{R} \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
-The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the build-in function \code{mcsim_instal} in \pkg{pksensi}.
+The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the built-in function \code{mcsim_instal} in \pkg{pksensi}.
 The \proglang{C} compiler had originated in Linux or MacOS.
-The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as,
+The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as:
 
@@
 Considering the model that is built under $y=f(x_{i})$.
-The sampling scheme can be written as:
+The sampling scheme can be written as,
 
@@
 Through random phase shift procedure, the replicate SI values can be repeatedly sampled independently.
-Therefore, they can be used to investigate the convergence of SI as
+Therefore, they can be used to investigate the convergence of SI as,
 
@@
 In this section, two types of PK model were applied here to demonstrate how \pkg{pksensi} work.
-The first one is generic model and the second is the chemical-specific model.
+The first one is a generic model and the second is a chemical-specific model.
 The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current example.
@@
 In the first example, we use a generic, one-compartment PK model from \pkg{httk} package [@JSSv079i04] to demonstrate how \pkg{pksensi} can be applied to PK studies.
-The differential equations for this model is as:
+The differential equations for this model is as,
 
@@
 
-The concentration of the chemical in the rest of body and blood ($C_{rest}$, mg/L) can be calculated as
+The concentration of the chemical in the rest of body and blood ($C_{rest}$, mg/L) can be calculated as,
 
@@
 
-On the other side, if the user only have the \proglang{C} version of the model code (\code{pbtk1cpt.c}), it can be compiled and loaded as:
+On the other side, if the user only has the \proglang{C} version of the model code (\code{pbtk1cpt.c}), it can be compiled and loaded as:
 
@@
 These model parameters are derived from the *in-vivo* or *in-vitro* experiment results.
-The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package as,
+The parameter value can be generated from \code{parameterize_1comp} function in \pkg{httk} package as:
 
@@
 
-The \code{Outputs} is a "character" that specify the estimated model output of concentration in rest body.
+The \code{Outputs} is a "character" that specifies the estimated model output of concentration in rest body.
 
@@
 
-The generated parameters are stored as a 3-D array under the named \code{a}, with the dimension of sample size, number of replication, and number of parameter, respectively.
+The generated parameters are stored as a 3-D array under the named \code{a}, with the dimension of sample size, number of replications, and number of parameters, respectively.
 
@@
 However, the current result is only based on the setting distribution of model parameters for APAP.
-The different given input conditions (e.g., range of parameter uncertainty, chemical dependent parameter value) can change the result (result not shown).
+The different given input conditions (e.g., range of parameter uncertainty, chemical-dependent parameter value) can change the result (result not shown).
 
@@
 
-The Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is also determined by parameter \code{ke}.
+Figure \ref{fig:pk-solvefun-Amet} shows that the amount of metabolized is also determined by parameter \code{ke}.
 Same as \code{Ccompartment}, the \code{kgutabs} contribute about 30 - 40% variation of model output in the first hour.
@@
 
-Some functions in \pkg{pksensi} provide a efficient way to check the result from global SA.
+Some functions in \pkg{pksensi} provide an efficient way to check the result from global SA.
 The \code{check} can determine which parameters have relatively lower sensitivity measurement across the given time points and model outputs, and therefore can be applied parameter fixing in model calibration.
@@
 
-Similar to \code{solve_fun} function that can define the initial value of parameter and state variable through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as exposure dose, fixed parameter value or initial condition of state variable.
+Similar to \code{solve_fun} function that can define the initial value of parameters and state variable through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as exposure dose, fixed parameter value or initial condition of state variables.
 
@@
 
-In global SA, we have to additionally generate the parameter matrix from the eFAST method. The current setting uses 512 sample size with 10 replication.
+In global SA, we have to generate the parameter matrix from the eFAST method additionally. The current setting uses sample sizes of 512 with 10 replication.
 

### 2019-07-03 manuscript.Rmd
"W"orkflow

@@
 ```{r workflow, echo = FALSE, out.width = "100%", fig.cap = '\\label{fig:Workflow}Workflow for performing uncertainty and sensitivity analysis.'}
-knitr::include_graphics("diagram/workflow.png")
+knitr::include_graphics("diagram/Workflow.png")
 ```

### 2019-07-07 manuscript.Rmd
manuscript

@@
 Although these tools provide various approaches to perform SA, they each individually have limitations for application to pharmacokinetic and PK modeling.
+For example, the estimated index in SA is not robust when using the inadequate size of the sample number.
+However, there are no software packages that provide the available tools in the assessment of reliability.
+In addition, the developed eFAST function doesn't include the feature to generate the random sampling curve.
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make SA more accessible and reproducible in pharmacological and toxicological researches.

### 2019-07-10 manuscript.Rmd
Revised (based on ET comments)

@@
 $$\frac{dA_{gutlumen}}{dt} = -k_{gutabs} \cdot A_{gutlumen} + g(t)$$
-$$\frac{dA_{rest}}{dt} = \frac{dA_{gutlumen}}{dt}-k_{elim} \cdot A_{rest}$$
+$$\frac{dA_{rest}}{dt} = k_{gutabs} \cdot A_{gutlumen}-k_{elim} \cdot A_{rest}$$
 
@@
 The \code{check} also provides some feasible argument to specify the target output or change the cut-off value.
-The argument of \code{SI.cutoff} is setting at 0.05 to detect the relative non-influential parameters as default.
+The argument of \code{SI.cutoff} is setting at 0.05 to detect the relative non-influential parameters as default, representing the 5% change of the output is contributed from the specific parameter variation.
 
@@
 The aim of this section is to apply global SA for APAP-PBPK model through \pkg{pksensi}.
+The model describes the pharmacokinetics of parent APAP and its two metabolites glucuronide (APAP-G) and sulfate (APAP-S) through the dozens of parameters.
 The model code is included in this package in https://github.com/nanhung/pksensi/blob/master/tests/pbpk_apap.model and can be generated through \code{pbpk_apap_model} function.
@@
        "Sulfation pathway acetaminophen, Vmax",
-       "Glucronidation pathway acetaminophen, Km",
-       "Glucronidation pathway substrate inhibition, Ki",
-       "Glucronidation pathway GA, Km",
-       "Glucronidation pathway acetaminophen, Vmax",
+       "Glucuronidation pathway acetaminophen, Km",
+       "Glucuronidation pathway substrate inhibition, Ki",
+       "Glucuronidation pathway GA, Km",
+       "Glucuronidation pathway acetaminophen, Vmax",
        "APAP-G hepatic transporter, Km",
@@
 
-To optimize the computing speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of APAP and its metabolites glucuronide (APAP-G) and sulfate (APAP-S) in plasma.
+To optimize the computing speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of APAP and its metabolites in plasma.
 The oral dose of APAP is 20 mg/kg in this example.

### 2019-07-12 manuscript.Rmd
MANUSCRIPT

@@
 This work was funded, in part, by grant 1U01FD005838 from the U.S.
-Food and Drug Administration and grant P42 ES027704 from the U.S.
+Food and Drug Administration (FDA) and grant P42 ES027704 from the U.S.
 National Institute of Environmental Health Sciences.
@@
 Yasha Hartberg and Dr.
-Barbara Gastel at the Texas A&M University and Dr.
-Frederic Bois from Certara for reviewing the manuscript and consultation.
+Barbara Gastel at the Texas A&M University, Dr.
+Frederic Bois from Certara and Dr.
+Eleftheria Tsakalozou from U.S.
+FDA for reviewing the manuscript and consultation.
 

### 2019-08-08 manuscript.Rmd
documentclass: jss

@@
 ---
+documentclass: jss
 author:

### 2019-08-11 manuscript.Rmd
revise (v3)

@@
 Therefore, we developed an open-source \proglang{R} package, called **pksensi**, to fill this need and make SA more accessible in pharmacological and toxicological research.
-This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, such as PBPK model with multivariate model outputs (multiple time-points and tissue compartments).
+This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, including those with multivariate model outputs, such as multiple time-points and tissue compartments.
 It also refined the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
-Furthermore, pksensi includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the SA results.
+Furthermore, pksensi includes functions to check the convergence of the SA results.
 Utilizing pksensi, we successfully reproduced our previously published SA results of human PBPK modeling for acetaminophen and its two primary metabolites.
@@
 Pharmacokinetic models describe the changes in the concentrations or amounts of a substance in the body over time (the various terms for the kinetic models, including "pharmacokinetic", "toxicokinetic", and "biokinetic" models, are used exchangeably here).
-The goal of SA in pharmacokinetic research is to examine the sensitivity of output variables (e.g.
-chemical concentration in blood or tissues) in model responds to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
-In addition, this SA approach can be a guidance of experimental design, parameter estimation during the model establishment, and qualification processes [@zhang2015sobol].
-Through SA, it can optimal experimental design and distinguish the "unidentifiable" parameters that can lead to problems with numerical procedures used in parameter estimation [@chu2010quantitative].
+The goal of SA in pharmacokinetic research is to examine the sensitivity of output variables,such as chemical concentration in blood or tissues, with respect to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
+In addition, this SA approach can guide experimental design and the parameter estimation processes [@zhang2015sobol].
+For instance, SA can identify "unidentifiable" parameters that can lead to problems with numerical procedures used in parameter estimation [@chu2010quantitative].
 It can be further applied to parameter prioritization and parameter fixing before model calibration [@fphar201800588].
@@
 We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based global SA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
-This method was first introduced in @cukier1973study (the classic FAST method with first-order effect only) that study of the sensitivity of coupled reaction systems that are constructed by differential equations with the numerous rate coefficients and then evolved to Sobol' sensitivity measure in @saltelli1999quantitative (the eFAST method with first- and total-order effects).
-The FAST algorithm creates the multidimensional space of input parameters through "search-curve".
-It's more computationally efficient in calculating the influence of model parameter than other global SA methods such as the Monte Carlo based approaches from @jansen1999analysis and @owen2013better.
+This method was first introduced in @cukier1973study (the classic FAST method with first-order effect only) in a study of the sensitivity of coupled reaction systems that are constructed by differential equations with the numerous rate coefficients.
+It then evolved to Sobol' sensitivity measure in @saltelli1999quantitative (the eFAST method with first- and total-order effects).
+The FAST algorithm creates the multidimensional space of input parameters through a "search-curve".
+It's more computationally efficient in calculating the influence of model parameter than other global SA methods such as the Monte Carlo based approaches [@jansen1999analysis; @owen2013better].
 Our previous work also found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
-This approach address common identifiability issues of PBPK model parameter that increase the computational burden of Bayesian analysis [@garcia2015identifiability], and therefore reduce the computational efficiency.
-We developed approaches for communicating parameter sensitivity in decision making.
+This approach addresses common identifiability issues of PBPK model parameter that increase the computational burden of Bayesian analysis [@garcia2015identifiability], and therefore reduce the computational efficiency.
+We also developed approaches for communicating parameter sensitivity in decision making.
 
@@
 However, there are no software packages that provide the available tools in the assessment of reliability.
-In addition, the developed eFAST function doesn't include the feature to generate the random sampling curve.
+In addition, the existing eFAST function does not includes the feature to generate the random sampling curve.
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make SA more accessible and reproducible in pharmacological and toxicological researches.
@@
 Moreover, it seamlessly integrates with both general ODE solvers available in R or as external \proglang{C} code, as well as the \pkg{GNU MCSim} software used for conducting sophisticated PK modeling, including MCMC.
-The analytical framework in \pkg{pksensi} not only include global SA but the uncertainty analysis and diagnostic tool to support the modeling process.
+The analytical framework in \pkg{pksensi} not only includes global SA but also the uncertainty analysis and diagnostic tools to support the modeling process.
 The design concepts of \pkg{pksensi} are:
@@
 
-In the beginning, the user need to install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
-In addition, \pkg{pksensi} can link with \proglang{R} \pkg{deSolve} package, which includes the comprehensive integrator to solve ODE.
-The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the built-in function \code{mcsim_instal} in \pkg{pksensi}.
-The \proglang{C} compiler had originated in Linux or MacOS.
+To start, the users need to install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
+In addition, \pkg{pksensi} can link with \proglang{R} \pkg{deSolve} package, which includes a comprehensive integrator to solve ODE.
+The \pkg{GNU MCSim} can be installed by following the instruction in \pkg{GNU MCSim}’s manual on https://www.gnu.org/software/mcsim/mcsim.html or using the built-in function \code{mcsim_install} in \pkg{pksensi}.
+The \proglang{C} compiler is already available in Linux or MacOS.
 The Windows user can install Rtools on https://cran.r-project.org/bin/windows/Rtools/ and set the working path of compiler (\code{/Rtools/mingw_64/bin} and \code{/Rtools/bin}) before using this function as:
@@
 
-We adopted eFAST, a widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimensional parameters in the numerical models, especially for ODE-based dynamical models, in this package.
+We adopted eFAST, a widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimensional parameters in numerical models in this package.
 Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity].
@@
 Also, time-dependent sensitivity measurements of first and total order effects for all parameters can be plotted simultaneously.
-The computed SI might show different estimation across replications under the same sample size.
-Therefore, the inadequate sample number can create an unreliable result and bias.
-It is essential to use a sufficient sample size to prevent incorrect judgment in parameter fixing.
+If the computed SI show different values across replications under the same sample size, this indicates inadequate sample number, which can create an unreliable results.
+It is essential to use a sufficient sample size to prevent incorrect judgments in parameter fixing.
 \pkg{pksensi} includes functions to check the convergence and sensitivity of model parameters, providing a means to assess the robustness of the sensitivity measurement.
-We also developed a "cut-off"-based approach in this package to accurately distinguish between "influential" and "non-influential" parameters.
+We also developed a "cut-off"-based approach in this package to distinguish between "influential" and "non-influential" parameters.
 Finally, \pkg{pksensi} provides visualization tools for the effective investigation and communication of results in decision support.
@@
 
-In this section, two types of PK model were applied here to demonstrate how \pkg{pksensi} work.
+In this section, two PK models were applied to demonstrate how \pkg{pksensi} works.
 The first one is a generic model and the second is a chemical-specific model.
@@
 The aim of this section is to apply global SA for APAP-PBPK model through \pkg{pksensi}.
-The model describes the pharmacokinetics of parent APAP and its two metabolites glucuronide (APAP-G) and sulfate (APAP-S) through the dozens of parameters.
+The model describes the pharmacokinetics of parent APAP and its two metabolites glucuronide (APAP-G) and sulfate (APAP-S).
 The model code is included in this package in https://github.com/nanhung/pksensi/blob/master/tests/pbpk_apap.model and can be generated through \code{pbpk_apap_model} function.
@@
 The model code needs to be prepared in the following global SA workflow.
-After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through \code{compile_model} function.
+After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through the \code{compile_model} function.
 
@@
 
-The 21 tested model parameters are defined in this step, including parameter name, probability distribution, and distributed parameter value.
+The 21 tested model parameters are defined in this step, including parameter names and probability distributions.
 To prevent errors, the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} needed to be adjusted from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
@@
 Since the unit of the given dose is mg/kg, the \code{mgkg_flag} is set to 1.
-More definition of input schedule functions can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
+Additional input schedule functions can be found in the section of input functions in \pkg{GNU MCSim} User’s Manual (https://www.gnu.org/software/mcsim/mcsim.html#Input-functions).
 
@@
 We apply uncertainty analysis through the \code{solve_mcsim} and visualize the result by \code{pksim} function.
-Some example data are included in the \pkg{pksensi} with experiment time (h) and concentration (mg/L).
+Some example data are included in \pkg{pksensi} with experiment time (h) and concentration (mg/L).
 
@@
 
-In the setting condition of simulation, The relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the numerical errors.
+In setting simulation conditions, the relative and absolute error tolerance (\code{rtol} & \code{atol}) were set to 1e-7 and 1e-9, respectively, to prevent the numerical errors.
 The Monte Carlo simulation is run for 1000 iterations in the assignment of \code{monte_carlo}.
@@
 Figure \ref{fig:pksim-MC} shows the uncertainty analysis of APAP data and modeling results.
-For parent compound, all data points are located in the simulated interval of 25-75%.
-Through this result, we can determine that the simulated outputs can accurately generate the same concentration profile as the in-vivo experiment under the setting of parameter ranges for APAP.
-The simulated result of metabolites APAP-G shows the different pharmacokinetic profile with experiment data.
-However, all data points are located in the simulated interval.
+For parent compound, all data points are located in the simulated interval of 25-75%, suggesting that the simulated outputs can accurately cover the experimental concentration profiles under the assumed parameter ranges for APAP.
+The simulated results of metabolites APAP-G and APAP-S, while not fully in the 25-75% range, and nonetheless within the envelope of the simulated intervals.
 
@@
 
-In global SA, we have to generate the parameter matrix from the eFAST method additionally. The current setting uses sample sizes of 512 with 10 replication.
+To perform global SA, we next have to generate the parameter matrix from the eFAST method. An example with sample size 512 with 10 replications would be as follows:
 
@@
 
-To conduct the global SA with \pkg{GNU MCSim} and \pkg{pksensi}, the input file with given "setpoint" condition should be generated before modeling.
+To conduct the global SA with \pkg{GNU MCSim} and \pkg{pksensi}, an input file needs to be generate before modeling.
 The file can create by \code{generate_infile} function.
-The \code{solve_mcsim} can also automatically create the input file and compute the output.
+Alternatively, \code{solve_mcsim} can also automatically create the input file and compute the output.
 
@@
 
-In addition, through using \code{check}, the parameter with sensitivity and convergence indices over the given condition can be preliminary detected for all output variables.
+In addition, through using \code{check}, the sensitivity and convergence indices can be preliminary checked for all output variables.
 Based on our previous study, we proposed the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.
@@
 
-We initiated \pkg{pksensi} that aim to conduct comprehensive and efficient global SA workflow in PK modeling, filling a key gap for open source modeling communities.
+We initiated \pkg{pksensi} in order to facilitiate a comprehensive and efficient global SA workflow in PK modeling, filling a key gap for open source modeling communities.
 It provides a straightforward application to investigate the impact of parameters on model outputs across simulation time-points and variables.
-We adopted the eFAST method, a common variance-based SA that had better efficiency and accuracy in the model assessment.
-It can exam the parameter effect from first order to the total order.
-In addition, the package can not only conduct global SA but also include the function to assess the convergence of sensitivity measurements, which is rarely discussed in most global SA studies and its software package.
-We also develop the function to visualize the output result that can help us distinguish the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model calibration.
-In addition to eFAST, Some studies applied another global SA algorithm in the physiologically and biologically based ODE models, such as compartmental absorption and transit (also known as CAT) and other system biology models [@melillo2019variance; @sumner2012methodology].
-Base on our experience, eFAST is the least computationally expensive approach compared to other Monte Carlo based Sobol’ method.
-In the future version, we expect to add additional global SA approach that had been applied in advanced studies [@glen2012estimating; @ring2017identifying].
-The other features such as principle component analysis [@lamboni2011multivariate] were considered to further include in our global SA workflow.
+We adopted the eFAST method, a common variance-based SA that we found to have the best balance of efficiency and accuracy.
+In addition, the package also includes the ability to assess the convergence of SA results, which is rarely addressed in most global SA studies and software packages.
+We also developed functions to visualize the output results and help distinguish the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model calibration.
+In the future version, we expect to add additional global SA approaches [@glen2012estimating; @ring2017identifying].
+The other features such as principle component analysis [@lamboni2011multivariate] will also be considered.
 
-We chose to integrate sensitivity analysis with \pkg{GNU MCSim} because it is a powerful simulation software package for dynamical simulation of biological-based models, mainly built for PK research with probabilistic approaches [@bois2009gnu] and had developed over decades.
+We chose to integrate sensitivity analysis with \pkg{GNU MCSim} because it is a powerful open-source simulation software package for dynamical simulation of biological-based models, mainly built for PK research with probabilistic approaches [@bois2009gnu].
 In addition, the current version of \pkg{pksensi} includes both \pkg{GNU MCSim} and \pkg{deSolve} as the main ODE solvers.
@@
 
-It is worth noting that the global SA results are limited to the specific model and the given range of the model parameter.
-The provided conditions, such as the number of parameters examined, the parameter distributions/ranges, and the given dosing level in PK model might have a potential impact on the level of importance of each parameter.
-In other words, if we have various types of experiment data (e.g.
-different exposure routes and biological measurements) and will use them in model calibration, it will enhance the difficulties in global SA.
-The complex model with a high dimension of the parameter is also a challenge in global SA, which will need higher computer demand in the model assessment.
+It is worth noting that the current global SA results are conditional on the specific model and the given range of the model parameters.
+Other factors, such as the number of parameters examined, the parameter distributions/ranges, and the given dosing level in PK model might also have a potential impact on the level of importance of each parameter, but these "meta-uncertainties" are not currently addressed.
+The complex models with a high dimensionality in the the parameter is also a challenge in global SA as they have higher computer demands.
 As approaches are developed to address these complexities, we hope to incorporate them into \pkg{pksensi}.
 
-Overall, \pkg{pksensi} provides comprehensive features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model calibration.
-Although \pkg{pksensi} used here is mainly for pharmacokinetic modeling, it can also work for the other ODE-based dynamic models to investigate the sensitivity of model output from the input parameters.
+Overall, \pkg{pksensi} provides a comprehensive suite features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model development and calibration.
+Although \pkg{pksensi} used here is mainly for pharmacokinetic modeling, it can also be applied to other ODE-based dynamic models in order to investigate the sensitivity of model outputs to input parameters.
 

### 2019-08-12 manuscript.Rmd
manuscript

@@
 
-To start, the users need to install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
 In addition, \pkg{pksensi} can link with \proglang{R} \pkg{deSolve} package, which includes a comprehensive integrator to solve ODE.
@@
 
-To start, we implemented the one-compartment PK model in \proglang{R} by compiling the file written under the \proglang{C} or \pkg{GNU MCSim} format.
+To start, we implemented the one-compartment PK model in \proglang{R}.
 \pkg{pksensi} allows users select the preferred method to solve the PK model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
 Running model under \pkg{GNU MCSim} native code can have a faster speed to obtain the model outputs.
-This section mainly focus on how to conduct global SA with \pkg{deSolve} package.
+This section mainly focus on how to conduct global SA with \pkg{deSolve} package with pure \proglang{R} environment.
 Then compare the computing time with \pkg{GNU MCSim}.
 
-The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.
-The model was developed on \pkg{GNU MCSim}'s format that mainly includes two state variables, which are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
+The one-compartment PK model was mainly includes two state variables, which are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
 The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.
@@
 ```{r, eval=T}
-pbtk1cpt_model()
-cat(readLines("pbtk1cpt.model"), sep = "\n")
+pbtk1cpt <- function(t, state, parameters) {
+  with(as.list(c(state, parameters)), {
+    dAgutlument = - kgutabs * Agutlument
+    dAcompartment = kgutabs * Agutlument - ke * Acompartment
+    dAmetabolized = ke * Acompartment
+    Ccompartment = Acompartment / vdist * BW;
+    list(c(dAgutlument, dAcompartment, dAmetabolized), 
+         "Ccompartment" = Ccompartment) 
+  })
+}
 ```
 
-The code must first be compiled to run the model.
-After create the model file, we can use \code{compile_model} to generate the file that has dynamic-link library (.dll) or share object (.so) extension and can be linked dynamically into an \proglang{R} session (\code{pbtk1cpt.dll} on Windows or \code{pbtk1cpt.so} on other systems) and \proglang{R} file (\code{pbtk1cpt_inits.R}) with default input parameters and initial state settings with the definition of \code{application = "R"}.
-
-```{r}
-mName <- "pbtk1cpt"
-```
-
-```{r eval=T, message=FALSE, warning=FALSE}
-compile_model(mName, application = "R")
-```
-
-The \code{pbtk1cpt_inits.R} file includes \code{initParms} and \code{initStates} functions and \code{Outputs} variable.
-The created function have default value of model parameters and initial conditions that can further use to customize in the simulation.
-
-```{r, eval = T}
-source(paste0(mName, "_inits.R"))
-```
-
-On the other side, if the user only has the \proglang{C} version of the model code (\code{pbtk1cpt.c}), it can be compiled and loaded as:
-
-```{r, eval = F}
-system("R CMD SHLIB pbtk1cpt.c")
-dyn.load(paste0("pbtk1cpt", .Platform$dynlib.ext))
-```
-
-The parameter values and initial states can be customized to the specific condition.
-It can also schedule for the given dosing scenario.
-In the current setting, we assumed the initial condition of the intake dose to be 10 mg.
-We can use \code{initParms} and \code{initStates} functions to customize the parameter values and the initial state that will be used in the following modeling and SA.
-These additional functions are generated when compiling the model file through \pkg{GNU MCSim}.
-Therefore, if the model was compiled through \code{R CMD SHLIB}, the user need to assign the baseline value of model parameters and state variable on their own.
-
+The parameter values and initial states need to assign to the specific condition.
 We used the corresponding parameter value of acetaminophen (APAP) in this example.
@@
 ```{r, eval=T}
-parms <- initParms()
-parms["vdist"] <- 1.1
-parms["ke"] <- 0.23
-parms["kgutabs"] <- 2.18
-parms["BW"] <- 70
-initState <- initStates(parms=parms)
-initState["Agutlument"] <- 10
+parms <- c(vdist = pars1comp$Vdist, 
+           ke = pars1comp$kelim, 
+           kgutabs = pars1comp$kgutabs, 
+           BW = pars1comp$BW)
+initState <- c(Agutlument = 10, Acompartment = 0, Ametabolized = 0)
 ```
 
-Here shows the given parameter value (\code{parms}), initial state condition (\code{initState}), and the output variable (\code{Outputs}) that need to specify in model solving.
-
-```{r, eval=T}
-parms
-```
-
-```{r, eval=T}
-initState
-```
-
+Here shows the given parameter value (\code{parms}) and initial state condition (\code{initState}) that need to specify in model solving.
 Both \code{parms} and \code{initState} are "numeric variables" that contain the value of parameter and initial state condition.
-
-```{r, eval=T}
-Outputs
-```
-
-The \code{Outputs} is a "character" that specifies the estimated model output of concentration in rest body.
 
@@
 ```{r, eval=T}
+library(deSolve)
 t <- seq(from = 0.01, to = 24.01, by = 1)
-y <- deSolve::ode(initState, t, func = "derivs", parms = parms, 
-         dllname = mName, initfunc = "initmod", 
-         nout = 1, outnames = Outputs)
+y <- ode(y = initState, times = t, func = pbtk1cpt, parms = parms)
 ```
@@
 q <- c("qunif", "qunif", "qunif", "qnorm")
-q.arg <- list(list(min = parms["vdist"] / 2, max = parms["vdist"] * 2),
-             list(min = parms["ke"] / 2, max = parms["ke"] * 2), 
-             list(min = parms["kgutabs"] / 2, max = parms["kgutabs"] * 2),
-             list(mean = parms["BW"], sd = 5)) 
+q.arg <- list(list(min = pars1comp$Vdist / 2, max = pars1comp$Vdist * 2),
+              list(min = pars1comp$kelim / 2, max = pars1comp$kelim * 2),
+              list(min = pars1comp$kgutabs / 2, max = pars1comp$kgutabs * 2),
+              list(mean = pars1comp$BW, sd = 5))
 params <- c("vdist", "ke", "kgutabs", "BW")
@@
 set.seed(1234)
-x <- rfast99(params, n = 200, q = q, q.arg = q.arg, replicate = 20)
+x <- rfast99(params, n = 200, q = q, q.arg = q.arg, replicate = 10)
 ```
@@
 Because the PK model is being used to describe a continuous process for the chemical concentration over time, the sensitivity measurements, therefore, have the time-dependent relationships for each model parameter.
-Here we define the output time points (\code{t}) to examine the change of the parameter sensitivity over time.
-To solve the model through \pkg{deSolve}, we need to provide the details of the argument, which include time (\code{t}), initial  conditions of state variable (\code{initState}), output variables (\code{outnames}), and name of the shared library (\code{mName}, without extension).
+Here we use the defined output time points (\code{t}) to examine the change of the parameter sensitivity over time.
+To solve the model through \pkg{deSolve}, we need to provide the details of the argument, which include time (\code{t}), initial  conditions of state variable (\code{initState}), output variables (\code{outnames}), and name of the model function (\code{func}).
 To create the time-dependent sensitivity measurement, we set the time duration from 0.01 to 24.01 hours with the time segment of 1 hour as the above definition in \code{ode} function in this example.
 The initial time point should avoid 0 to prevent computational error in misconduct.
-The \code{outnames}, \code{dllname}, are based on the arguments from the \code{ode} function in \pkg{deSolve} package.
-The details of the model structure and these arguments are defined in \code{pbtk1comp.c}.
-and \code{pbtk1comp_inits.R} files.
+The \code{outnames} is based on the arguments from the \code{ode} function in \pkg{deSolve} package.
 
@@
 outputs <- c("Ccompartment", "Ametabolized")
-out <- solve_fun(x, time = t, initState = initState, 
-                 outnames = outputs, dllname = mName)
+out <- solve_fun(x, time = t, func = pbtk1cpt, initState = initState, outnames = outputs)
 ```
@@
 In addition to using \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can also be used, and provides better computational efficiency.
+To start, the users need to install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
 To solve ODEs through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
+The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.
+
+```{r, eval=T}
+pbtk1cpt_model()
+cat(readLines("pbtk1cpt.model"), sep = "\n")
+```
+
 The computing time of using \code{solve_fun} in SA is estimated as,
@@
 ```{r, eval=T}
-system.time(out<-solve_fun(x, t, initState = initState, 
-                           outnames = Outputs, dllname = mName))
+system.time(out <- solve_fun(x, time = t, 
+                             func = pbtk1cpt, initState = initState, 
+                             outnames = outputs))
 ```
@@
 ```{r, eval=T, message=F, warning=F}
+mName <- "pbtk1cpt"
 compile_model(mName, application = "mcsim")
@@
 system.time(out <- solve_mcsim(x, mName = mName, params = params, 
-                               vars = Outputs, time = t, 
+                               vars = outputs, time = t, 
                                condition = conditions))
@@
 
-After solving the equations under the same given condition, we can find that \pkg{GNU MCSim} has about 4 - 5 times faster in computing performance than using \pkg{deSolve}.
+After solving the equations under the same given condition, we can find that \pkg{GNU MCSim} has more than 50 times faster in computing performance than using \pkg{deSolve}.
 In this case, we only focus on performing the global SA alone for generic PK model without additional comparison with experimental data.

### 2019-08-12 manuscript.Rmd
manuscript

@@
 \pkg{pksensi} allows users select the preferred method to solve the PK model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
-Running model under \pkg{GNU MCSim} native code can have a faster speed to obtain the model outputs.
 This section mainly focus on how to conduct global SA with \pkg{deSolve} package with pure \proglang{R} environment.
@@
 
-The one-compartment PK model was mainly includes two state variables, which are the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
+The one-compartment PK model can describe the quantity of compound in the gut lumen (\code{Agutlument}) and the rest of body (\code{Acompartmant}).
 The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.

### 2019-08-12 manuscript.Rmd
manuscript

@@
 This section mainly focus on how to conduct global SA with \pkg{deSolve} package with pure \proglang{R} environment.
-Then compare the computing time with \pkg{GNU MCSim}.
 
@@
 The \code{Ametabolized} is the quantity of compound transform and metabolize through hepatic clearance.
-The \code{AUC} is the calculated area under the curve of the rest body ($mg \cdot h/L$).
 
@@
 
-The parameter values and initial states need to assign to the specific condition.
+The parameter values and initial states need to assign to the specific value before simulation.
 We used the corresponding parameter value of acetaminophen (APAP) in this example.
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model.'}
+```{r, eval=T, fig.height=4, fig.width=4, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model.'}
 plot(y)
@@
 The \code{BW} is assumed to a normal distribution with mean = 60 kg and sd = 5 kg.
-The number of sample size determines the robustness of the result of SA.
-Higher number of sample size lead to narrower confidence intervals for sensitivity measurements across different replications.
-However, it will take a longer time in computation.
-Here we use a sample size of 200 with 20 replications.
 
@@
 
-Through \code{rfast99} function, a S3 object with class 'rfast99' will be created. The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
+Here we use a sample size of 200 with 10 replications.
+Through \code{rfast99} function, a S3 object with class 'rfast99' will be created.
+The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
+The number of sample size determines the robustness of the result of SA.
+Higher number of sample size lead to narrower confidence intervals for sensitivity measurements across different replications.
+However, it will take a longer time in computation.
 
@@
 The sample number is 200 with 4 model parameters, which generates 800 model evaluations.
-The replication is set to 20.
-Therefore, the total of 16,000 parameter sequence will be used to compute the corresponding outputs.
+The replication is set to 10.
+Therefore, the total of 8,000 parameter sequence will be used to compute the corresponding outputs.
 Figure \ref{fig:rep-sampl} plotted the sampling process for each parameter from the first 3 replications.
@@
 outputs <- c("Ccompartment", "Ametabolized")
-out <- solve_fun(x, time = t, func = pbtk1cpt, initState = initState, outnames = outputs)
+out <- solve_fun(x, time = t, func = pbtk1cpt, 
+                 initState = initState, outnames = outputs)
 ```
@@
 In addition to using \pkg{deSolve} to solve differential equations in PK model, the \pkg{GNU MCSim} can also be used, and provides better computational efficiency.
-To start, the users need to install \pkg{GNU MCSim} [@JSSv002i09] and related standard \proglang{C} compiler to solve the ODE function.
+To start, the users need to install \pkg{GNU MCSim} and related standard \proglang{C} compiler to solve the ODE function.
 To solve ODEs through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
@@
 
-We initiated \pkg{pksensi} in order to facilitiate a comprehensive and efficient global SA workflow in PK modeling, filling a key gap for open source modeling communities.
+We initiated \pkg{pksensi} in order to facilitate a comprehensive and efficient global SA workflow in PK modeling, filling a key gap for open source modeling communities.
 It provides a straightforward application to investigate the impact of parameters on model outputs across simulation time-points and variables.
@@
 
-We chose to integrate sensitivity analysis with \pkg{GNU MCSim} because it is a powerful open-source simulation software package for dynamical simulation of biological-based models, mainly built for PK research with probabilistic approaches [@bois2009gnu].
+We chose to integrate sensitivity analysis with \pkg{GNU MCSim} because it is a powerful open-source software package for dynamical simulation of biological-based models, mainly built for PK research with probabilistic approaches [@bois2009gnu].
 In addition, the current version of \pkg{pksensi} includes both \pkg{GNU MCSim} and \pkg{deSolve} as the main ODE solvers.

### 2019-08-13 manuscript.Rmd
update

@@
 Therefore, the alternative method to prevent this issue is to detect the crucial parameter range that causes the problem.
-Also, setting the maximum number of steps to higher value instead of using the default value (500) in \pkg{GNU MCSim} can prevent this problem (internally defined).
-The maximum number of step is set to 5000 in this case.
+Also, setting the maximum number of steps to higher value in \pkg{GNU MCSim} can prevent this problem (internally defined).
+The maximum number of step is set to 5000 (default) in this case.
+The value can reset through reinstall the the \pkg{GNU MCSim}.
 Here we separate the global SA of APAP-PBPK model process to several steps.
-The value can reset through reinstall the the \pkg{GNU MCSim} by using \code{mcsim_install(mxstp = 5000)}.
 
@@
 
-In addition, through using \code{check}, the sensitivity and convergence indices can be preliminary checked for all output variables.
 Based on our previous study, we proposed the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.

### 2019-08-13 manuscript.Rmd
manuscript v3

@@
 output: rticles::jss_article
-#output: word_document
+# output: word_document
 ---
@@
 
-```{r, eval=T, fig.height=4, fig.width=4, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model.'}
+```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model.'}
 plot(y)
@@
 To start, the users need to install \pkg{GNU MCSim} and related standard \proglang{C} compiler to solve the ODE function.
-To solve ODEs through \pkg{GNU MCSim}, we need to change the argument to \code{application = mcsim} in \code{compile_model}.
 The following \proglang{R} script will download the one-compartment PK model from https://github.com/nanhung/pksensi/blob/master/tests/pbtk1cpt.model.
@@
 Then, before we conduct the SA through \pkg{GNU MCSim}, The following code is used to compile the \pkg{GNU MCSim} model code to the executable program.
+To solve ODEs through \pkg{GNU MCSim}, we need to use \code{compile_model} with argument \code{application = mcsim}.
 

### 2019-08-25 manuscript.Rmd
manuscript

@@
 Pharmacokinetic models describe the changes in the concentrations or amounts of a substance in the body over time (the various terms for the kinetic models, including "pharmacokinetic", "toxicokinetic", and "biokinetic" models, are used exchangeably here).
-The goal of SA in pharmacokinetic research is to examine the sensitivity of output variables,such as chemical concentration in blood or tissues, with respect to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
+The goal of SA in pharmacokinetic research is to examine the sensitivity of output variables, such as chemical concentration in blood or tissues, with respect to input parameters, such as anatomical, physiological, and kinetic constants [@mcnally2011workflow].
 In addition, this SA approach can guide experimental design and the parameter estimation processes [@zhang2015sobol].
@@
 We used \pkg{GNU MCSim} [@JSSv002i09], an effective simulation package for Bayesian population PBPK modeling, to calibrate the model.
-We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based global SA algorithm, had the best balance of efficiency and accuracy for a complex, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
-This method was first introduced in @cukier1973study (the classic FAST method with first-order effect only) in a study of the sensitivity of coupled reaction systems that are constructed by differential equations with the numerous rate coefficients.
+We found that the extended Fourier Amplitude Sensitivity Test (eFAST), a type of variance-based global SA algorithm, had the best balance of efficiency and accuracy for a sophisticated, multi-compartment, multi-dataset, and multi-metabolite PBPK model.
+This method was first introduced in @cukier1973study (the classic FAST method with the first-order effect only) in a study of the sensitivity of coupled reaction systems that are constructed by differential equations with the numerous rate coefficients.
 It then evolved to Sobol' sensitivity measure in @saltelli1999quantitative (the eFAST method with first- and total-order effects).
 The FAST algorithm creates the multidimensional space of input parameters through a "search-curve".
-It's more computationally efficient in calculating the influence of model parameter than other global SA methods such as the Monte Carlo based approaches [@jansen1999analysis; @owen2013better].
+It is more computationally efficient in calculating the influence of model parameter than other global SA methods such as the Monte Carlo based approaches [@jansen1999analysis; @owen2013better].
 Our previous work also found some efficient visualization approaches that can be used to distinguish between “influential” and “non-influential” parameters.
@@
 However, there are no software packages that provide the available tools in the assessment of reliability.
-In addition, the existing eFAST function does not includes the feature to generate the random sampling curve.
+In addition, the existing eFAST function does not include the feature to generate the random sampling curve.
 We present here an \proglang{R} package, called \pkg{pksensi}, which is designed to make SA more accessible and reproducible in pharmacological and toxicological researches.
@@
 Moreover, it seamlessly integrates with both general ODE solvers available in R or as external \proglang{C} code, as well as the \pkg{GNU MCSim} software used for conducting sophisticated PK modeling, including MCMC.
-The analytical framework in \pkg{pksensi} not only includes global SA but also the uncertainty analysis and diagnostic tools to support the modeling process.
+The analytical framework in \pkg{pksensi} includes not only global SA but also the uncertainty analysis and diagnostic tools to support the modeling process.
 The design concepts of \pkg{pksensi} are:
@@
 *	Cross-platform: Models can run on Windows/MacOS/Linux
-*	Freely available: All related packages are free and open source
+*	Freely available: All related packages are free and open-source
 *	Integrated application: Users can run pharmacokinetic models in \proglang{R} with script that were written in \proglang{C} or \pkg{GNU MCSim} 
@@
 
-We adopted eFAST, a widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimensional parameters in numerical models in this package.
+We adopted eFAST, the widely used analysis of variance (ANOVA)-like global SA approach to decompose the variance to partial variances for multidimensional parameters in numerical models in this package.
 Most of the available eFAST functions, such as the algorithm to generate the parameter space and to set the sampling frequency, were sourced from the \proglang{R} \pkg{sensitivity} package [@R-sensitivity].
@@
 where $x_i$ is the nominal value of the $i$-th parameter, and $\omega_i$ is a vector giving the set of frequencies, one frequency for each parameter, and $\varphi$ is a random phase-shift coefficient ranged from 0 to $2\pi$.
-The default set of frequencies is based on the study from @saltelli1999quantitative.
+The default set of frequencies is according to the study from @saltelli1999quantitative.
 Through the random phase-shift, the robust result of the sensitivity measurement should be similar across replications under the same sample size.
@@
 This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed SI for all parameters across all time-points and output variables.
-Through random phase shift procedure, the replicate SI values can be repeatedly sampled independently.
+Through random phase shift procedure, the replicated SI values can repeatedly sample independently.
 Therefore, they can be used to investigate the convergence of SI as,
@@
 The R package \pkg{pksensi} provides a useful function to solve ODEs in pharmacokinetic and PK models using analytical and numerical approaches.
-The solution can be performed under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package [@JSSv033i09].
+The solution can perform under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package [@JSSv033i09].
 However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as \proglang{FORTRAN}, \proglang{C}, or \proglang{C++}.
@@
 However, Windows users need to install Rtools to compile the source code by using the GNU GCC compiler.
-More requirements are detailed in @soetaertr.
+More requirements were detailed in @soetaertr.
 The \proglang{C} implementation of the example model can be found as an example within \pkg{GNU MCSim} [@JSSv002i09].
@@
 The uncertainty analysis is a crucial modeling process within SA.
-Sometimes, it's difficult to have informative data to set up the parameter distribution.
+Sometimes, it is difficult to have informative data to set up the parameter distribution.
 A wider range for a parameter distribution might cause a numerical error at extreme values.
@@
 Before fitting the data to model, the Monte Carlo can be applied with the given distribution of model parameter to examine whether the experiment data within the range of in the distribution of model output.
-\pkg{pksensi} also has a function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built in \pkg{GNU MCSim} and exam the simulation result.
-The general workflow of uncertainty and sensitivity analysis is illustrated in Figure \ref{fig:Workflow}.
+\pkg{pksensi} also has a function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built-in \pkg{GNU MCSim} and exam the simulation result.
+The general workflow of uncertainty and SA is illustrated in Figure \ref{fig:Workflow}.
 
@@
 This format is particularly suited for further graphical routines within \pkg{pksensi}.
-Also, time-dependent sensitivity measurements of first and total order effects for all parameters can be plotted simultaneously.
-If the computed SI show different values across replications under the same sample size, this indicates inadequate sample number, which can create an unreliable results.
+Also, time-dependent sensitivity measurements of first and total effects for all parameters can be plotted simultaneously.
+If the computed SI show different values across replications under the same sample size, this indicates an inadequate sample number, which can create an unreliable result.
 It is essential to use a sufficient sample size to prevent incorrect judgments in parameter fixing.
@@
 
-In this section, two PK models were applied to demonstrate how \pkg{pksensi} works.
-The first one is a generic model and the second is a chemical-specific model.
-The following code shows the version of the platform, version of \proglang{R} and \pkg{pksensi} that use to conduct the current example.
+In this section, two PK models will apply to demonstrate how \pkg{pksensi} works.
+The first one is a generic model, and the second is a chemical-specific model.
+The following code shows the version of the platform, version of \proglang{R}, and \pkg{pksensi} that use to conduct the current example.
 We use \pkg{GNU MCSim} v6.1.0 in these examples.
@@
 
-In the first example, we use a generic, one-compartment PK model from \pkg{httk} package [@JSSv079i04] to demonstrate how \pkg{pksensi} can be applied to PK studies.
+In the first example, we use a generic, one-compartment PK model from \pkg{httk} package [@JSSv079i04] to demonstrate how \pkg{pksensi} can apply to PK studies.
 The differential equations for this model is as,
@@
 
-The concentration of the chemical in the rest of body and blood ($C_{rest}$, mg/L) can be calculated as,
+The concentration of the chemical in the rest of body and blood ($C_{rest}$, mg/L) can calculate as,
 
@@
 where $V_{dist}$ is the volume of distribution (L/kg BW) and $BW$ is the body weight (kg).
-The $C_{rest}$ can also be seen as the chemical concentration in plasma that can be further used to compare with observed results in a pharmacokinetic experiment.
+The $C_{rest}$ can also represent as the chemical concentration in plasma that can be further used to compare with observed results in a pharmacokinetic experiment.
 The bioavailability is assumed to be 100% in this model.
@@
 \pkg{pksensi} allows users select the preferred method to solve the PK model, either with the \pkg{deSolve} package or with \pkg{GNU MCSim} through the compile function.
-This section mainly focus on how to conduct global SA with \pkg{deSolve} package with pure \proglang{R} environment.
+This section mainly focuses on how to conduct global SA with \pkg{deSolve} package with pure \proglang{R} environment.
 
@@
 The parameter values and initial states need to assign to the specific value before simulation.
-We used the corresponding parameter value of acetaminophen (APAP) in this example.
+Here, we use the corresponding parameter value of acetaminophen (APAP) in this example.
 These model parameters are derived from the *in-vivo* or *in-vitro* experiment results.
@@
 
-We conduct global SA for the four parameters in one-compartment PK model to investigate the parameter impact on the plasma concentration during the 24-hr time period post intake.
+We conduct global SA for the four parameters in one-compartment PK model to investigate the parameter impact on the plasma concentration during the 24-hr time post intake.
 The distribution of parameter is taken to be uniform with bounds corresponding to 50% and 200% of the nominal value.
@@
 
-The generated parameters are stored as a 3-D array under the named \code{a}, with the dimension of sample size, number of replications, and number of parameters, respectively.
+The generated parameters are stored as a 3-D array under the named \code{a}, with the dimension of sample size, the number of replications, and the number of parameters, respectively.
 
@@
 
-The sample number is 200 with 4 model parameters, which generates 800 model evaluations.
+The sample number is 200, with 4 model parameters, which generates 800 model evaluations.
 The replication is set to 10.
@@
 
-Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} are dominating the plasma concentration before and after 5-hr post chemical intake, respectively.
+Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} dominate the plasma concentration before and after 5-hr post chemical intake, respectively.
 Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-However, the current result is only based on the setting distribution of model parameters for APAP.
+However, the current result only based on the setting distribution of model parameters for APAP.
 The different given input conditions (e.g., range of parameter uncertainty, chemical-dependent parameter value) can change the result (result not shown).
@@
 Same as \code{Ccompartment}, the \code{kgutabs} contribute about 30 - 40% variation of model output in the first hour.
-The \code{BW} is the least important parameter in the current analysis, and therefore can be fixed in the model fitting to data and additional applications.
+The \code{BW} is the least important parameter in the current analysis, and therefore, can be fixed in the model fitting to data and additional applications.
 
-In addition to use time-SI profile to investigate the parameter impact on model output, we can construct the relationship between parameter and the corresponding value of model output in the examination.
+Also, to use time-SI profile to investigate the parameter impact on model output, we can construct the relationship between parameter and the corresponding value of model output in the examination.
 
@@
 The output variable \code{out} containing all the input arguments detailed before and the calculated SI of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
-The convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
-The output are formatted as 4-D array in \code{y} with the dimension name of model evaluation, number of replications, number of time points, and number of output variables, respectively.
+Convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
+The outputs are formatted as 4-D array in \code{y} with the dimension name of model evaluation, number of replications, number of time points, and number of output variables, respectively.
 
@@
 
-Some functions in \pkg{pksensi} provide an efficient way to check the result from global SA.
-The \code{check} can determine which parameters have relatively lower sensitivity measurement across the given time points and model outputs, and therefore can be applied parameter fixing in model calibration.
+Some functions in \pkg{pksensi} provide efficient ways to check the result from global SA.
+The \code{check} can determine which parameters have relatively lowered sensitivity measurement across the given time points and model outputs, and therefore can be applied parameter fixing in model calibration.
 The \code{check} also provides some feasible argument to specify the target output or change the cut-off value.
@@
 
-The computing time of using \code{solve_fun} in SA is estimated as,
+The computing time of using \code{solve_fun} in SA estimate as,
 
@@
 
-Then, before we conduct the SA through \pkg{GNU MCSim}, The following code is used to compile the \pkg{GNU MCSim} model code to the executable program.
+Then, before we conduct the SA through \pkg{GNU MCSim}, The following codes are used to compile the \pkg{GNU MCSim} model code to the executable program.
 To solve ODEs through \pkg{GNU MCSim}, we need to use \code{compile_model} with argument \code{application = mcsim}.
@@
 
-Similar to \code{solve_fun} function that can define the initial value of parameters and state variable through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as exposure dose, fixed parameter value or initial condition of state variables.
+Similar to \code{solve_fun} function that can define the initial values of parameters and state variable through generated functions, the \code{solve_mcsim} also has a \code{condition} argument that can be used to give the specific input value such as exposure dose, nominated parameter value, or initial condition of state variables.
 
@@
 
-The aim of this section is to apply global SA for APAP-PBPK model through \pkg{pksensi}.
+This section aims to apply global SA for APAP-PBPK model through \pkg{pksensi}.
 The model describes the pharmacokinetics of parent APAP and its two metabolites glucuronide (APAP-G) and sulfate (APAP-S).
-The model code is included in this package in https://github.com/nanhung/pksensi/blob/master/tests/pbpk_apap.model and can be generated through \code{pbpk_apap_model} function.
-We applied the global SA workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
-The descriptions of each parameter and the sampling ranges are list in Table \ref{tab:data}.
+The model code is included in this package in https://github.com/nanhung/pksensi/blob/master/tests/pbpk_apap.model and can generate through \code{pbpk_apap_model} function.
+We apply the global SA workflow to the original published model with 21 model parameters [@s13318-015-0253-x].
+The descriptions of each parameter and the sampling range are list in Table \ref{tab:data}.
 
@@
 To rapidly reach the acceptance convergence in this example, we apply uniform distribution for all testing parameters.
-The ranges of informative parameters are set to 1.96-times difference for the single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
-The nominal values of informative model parameters were defined as:
+The ranges of informative parameters set to 1.96-times difference for the single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
+The nominal values of informative model parameters define as:
 
@@
 
-Generally, wide range of parameter value might cause the computing error when solving the differential equation.
+Generally, a wide range of parameter value might cause the computing error when solving the differential equation.
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
@@
 Therefore, the alternative method to prevent this issue is to detect the crucial parameter range that causes the problem.
-Also, setting the maximum number of steps to higher value in \pkg{GNU MCSim} can prevent this problem (internally defined).
-The maximum number of step is set to 5000 (default) in this case.
-The value can reset through reinstall the the \pkg{GNU MCSim}.
+Also, setting the maximum number of steps to the higher value in \pkg{GNU MCSim} can prevent this problem (internally defined).
+The maximum number of step is 5000 (default) in this case.
+The value can reset through reinstall the \pkg{GNU MCSim}.
 Here we separate the global SA of APAP-PBPK model process to several steps.
@@
 
-The model code needs to be prepared in the following global SA workflow.
+The model code needs to prepare in the following global SA workflow.
 After creating the \code{pbpk_apap.model} file in the working directory, the next step is to generate the executable program (\code{mcsim.pbpk_apap}) through the \code{compile_model} function.
@@
 
-To optimize the computing speed, this case only uses \pkg{GNU MCSim} to estimate the concentration of APAP and its metabolites in plasma.
+In this case, we only use \pkg{GNU MCSim} to estimate the concentration of APAP and its metabolites in plasma to optimize the computing speed.
 The oral dose of APAP is 20 mg/kg in this example.
@@
 Based on our previous study, we proposed the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.
-Through the given argument \code{order}, we can select the specific order of sensitivity measurement that we're interested in (Figure \ref{fig:Heat-SI}).
+Through the given argument \code{order}, we can select the specific order of sensitivity measurement that we are interested in (Figure \ref{fig:Heat-SI}).
 
@@
 The sample size of convergence in the current PBPK model is 8,192 [@fphar201800588].
-However, based on the current sample size we still can find 6 parameters that can be an important parameter for the plasma APAP concentration.
+However, based on the current sample size, we still can find 6 parameters that can be an important parameter for the plasma APAP concentration.
 
@@
 
-Computational modeling has become an important technique in modern scientific research to support hypotheses, experiments, and discoveries.
+Computational modeling has become an essential technique in modern scientific research to support hypotheses, experiments, and discoveries.
 Uncertainty and SA are essential tools for modelers in the conduct of model development and examination.
@@
 
-We initiated \pkg{pksensi} in order to facilitate a comprehensive and efficient global SA workflow in PK modeling, filling a key gap for open source modeling communities.
+We initiated \pkg{pksensi} in order to facilitate a comprehensive and efficient global SA workflow in PK modeling, filling a crucial gap for open-source modeling communities.
 It provides a straightforward application to investigate the impact of parameters on model outputs across simulation time-points and variables.
@@
 In the future version, we expect to add additional global SA approaches [@glen2012estimating; @ring2017identifying].
-The other features such as principle component analysis [@lamboni2011multivariate] will also be considered.
+The other features, such as principle component analysis [@lamboni2011multivariate] will also be considered.
 
@@
 Other factors, such as the number of parameters examined, the parameter distributions/ranges, and the given dosing level in PK model might also have a potential impact on the level of importance of each parameter, but these "meta-uncertainties" are not currently addressed.
-The complex models with a high dimensionality in the the parameter is also a challenge in global SA as they have higher computer demands.
+The complex models with high dimensionality in the parameter is also a challenge in global SA as they have higher computer demands.
 As approaches are developed to address these complexities, we hope to incorporate them into \pkg{pksensi}.
 
-Overall, \pkg{pksensi} provides a comprehensive suite features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model development and calibration.
+Overall, \pkg{pksensi} provides full suite features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model development and calibration.
 Although \pkg{pksensi} used here is mainly for pharmacokinetic modeling, it can also be applied to other ODE-based dynamic models in order to investigate the sensitivity of model outputs to input parameters.

### 2019-08-25 manuscript.Rmd
manuscript

@@
 For instance, we recently applied a global SA workflow to reduce the computational burden in the Bayesian Markov Chain Monte Carlo-based calibration process of a physiologically based pharmacokinetic (PBPK) model.
-Although several SA algorithms and software packages are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, run sensitivity analyses, visualize SA results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
-Therefore, we developed an open-source \proglang{R} package, called **pksensi**, to fill this need and make SA more accessible in pharmacological and toxicological research.
+Although several SA algorithms and software packages are available, no comprehensive software package exists that allows users to seamlessly solve the PBPK model differential equations, conduct and visualize SA results, and discriminate between the “non-influential” model parameters that can be fixed and those that need calibration.
+Therefore, we developed an \proglang{R} package, called **pksensi**, to fill this need and make SA more accessible in pharmacological and toxicological research.
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, including those with multivariate model outputs, such as multiple time-points and tissue compartments.
-It also refined the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
+It also refines the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
 Furthermore, pksensi includes functions to check the convergence of the SA results.
@@
 The \pkg{FME} package contains a basic method for performing local SA for dynamic models [@JSSv033i03].
-It can also conduct collinearity and identifiability analysis to evaluate which model parameters can be estimated or exclude based on available observation data.
+It can also conduct collinearity and identifiability analysis to evaluate which model parameters can estimate or exclude based on available observation data.
 The \pkg{multisensi} package provides SA for models with multivariate output [@R-multisensi].
@@
 This type of global SA approach and SI also know as Sobol' method and Sobol' index.
-Since the estimated SI is not stable (high variation) when using smaller sample sizes, we included a random phase-shift approach to conduct replicating sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
+Since the estimated SI is not stable (high variation) when using smaller sample size, we included a random phase-shift approach to conduct replicating sampling from random starting points across parameter space to test the convergence and robustness of the sensitivity measurement.
 Considering the model that is built under $y=f(x_{i})$.
-The sampling scheme can be written as,
+The sampling scheme can describe as,
 
@@
 The default set of frequencies is according to the study from @saltelli1999quantitative.
-Through the random phase-shift, the robust result of the sensitivity measurement should be similar across replications under the same sample size.
+Through the random phase-shift, the robust result of the sensitivity measurement should be similar across each replication under the same sample size.
 For most Sobol'-based SA algorithm, the larger sample size is needed to reduce the probability of having negative estimates. However, FAST estimates are always positive according to its construction.
@@
 This method quantitatively assesses convergence by computing the width of 95% confidence intervals of computed SI for all parameters across all time-points and output variables.
-Through random phase shift procedure, the replicated SI values can repeatedly sample independently.
+Through random phase-shift procedure, the replicated SI values can repeatedly sample independently.
 Therefore, they can be used to investigate the convergence of SI as,
@@
 The R package \pkg{pksensi} provides a useful function to solve ODEs in pharmacokinetic and PK models using analytical and numerical approaches.
-The solution can perform under the pure \proglang{R} programming environment by linking pksensi with the \pkg{deSolve} package [@JSSv033i09].
+The solution can perform under the pure \proglang{R} programming environment by linking \pkg{pksensi} with the \pkg{deSolve} package [@JSSv033i09].
 However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as \proglang{FORTRAN}, \proglang{C}, or \proglang{C++}.
@@
 
-In this section, two PK models will apply to demonstrate how \pkg{pksensi} works.
-The first one is a generic model, and the second is a chemical-specific model.
-The following code shows the version of the platform, version of \proglang{R}, and \pkg{pksensi} that use to conduct the current example.
-We use \pkg{GNU MCSim} v6.1.0 in these examples.
-
-```{r}
-sessionInfo()$R.version$version.string
-```
-
-```{r}
-sessionInfo()$otherPkgs$pksensi$Version
-```
+In this section, two PK models will apply to demonstrate how \pkg{pksensi} works. The first one is a generic model, and the second is a chemical-specific model.
 
@@
 
-```{r, eval=T, fig.height=5, fig.cap = '\\label{fig:rep-sampl}The parameter search curve generated by eFAST method.
+```{r, eval=T, fig.height=5, fig.cap = '\\label{fig:rep-sampl}The parameter search curves generated by eFAST method.
 The x-axis is the sample number and y-axis is the parameter value for \\code{vdist}, \\code{ke}, \\code{kgutabs}, and \\code{BW} (top to down), respectively.
@@
 The time-dependent SI are shown in Figure \ref{fig:pk-solvefun-rest}.
-The SI has computed range from 0 (no impact) to 1 (high impact) and represent the contribution percentage of output variance under the given parameter distributions.
+The SI has computed range from 0 (no impact) to 1 (high impact) and represents the contribution percentage of output variance under the given parameter distributions.
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-rest}Time-dependent SI of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.
-The solid line represent the total (black) and first (red) order SI with 95\\% confidence interval (polygon).
-The dash line is the cut-off with default value of 0.05.'}
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-rest}Time-dependent SI of the plasma concentration estimated from the one-compartment PK model during 24-hr time period intake.
+The solid line represents the total (black) and first (red) order SI with 95\\% confidence interval (polygon).
+The dashed line is the cut-off with the default value of 0.05.'}
 plot(out)
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent SI of the plasma concentration estimated from one-compartment PK model during 24-hr time period intake.'}
+```{r, eval=T, fig.cap = '\\label{fig:pk-solvefun-Amet}Time-dependent SI of the plasma concentration estimated from the one-compartment PK model during 24-hr time period intake.'}
 plot(out, vars = "Ametabolized")
@@
 
-Figure \ref{fig:SI-output} show the concentration of rest body (\code{Ccompartment}) and the model parameters.
+Figure \ref{fig:SI-output} shows the concentration of rest body (\code{Ccompartment}) and the model parameters.
 We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameters have high impact on the modeling result.
 The \code{ke} shows the high correlation at the later time period (t = 24.01 h).
-The \code{BW} doesn't show any obvious relationship with \code{Ccompartment}.
+The \code{BW} does not show any obvious relationship with \code{Ccompartment}.
 
@@
 
-The computing time of using \code{solve_fun} in SA estimate as,
+The computing time of using \code{solve_fun} in SA is estimated as,
 
@@
 Previously, the probability distributions of model parameters were set to either truncated normal or uniform distributions when the parameters have informative prior information or not, respectively.
-To rapidly reach the acceptance convergence in this example, we apply uniform distribution for all testing parameters.
+To rapidly reach the acceptable convergence in this example, we apply uniform distribution for all testing parameters.
 The ranges of informative parameters set to 1.96-times difference for the single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
@@
 Km_AS <- log(2.29e4)
-
 rng <- 1.96 
@@
 Generally, a wide range of parameter value might cause the computing error when solving the differential equation.
-One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in a lower value.
+One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in the lower levels.
 The \code{generate_infile} and \code{solve_mcsim} provide the arguments of \code{rtol} and \code{atol} that adjust the error tolerance to prevent the unwanted error.
@@
 The 21 tested model parameters are defined in this step, including parameter names and probability distributions.
-To prevent errors, the range of \code{SULT_VmaxC} and \code{UGT_VmaxC} needed to be adjusted from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
+To prevent errors, the ranges of \code{SULT_VmaxC} and \code{UGT_VmaxC} need to be adjusted from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
 The objects \code{q} and \code{dist} are set to the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
@@
 The Monte Carlo simulation is run for 1000 iterations in the assignment of \code{monte_carlo}.
-The input file ('sim.in') and output file ('simmc.out') will be generated under the standard ASCII format.
+The input file (`sim.in`) and output file (`simmc.out`) will be generated under the standard ASCII format.
 
@@
 
-```{r, fig.cap = '\\label{fig:pksim-MC}Coverage checks of prior PBPK model predictions with calibrated APAP data', fig.height=3.5, fig.width=9}
+```{r, fig.cap = '\\label{fig:pksim-MC}Coverage checks of prior PBPK model predictions with calibrated APAP data.', fig.height=3.5, fig.width=9}
 par(mfrow = c(1,3), mar = c(4,4,1,1))
@@
 
-Based on our previous study, we proposed the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.
+Based on our previous study, we propose the heatmap visualization approach \code{heat_check} to distinguish "influential" and "non-influential" parameters with a "cut-off" point.
 Through the given argument \code{order}, we can select the specific order of sensitivity measurement that we are interested in (Figure \ref{fig:Heat-SI}).
 
-```{r, eval=T, fig.cap = '\\label{fig:Heat-SI}Heat map representation of time-dependent total SI computed for global SA method', fig.height=5}
+```{r, eval=T, fig.cap = '\\label{fig:Heat-SI}Heat map representation of time-dependent total SI computed in global SA method.', fig.height=5}
 heat_check(out, order = "total order")
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:Heat-CI}Heat map representation of time-dependent total CI computed for global SA method', fig.height=5}
+```{r, eval=T, fig.cap = '\\label{fig:Heat-CI}Heat map representation of time-dependent total CI computed in global SA method.', fig.height=5}
 heat_check(out, index = "CI", order = "total order")

### 2019-09-11 manuscript.Rmd
manuscript

@@
 This method was first introduced in @cukier1973study (the classic FAST method with the first-order effect only) in a study of the sensitivity of coupled reaction systems that are constructed by differential equations with the numerous rate coefficients.
-It then evolved to Sobol' sensitivity measure in @saltelli1999quantitative (the eFAST method with first- and total-order effects).
+It then evolved to estimate the Sobol' sensitivity measure in @saltelli1999quantitative (the eFAST method with first- and total-order effects).
 The FAST algorithm creates the multidimensional space of input parameters through a "search-curve".
@@
 Most of SA tools are constructed using scientific computing software, such as \proglang{GNU Octave}/\proglang{Matlab} [@pianosi2015matlab], \proglang{python} [@herman2017salib], and \proglang{R} [@iooss2017introduction].
-The \proglang{R} software is widely used and had numerous developed packages, such as \pkg{sensitivity} [@R-sensitivity], which is a practical tool to conduct local and global SA.
+The \proglang{R} software is widely used and includes numerous developed packages, such as \pkg{sensitivity} [@R-sensitivity], which is a practical tool to conduct local and global SA.
 The \pkg{sensitivity} package includes some functions to generate the parameter sequences by using different computing algorithms.
@@
 The uncertainty analysis using Monte Carlo is an appropriate approach in the data-driven modeling process to address this problem.
-Before fitting the data to model, the Monte Carlo can be applied with the given distribution of model parameter to examine whether the experiment data within the range of in the distribution of model output.
-\pkg{pksensi} also has a function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis which is built-in \pkg{GNU MCSim} and exam the simulation result.
+Before fitting the data to model, the Monte Carlo can be applied with the given distribution of model parameter to examine whether the experimental data are within the range of the distribution of model output.
+\pkg{pksensi} also has a function that can link to \pkg{GNU MCSim} to conduct Monte Carlo analysis, which is built-into \pkg{GNU MCSim} and examine the simulation result.
 The general workflow of uncertainty and SA is illustrated in Figure \ref{fig:Workflow}.
@@
 where $V_{dist}$ is the volume of distribution (L/kg BW) and $BW$ is the body weight (kg).
-The $C_{rest}$ can also represent as the chemical concentration in plasma that can be further used to compare with observed results in a pharmacokinetic experiment.
+$C_{rest}$ can also represent as the chemical concentration in plasma that can be further used to compare with observed results in a pharmacokinetic experiment.
 The bioavailability is assumed to be 100% in this model.
@@
 
-The parameter values and initial states need to assign to the specific value before simulation.
+The parameter values and initial states need to be assigned to specific values before simulation.
 Here, we use the corresponding parameter value of acetaminophen (APAP) in this example.
@@
 
-```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model.'}
+```{r, eval=T, fig.cap = '\\label{fig:ode-PK}Simulation results of one-compartment PK model for fixed default parameters.'}
 plot(y)
@@
 The \code{set.seed} can use to reproduce the same parameter matrix in the random sampling.
-The number of sample size determines the robustness of the result of SA.
+The sample size determines the robustness of the result of SA.
 Higher number of sample size lead to narrower confidence intervals for sensitivity measurements across different replications.
@@
 Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} dominate the plasma concentration before and after 5-hr post chemical intake, respectively.
-Besides, the \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
-However, the current result only based on the setting distribution of model parameters for APAP.
-The different given input conditions (e.g., range of parameter uncertainty, chemical-dependent parameter value) can change the result (result not shown).
+The \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
+However, the current result only based on the distribution of model parameters for APAP.
+Given different input conditions (e.g., range of parameter uncertainty, chemical-dependent parameter value) the result can of course change (result not shown).
 
@@
 
-Also, to use time-SI profile to investigate the parameter impact on model output, we can construct the relationship between parameter and the corresponding value of model output in the examination.
+In addition to using the time-SI profile to investigate the parameter impact on model output, we can directly examine the relationship between parameters and model output graphically.
 
@@
 
-Figure \ref{fig:SI-output} shows the concentration of rest body (\code{Ccompartment}) and the model parameters.
-We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, representing that the parameters have high impact on the modeling result.
-The \code{ke} shows the high correlation at the later time period (t = 24.01 h).
+Figure \ref{fig:SI-output} shows the relationship between the concentration of the rest of the body (\code{Ccompartment}) and the model parameters at different time points.
+We can find that \code{kgutabs} and \code{vdist} have higher correlation with \code{Ccompartment} in the beginning (t = 0.01 h) of post-intake duration compared with other parameters, suggesting that the parameters have high impact on the modeling result.
+The \code{ke} shows a high correlation at the later time period (t = 24.01 h).
 The \code{BW} does not show any obvious relationship with \code{Ccompartment}.
 
-The output variable \code{out} containing all the input arguments detailed before and the calculated SI of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
+The output variable \code{out} containing all the input arguments detailed above and the calculated SI of first order (\code{mSI}), interaction (\code{iSI}), and total order (\code{tSI}).
 Convergence indices are also stored in the list named \code{mCI}, \code{iCI}, and \code{tCI}.
@@
 The \code{check} can determine which parameters have relatively lowered sensitivity measurement across the given time points and model outputs, and therefore can be applied parameter fixing in model calibration.
-The \code{check} also provides some feasible argument to specify the target output or change the cut-off value.
-The argument of \code{SI.cutoff} is setting at 0.05 to detect the relative non-influential parameters as default, representing the 5% change of the output is contributed from the specific parameter variation.
+The \code{check} also provides an argument to specify the target output or change the cut-off value.
+The argument of \code{SI.cutoff} set for example at 0.05, is used to detect the relative non-influential parameters as default, in this case representing a 5% change of the output is contributed from the specific parameter variation.
 
@@
 However, all parameters do not converge to the setting cut-off, which means the larger sample size is required in further sensitivity testing.
-Similar to the \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to exam the given output.
+Similar to the \code{plot} function that can assign specific output variable in the examination, the \code{check} function can also use the assignment (\code{vars}) to examine a given output.
 
@@
 Previously, the probability distributions of model parameters were set to either truncated normal or uniform distributions when the parameters have informative prior information or not, respectively.
-To rapidly reach the acceptable convergence in this example, we apply uniform distribution for all testing parameters.
+To rapidly reach the acceptable convergence in this example, we use a uniform distribution for all testing parameters.
 The ranges of informative parameters set to 1.96-times difference for the single side under log-scaled (approximate 54.6 times difference between minimum and maximum in natural scaled).
@@
 One of the effective ways to prevent this problem is to adjust the value of relative and absolute error tolerance to control the error appearance by resetting these parameters in the lower levels.
-The \code{generate_infile} and \code{solve_mcsim} provide the arguments of \code{rtol} and \code{atol} that adjust the error tolerance to prevent the unwanted error.
+The \code{generate_infile} and \code{solve_mcsim} functions provide arguments of \code{rtol} and \code{atol} that adjust the error tolerance to prevent the unwanted error.
 However, the modification will decrease the computing speed.
@@
 The maximum number of step is 5000 (default) in this case.
-The value can reset through reinstall the \pkg{GNU MCSim}.
+The value can reset through reinstallation of \pkg{GNU MCSim}.
 Here we separate the global SA of APAP-PBPK model process to several steps.
@@
 To prevent errors, the ranges of \code{SULT_VmaxC} and \code{UGT_VmaxC} need to be adjusted from $U(0, 15)$ [@s13318-015-0253-x] to $U(0, 10)$ [@fphar201800588].
-The objects \code{q} and \code{dist} are set to the type of distribution that will use to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
+The objects \code{q} and \code{dist} are set to the type of distribution that will be used to generate the parameter matrix in \pkg{GNU MCSim} (for uncertainty analysis) and \proglang{R} (for SA).
 
@@
 
-We apply uncertainty analysis through the \code{solve_mcsim} and visualize the result by \code{pksim} function.
+We apply uncertainty analysis through the \code{solve_mcsim} function and visualize the results using \code{pksim} function.
 Some example data are included in \pkg{pksensi} with experiment time (h) and concentration (mg/L).
@@
 To conduct the global SA with \pkg{GNU MCSim} and \pkg{pksensi}, an input file needs to be generate before modeling.
-The file can create by \code{generate_infile} function.
+The file can be created by the \code{generate_infile} function.
 Alternatively, \code{solve_mcsim} can also automatically create the input file and compute the output.
@@
 
-The plotting function can create the result of time-dependent sensitivity measurement to determine the parameter impact on model output over time (Figure \ref{fig:APAP-sensi}).
+The plotting function can show the results of time-dependent sensitivity to determine the parameter impacts on model output over time (Figure \ref{fig:APAP-sensi}).
 
@@
 We also developed functions to visualize the output results and help distinguish the "influential" and "non-influential" parameters that can be applied in parameter fixing in the model calibration.
-In the future version, we expect to add additional global SA approaches [@glen2012estimating; @ring2017identifying].
-The other features, such as principle component analysis [@lamboni2011multivariate] will also be considered.
+In future versions, we expect to add additional global SA approaches [@glen2012estimating; @ring2017identifying].
+Other features, such as principle component analysis [@lamboni2011multivariate] will also be considered.
 
@@
 Other factors, such as the number of parameters examined, the parameter distributions/ranges, and the given dosing level in PK model might also have a potential impact on the level of importance of each parameter, but these "meta-uncertainties" are not currently addressed.
-The complex models with high dimensionality in the parameter is also a challenge in global SA as they have higher computer demands.
+Complex models with high paremeter dimensionality are also challenging in global SA as they have higher computer demands.
 As approaches are developed to address these complexities, we hope to incorporate them into \pkg{pksensi}.
 
-Overall, \pkg{pksensi} provides full suite features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model development and calibration.
+Overall, \pkg{pksensi} provides comprehensive suite of features and functions for performing global SA for PK models and can create robust and reproducible results for decision making in model development and calibration.
 Although \pkg{pksensi} used here is mainly for pharmacokinetic modeling, it can also be applied to other ODE-based dynamic models in order to investigate the sensitivity of model outputs to input parameters.

### 2019-09-11 manuscript.Rmd
manuscript

@@
 This package can investigate both parameter uncertainty and sensitivity in pharmacokinetic models, including those with multivariate model outputs, such as multiple time-points and tissue compartments.
-It also refines the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
+We also refines the extended Fourier Amplitude Sensitivity Test method for SA by adding a random phase-shift to analyze the statistical variability of the sensitivity index for each model parameter.
 Furthermore, pksensi includes functions to check the convergence of the SA results.
@@
 Sensitivity analysis (SA) is a mathematical technique to investigate how variations in model parameters affect model outputs [@saltelli2008global].
-An increasing number of studies use SA approach to determine which model parameters contribute to high variation in model predictions [@ferretti2016trends].
+An increasing number of studies use SA to determine which model parameters contribute to high variation in model predictions [@ferretti2016trends].
 This technique has also been applied in pharmacology and toxicology research [@loizou2015application; @mcnally2012reconstruction].
@@
 Through the random phase-shift, the robust result of the sensitivity measurement should be similar across each replication under the same sample size.
-For most Sobol'-based SA algorithm, the larger sample size is needed to reduce the probability of having negative estimates. However, FAST estimates are always positive according to its construction.
+For most Sobol'-based SA algorithms, the larger sample size is needed to reduce the probability of having negative estimates. However, FAST estimates are always positive according to its construction.
 
@@
 However, the most efficient way to achieve high computational speed is through compiled, lower-level languages, such as \proglang{FORTRAN}, \proglang{C}, or \proglang{C++}.
-The \pkg{pksensi} includes a function to compile and create dynamic-link libraries (.dll) on Windows and shared objects (.so) on Unix-liked systems (e.g., Linux and MacOS).
+The \pkg{pksensi} package includes a function to compile and create dynamic-link libraries (.dll) on Windows and shared objects (.so) on Unix-liked systems (e.g., Linux and MacOS).
 This compiled program can load and execute through the build-in function in \proglang{R}.
@@
 Here, we can see (Figure \ref{fig:pk-solvefun-rest}) that \code{vdist} and \code{ke} dominate the plasma concentration before and after 5-hr post chemical intake, respectively.
-The \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
+The parameter \code{kgutabs} only plays a crucial role to determine the plasma concentration in the first hour.
 However, the current result only based on the distribution of model parameters for APAP.
@@
 The \code{ke} shows a high correlation at the later time period (t = 24.01 h).
-The \code{BW} does not show any obvious relationship with \code{Ccompartment}.
+The parameter \code{BW} does not show any obvious relationship with \code{Ccompartment}.
 
